---
title: "4.1 County Level Person Weights"
author: "Suman Mitra"
date: "September 1, 2015"
output: html_document
---

## Overview 

This document explains how county-level weights are developed using the R statistical software for people who participated in the 2012 CHTS. The purpose of these weights is to scale up information provided by respondents to the population of each county based on the socio-economic variables listed below. The person level weights were developed at the county level. Users of these weights are cautioned that these weights should not be applied to lower-level geographies such as cities or zip codes. Similar weights can be developed for these geographies if they contain enough CHTS respondents.

## Methodology

Person-level county weights are calculated by raking at the person level. These weights adjust the relative importance of responses to reflect the different probabilities of selecting respondents, and align the sample distributions to population distributions based on the 2010 Census. In particular, person-level weights were adjusted so that the sum of these weights equals known population totals for certain subgroups of the population of each county.  Variables used for raking at the person level are:

  * Hispanic Status (Hispanic, Non-Hispanic) - County-wide distribution;
  * Ethnicity (White, African American, Asian, Other) - County-wide distribution;
  * Age (less than 20 years, 20 - 24 years, 25 - 34 years, 35 - 54 years, 55 - 64 years, 65 years or older) - County -wide distribution;
  * Employment Status (Part-time or full-time Employed, Not-employed)- County-wide distribution; and
  * Gender (Male, Female)- County-wide distribution

After the raking procedure, very large weights are capped to be no more than five times the mean of weight. To check whether the new weights represent the total population of the counties, we perform some random checks for different variables and the results show that the new weights align with the county populations.

We use this new county wide person weight to calculate the final  trip weight which is an outcome of Trip Correction Factor (TCF) * New County Wide Person Weight. 


## R Code and Implementation

The following commands are used to generate county wide person weight.


### Generate Person level weight

First we need to import the person data file and corresponding lookup file (containing location information) and then we merge them.

```{r pw-1,echo=FALSE}
if ( !exists(rd) ) { rd <- "/home/crindt/chts-book" }
p <- function(ps) { return(paste(rd,ps,sep="/")) }
```

```{r pw-2,cache=TRUE}
library(chts2011)
library(dplyr)
# Set the working directory
setwd(p("Weight/PersonWeight_County"))
# Import the person level variables csv file and give a new name
#persondata <- read.csv("c:/Users/user/Documents/Suman_research/Caltrans Project/Data/deliv_per.csv", header=TRUE)
persondata <- chts_per
# Import the lookup variable csv file and give a new name (containing x, y coordinate data)
#lookup_hh <- read.csv("c:/Users/user/Documents/Suman_research/Caltrans Project/Data/lookup_hh.csv", header=TRUE)
lookup_hh <- chts_hh %>% mutate(HCTFIPx=CTFIP)
lookup_hh$HCTFIP <- as.integer(lookup_hh$HCTFIPx) %% 1000
# Merging two data set to get the county information
perdata2<-merge(persondata,lookup_hh, by ="SAMPN", all = TRUE)
# Names of the new merged file
names(perdata2)
```

Now we have new merged file which is used to generate new person weight. We need to recode the raking variables to make them similar to the census data (Here we are using American Community Survey, 2010 five year estimate data). 

First we will assign zero for 'NA' observations of raking variables because the code we will use later for raking can't read 'NA'. 

```{r pw-3,cache=TRUE}
# Assign 0 for all NA observations
perdata2$HISP[is.na(perdata2$HISP)] <- 0
perdata2$AGE[is.na(perdata2$AGE)] <- 0
perdata2$RACE1[is.na(perdata2$RACE1)] <- 0
perdata2$EMPLY[is.na(perdata2$EMPLY)] <- 0
```

Then, we change the value of hispanic variable to make it similar to census data. 

```{r pw-4,cache=TRUE}
# For changing values of Hispanic
perdata2$HISPn<-ifelse(perdata2$HISP>2,2,perdata2$HISP)

# Show values
head(perdata2$HISPn,10)
```

Next, we recode the age and race variable.

```{r pw-5,cache=TRUE}
# For recoding AGE
perdata2$AGEn<-ifelse(perdata2$AGE<20,1,ifelse((perdata2$AGE>=20 & perdata2$AGE<=24),2,ifelse((perdata2$AGE>=25 & perdata2$AGE<=34),3,ifelse((perdata2$AGE>=35 & perdata2$AGE<=54),4,ifelse((perdata2$AGE>=55 & perdata2$AGE<=64),5,6)))))
# Show values
head(perdata2$AGE,10)

# For recoding race
perdata2$RACEn<-ifelse(perdata2$RACE1==1,1,ifelse(perdata2$RACE1==2,2,ifelse(perdata2$RACE1==4,3,4)))
# Show values
head(perdata2$RACEn,10)
```


Here, we recode employmnet (EMPLY) and Gender variable.

```{r pw-6,cache=TRUE}
# For recoding EMPLY
perdata2$EMPLYn<-ifelse(perdata2$EMPLY==1,1,ifelse(perdata2$EMPLY==2,2,ifelse(perdata2$EMPLY==8,1,2)))
# Show values
head(perdata2$EMPLYn,10)

# For recoding Gender
perdata2$GENDn<-ifelse(perdata2$GEND==1,1,2)
# Show the data
head(perdata2$GENDn,10)
```

Now we have data file (perdata2) which has similar coding  as census data for raking variables. 
We need county wide total population data which we collect from census and save it as csv (we organize them as per our requriment).Here we import that file and merge this data with perdata2. 


```{r pw-7,cache=TRUE}
# Assgin county wide total population into perdata2 file
# Import the countywise total population csv file and give a new name
tot.pop <- read.csv(p("Weight/Data_County/Total/county_totpop2.csv"), header=TRUE)
head(tot.pop)

# Merging two data set
perdata3<-merge(perdata2,tot.pop, by ="HCTFIP", all = TRUE)
names(perdata3)
```

So, now we have a new file named perdata3 which contains all the raking variables informaiton and county wide total population. This is a single file for all counties. 
But we need a single dataframe for each county. So we need to create a list of seperate dataframe for each county.


```{r pw-8,cache=TRUE}
# create list of dataframes for different counties 
percnty_list <- split(perdata3, as.factor(perdata3$HCTFIP))
# Show the list of one county
head(percnty_list[[11]])
```

We need census data of raking variables (Hispanic, Race, Employment status, Gender and Age). We collect this information from census and organize them as per our requriment and save as csv file. Then we import them and create seperate dataframe for each county (the csv file contain informaiton of all the counties).

First, we do it for hispanic data.

```{r pw-9,cache=TRUE}
# Hispanic data
hisp.county <- read.csv(p("Weight/Data_County/Hispanic/Hisp_county.csv"), header=TRUE)
head(hisp.county)

# Generate a list of seperate datafrmaes containing hispanic data for each county
hisp.county$id <- 1:nrow(hisp.county)
hisp.data <- lapply(split(hisp.county, hisp.county$id), function(x){data.frame(HISPn = c(1,2),  Freq  = c(x$Hispanic, x$Non.Hispanic))                                                           
}
)
# Show
head(hisp.data[1])
```

Here, we import the race data.

```{r pw-10,cache=TRUE}
# Race
race.county <- read.csv(p("Weight/Data_County/Race/Race_county.csv"), header=TRUE)
race.county$id <- 1:nrow(race.county)
race.data <- lapply(split(race.county, race.county$id), function(x){data.frame(RACEn = c(1:4),  Freq  = c(x$White, x$African.American,x$Asian,x$Others))                                                             
}
)
head(race.data[1])
```

Then, we repeat the process for employment data.

```{r pw-11,cache=TRUE}
## Employment data
emply.county <- read.csv(p("Weight/Data_County/Employment/Employment_county.csv"), header=TRUE)

emply.county$id <- 1:nrow(emply.county)
emply.data <- lapply(split(emply.county, emply.county$id), function(x){data.frame(EMPLYn = c(1,2),  Freq  = c(x$Employed, x$Unemployed))                                                             
}
)
head(emply.data[1])
```

Finally, we import gender and age data.

```{r pw-12,cache=TRUE}
#Gender
gend.county <- read.csv(p("Weight/Data_County/Gender/Gender_county.csv"), header=TRUE) 

gend.county$id <- 1:nrow(gend.county)
gend.data <- lapply(split(gend.county, gend.county$id), function(x){data.frame(GENDn = c(1,2),  Freq  = c(x$Male, x$Female))                                                             
}
)
head(gend.data[1])

# Age
age.county <- read.csv(p("Weight/Data_County/Age/Age_county.csv"), header=TRUE)
head(age.county)
age.county$id <- 1:nrow(age.county)
age.data <- lapply(split(age.county, age.county$id), function(x){data.frame(AGEn = c(1:6),  Freq  = c(x$x1, x$x2,x$x3,x$x4,x$x5,x$x6))                                                             
}
)
head(age.data[48])
```


Here we want to set the directory where we want to save our output.

```{r pw-13,cache=TRUE}
# Set directory where we want to save
setwd(p("Weight/PersonWeight_County"))
```


As our data is ready for raking, we proceed for raking next. We  do it in a for loop function which generate weight for each county. To do this we need to install the library "survey". First we design a survey using 'svydesign' command and we use sampling without replacement (pps ="brewer") option. The 'fpc (finite population correction)' is the ratio of sample population and county wide total population. Then we perform the raking. After raking, we trim the weight so that very large weights were capped to be no more than five times the mean of weight.The following codes also give us informaiotn about the errors (the fip code of counties for which the weight can't be generated). The generated expanded person weights for different counties are saved in our specified directory. 


```{r pw-14,cache=TRUE}
# Raking for all county to create new expanded weight
#Need library "survey"
library (survey)
# Loop function
for (i in 1:length(hisp.data)){ #length must be equal to 58=total number of county
  tryCatch({ #If there is any error it will give error message but still continue the loop 
  f=seq(6001, 6115, by=2)
  s<-as.character(f[i])
  name=c('perweight_',s, '.CSV')
  name=paste(name,collapse="")
  temp1<- svydesign(id=~1, weights=~PERWGT,data=percnty_list[[i]], fpc=~fpc,pps="brewer") # Survey design 
  temp3<- rake(temp1, list(~RACEn,~HISPn,~AGEn,~EMPLYn,~GENDn), list(race.data[[i]],hisp.data[[i]],age.data[[i]],emply.data[[i]],gend.data[[i]]),control = list(maxit = 100)) # Raking
  temp4<-trimWeights(temp3,lower=(min(weights(temp3))),upper=(5*mean(weights(temp3))))# Trim the weights (max value no greater than 5 times the mean)
  temp5<-cbind(temp4$variables[1:3],weights(temp4))# Combine weight with sample no and person no
  colnames(temp5)[4] <- "NEXPPERWGT" # change the column name of weight variable
  write.table(temp5, file =name,sep=",", row.names=FALSE, col.names=TRUE) # Save as csv file
  }, error=function(e){cat("ERROR :",conditionMessage(e), "\n",finally=print(name))}) # Show the errors and the file names
}
```

After running the codes, we get nine errors for nine county data sets and we need to treat them indiviudally (HCTFIP= 3,15,21,27,49,51,91,103,105). The errors show that some race categories are absent from sample of these counties, so we don't include race as raking variable for these counties in the following codes.

```{r pw-15,cache=TRUE}
# Loop function for the rest of the counties
d=c(2,8,11,14,25,26,46,52,53) # # only for counties which gave error for previous cases (HCTFIP=3(i=2);HCTFIP=15(i=8);HCTFIP=21(i=11) and so on)
for (i in d){ 
  tryCatch({ 
    f=seq(6001, 6115, by=2)
    s<-as.character(f[i])
    name=c('perweight_',s, '.CSV')
    name=paste(name,collapse="")
    temp1<- svydesign(id=~1, weights=~EXPPERWGT,data=percnty_list[[i]], fpc=~fpc,pps="brewer") # Survey design 
    # Raking
    temp3<- rake(temp1, list(~HISPn,~AGEn,~EMPLYn,~GENDn), list(hisp.data[[i]],age.data[[i]],emply.data[[i]],gend.data[[i]]),control = list(maxit = 100)) 
    temp4<-trimWeights(temp3,lower=(min(weights(temp3))),upper=(5*mean(weights(temp3))))# Trim the weights (max value no greater than 5 times the mean)
    temp5<-cbind(temp4$variables[1:3],weights(temp4))# Combine weight with sample no and person no
    colnames(temp5)[4] <- "NEXPPERWGT" # change the column name of weight variable
    write.table(temp5, file =name,sep=",", row.names=FALSE, col.names=TRUE) # Save as csv file
  }, error=function(e){cat("ERROR :",conditionMessage(e), "\n",finally=print(name))})# Show the errors and the file names
}
```

There is no error now. So, we have expanded person weights for all 58 counties and these weights are saved in the directory we specified earlier. We need to combine these weights of 58 counties into one file. The following codes are used to import and combine them.


```{r pw-16,cache=TRUE}
# Import multiple csv file of new expanded weight and combine into one file
setwd(p("Weight/PersonWeight_County"))
# Import the files
filenames <- list.files(path = p("Weight/PersonWeight_County"))
# combined the files into one
nperweight<-do.call("rbind", lapply(filenames, read.csv, header = TRUE)) 
head(nperweight)
```

Now we have on file which contains new county wide expanded person weight for all counties. We want to generate new county wide person weight (not expanded) from our expanded weight. We know that person weight = Expanded weight * fpc (finite population correction) and fpc = sample population/county total population. 

First, we merge the fpc data with expanded new person weight.

```{r pw-17,cache=TRUE}
# fpc =sample population/county total population
# Subest the fpc
fpc<-subset(perdata3,select=c(SAMPN,PERNO,fpc))
head(fpc)

# Merge fpc with expanded weight
nperweight<-merge(fpc,nperweight, by =c("SAMPN","PERNO"), all = TRUE)
names(nperweight)
```

Here, we generate the unexpanded new person weight.

```{r pw-18,cache=TRUE}
# Unexpanded person weight= Expanded weight* fpc
nperweight$NPERWGT<-nperweight$NEXPPERWGT*nperweight$fpc
head(nperweight)
```

Now, we subset the weight data and exclude the fpc variable (since we don't need it).

```{r pw-19,cache=TRUE}
# Subset the dataset (excluding the fpc variable)
nperweight2<-subset(nperweight,select=-c(3))
names(nperweight2)

# Reorder the variables
nperweight2 <- nperweight2[,c(1,2,5,4,3)]
head(nperweight2)
```

Now we have a file with new county wide expaned and unexpanded person weights. Here we save it in our desired directory.

```{r pw-20,cache=TRUE}
# Save 
setwd(p("Weight/NewWeightData"))
write.csv(nperweight2, file="nperweight922015.csv",row.names = FALSE) # save as csv
```

We want to merge the new weight with person data file.

```{r pw-21,cache=TRUE}
# Merge the new weight file with original person data file
npersondata<-merge(persondata,nperweight2, by =c("SAMPN","PERNO"), all = TRUE)
head(npersondata)

write.csv(npersondata, file="npersondata922015.csv",row.names=FALSE)
```

Now we have a person data file with new county wide weight. 
As we mentioned earlier we want check whether the new weights represent the total population of the counties. So we perform some random checks for different variables using following codes. We do it for Mendocino county as an example.

```{r pw-22,cache=TRUE}
# Random Check
setwd(p("Weight/Check"))
# We do it for Mendocino county (FIP code=45)
popdata_45<-subset(npersondata,HCTFIP==45)
head(popdata_45$HCTFIP)

# For race variable
race<-as.data.frame(table(popdata_45$RACE1))
race$Percent<-race$Freq*100/sum(race$Freq)
race
wrace<-aggregate(NEXPPERWGT~RACE1,FUN=sum,data=popdata_45)
wrace$Percent<-wrace$NEXPPERWGT*100/sum(wrace$NEXPPERWGT)
wrace<- rbind(wrace, c("Total", colSums(wrace[,2:3],na.rm = TRUE)))
wrace
```

Here, we do a random check for education variable.

```{r pw-23,cache=TRUE}
# For education
educ<-as.data.frame(table(popdata_45$EDUCA))
educ$Percent<-educ$Freq*100/sum(educ$Freq)
educ
weduc<-aggregate(NEXPPERWGT~EDUCA,FUN=sum,data=popdata_45)
weduc$Percent<-weduc$NEXPPERWGT*100/sum(weduc$NEXPPERWGT)
weduc<- rbind(weduc,c("Total",colSums(weduc[,2:3],na.rm = TRUE)))
weduc
```

The total population of Mendocino county is 87,487 (based on census). Our random check shows that the new weights align with the total population of Mendocino county for both variables. 


Now we need to calculate the new final  trip weight using the new county wide person weight. To do this, first, we need to import the place level file and merge it with new weight file.

```{r pw-24,cache=TRUE}
#Import the place level data  csv file and give a new name
#place_data1 <- read.csv("../Data/deliv_place.csv", header=TRUE)
place_data1 <- chts_pla

# Merge with new weight
nplacedata<-merge(place_data1,nperweight2, by =c("SAMPN","PERNO"), all = FALSE)
names(nplacedata)
```

The final trip weight is an outcome of Trip Correction Factor (TCF) * New county wide Person Weight. The follow codes are used to calculate the final trip weight and the expanded final trip weight for all counties.

```{r pw-25,cache=TRUE}
# Generate Final trip weight (TCF*NPERWGT) 
nplacedata$NTCFPERWGT<-nplacedata$TCF*nplacedata$NPERWGT
# Generate Final expanded trip weight (TCF*NEXPPERWGT) 
nplacedata$NEXPTCFPERWGT<-nplacedata$TCF*nplacedata$NEXPPERWGT
# show 
head(nplacedata)
```

If we want to save the new file with new trip weight as csv file, we can do it by using following codes.

```{r pw-26,cache=TRUE}
# Set the working directory
setwd(p("Weight/NewWeightData"))

write.csv(nplacedata, file="nplacedata932015.csv",row.names=FALSE)
# Now we have new place data file with new weight
```

Now we have new place data file with new trip weight for all counties.
