---
title: "4.2 County Level Household Weights"
author: "Suman Mitra"
date: "September 2, 2015"
output: html_document
---

```{r hw-setup-0,echo=FALSE,cache=TRUE}
if ( !exists("rd") ) { rd <<- gsub("(.*?chts-book[^/]*).*$","\\1",getwd()) } 
source(paste(rd,"R/setup-hooks.R",sep="/"))
write(paste("Using Root Directory",rd,"\n"),stderr())
```

This document explains how to develop county-level household weights using
the R statistical software for households who participated in the 2012
CHTS. The purpose of these weights is to scale up information provided
by respondents to the population of each county based on the
socio-economic variables listed below. Users of county-level household weights are
cautioned that these weights should not be applied to lower-level
geographies such as cities or zip codes. Similar weights can be
developed for these geographies using the approach described below if they contain enough CHTS
respondents.

# Methodology

Household-level county weights are calculated by raking at the
household level. These weights adjust the relative importance of
responses to reflect the different probabilities of selection of CHTS
respondents, and they align sample distributions to population
distributions based on the 2010 Census. In particular, household
weights were adjusted so that the sums of adjusted weights are equal
to known population totals for some subgroups of the population of
each county.  Following NuStats' work for Caltrans, the variables used for
raking at the household level are:

  * Household size (1, 2, 3, 4 or more) - County wide distribution;
  * Household income (Less than $24,999, $25,000 - $49,999, $50,000 - $74,999, $75,000-$99,999, $100,000-$149,999, $150,000 or above) - County wide distribution;
  * Total number of workers in the household (0, 1, 2, 3 or more) - County wide distribution; and
  * Total number of vehicles in the household (0, 1, 2, 3 or more) - County wide distribution.

After the raking procedure, very large weights are capped to be no
more than five times the mean of weights. To check whether the new
weights represent the total population of a county, we perform
some random checks for different variables. The results show that
the new weights align well with the population of the different counties.


\ifmydocumentation

# R Code and Implementation

The following commands are used to generate countywide household weights.

First we import the person data frame and the corresponding lookup
data frame (which contains location information) and then we merge them:

```{r hw-2,cache=TRUE}
# Set the working directory

setwd(pd("Weight/HHWeight_County"))
#Import the household level variables data frame and give a new name
library(chts2011)
library(chts2011pvt)
hhdata <- chts_hh
lookup_hh <- chts_lu_hh

#Merging two data set and giving a new name
hhdata2<-merge(hhdata,lookup_hh, by ="SAMPN", all = TRUE)

# Subset the dataset (excluding the duplicate variables)
nhhdata<-subset(hhdata2,select=-c(HCITY.x,HZIP.x,HXCORD.x,HYCORD.x,HPrimaryCity.x))
# Rename the new added variables
names(nhhdata)[c(76,77,78,79,83)]<-c("HCITY","HZIP","HXCORD","HYCORD","HPrimaryCity")
names(nhhdata)
```

Next, we force `HCTFIP` values to be integers:
```{r hw-3}
nhhdata$HCTFIP <- as.integer(nhhdata$HCTFIP) %% 1000
```

Now we have a new merged file that can be used to generate new household
weights. We need to recode the raking variables to make them similar to
the census data. Here we are using American Community Survey, 2010
five year estimate data.

First we assign zero to `NA` entries of the raking variables
because the code we use below for raking cannot read `NA`:

```{r hw-4,cache=TRUE}
# Assign 0 for all NA observations
nhhdata[is.na(nhhdata$INCOM)] <- 0
nhhdata[is.na(nhhdata$HHSIZ)] <- 0
nhhdata[is.na(nhhdata$HHVEH)] <- 0
nhhdata[is.na(nhhdata$HHEMP)] <- 0
```

Second, we transform the household income (`INCOM`) variable to make
it similar to the corresponding census data variable.

```{r hw-5,cache=TRUE}
# Recoding income variable (INCOM)
nhhdata$INCOME<-ifelse(nhhdata$INCOM<=2,1,ifelse((nhhdata$INCOM>=3 &nhhdata$INCOM<=4),2,ifelse(nhhdata$INCOM==5,3,ifelse(nhhdata$INCOM==6,4,ifelse(nhhdata$INCOM==7,5,6)))))
# Show the values
head(nhhdata$INCOME)
```

Third, we recode the household size (`HHSIZE`) variable:

```{r hw-6,cache=TRUE}
# Recoding household size
nhhdata$HHSIZE<-ifelse(nhhdata$HHSIZ>=4,4,nhhdata$HHSIZ)
# Show values
head(nhhdata$HHSIZE,10)
```

next, we recode the number of household vehicles (`HHVEH`) variable:

```{r hw-7,cache=TRUE}
# Recoding household vehicle(HHVEH)
nhhdata$HHVEHC<-ifelse(nhhdata$HHVEH>=3,3,nhhdata$HHVEH)
# Show values
head(nhhdata$HHVEHC,10)
```

Finally, we work on the number of household workers (`HHEMP`):

```{r hw-8,cache=TRUE}
# Recoding number of workers (HHEMP)
nhhdata$HHWRK<-ifelse(nhhdata$HHEMP>=3,3,nhhdata$HHEMP)
# See the data
head(nhhdata$HHWRK,10)
# Show the names
names(nhhdata)
```

Now we have a data frame (`nhhdata`) that is coded like the census
data for our raking variables.  To proceed, we need countywide total household data
which we collect from the census. We save those data in a CSV file. 
Let us first import that file and merge these data with `nhhdata`:


```{r hw-9,cache=TRUE}
# Assgin county wide total household into nhhdata file
#Import the county wide total hosuehold csv file and give a new name
tot.hh <- read.csv(pd("Weight/Data_County/Total/county_tothh2.csv"), header=TRUE)
# Merging two data set
hhdata3<-merge(nhhdata,tot.hh, by ="HCTFIP", all = TRUE)
names(hhdata3)
head(hhdata3$fpc)
```

Now we have a new data frame named `hhdata3` that contains all the
raking variables and countywide total household information in a single file for all counties.
Howeverm we need separate dataframes for each county so we create a list of separate dataframes for
each county:


```{r hw-10,cache=TRUE}
# create list of dataframes for different counties 
hhcnty_list <- split(hhdata3, as.factor(hhdata3$HCTFIP))
# Show the list of one county
head(hhcnty_list[[2]])
```

We also need census data for our raking variables (Household size, Household
income, Number of household vehicles and Number of household
workers). We collect this information from the census, organize it, and save it in a CSV file. Then we import this
file and create separate dataframes for each county:

First, let us process the Household size variable:

```{r hw-11,cache=TRUE}
# Household size data
hhsize.county <- read.csv(pd("Weight/Data_County/County_HH/HHSize/hhsize.csv"),
                          header=TRUE)
head(hhsize.county)
# Generate a list of seperate datafrmaes containing household size variable
# for each county
hhsize.county$id <- 1:nrow(hhsize.county)
hhsize.data <- lapply(
  split(hhsize.county, hhsize.county$id),
  function(x){
    data.frame(HHSIZE = c(1:4),
               Freq  = c(x$HS1, x$HS2,x$HS3,x$HS4))                                                      
  }
)
# Show
head(hhsize.data[1])
```

Second, let us import the Household income data:

```{r hw-12,cache=TRUE}
# Income
hhincome.county <- read.csv(pd("Weight/Data_County/County_HH/Income/hhincome.csv"),
                            header=TRUE)
head(hhincome.county)
hhincome.county$id <- 1:nrow(hhincome.county)
hhincome.data <- lapply(
  split(hhincome.county, hhincome.county$id),
  function(x){
    data.frame(INCOME = c(1:6),
               Freq  = c(x$INC1, x$INC2,x$INC3,x$INC4,x$INC5,x$INC6))
  }
)
head(hhincome.data[1])
```

Third, repeat the same process for the number of household vehicles:

```{r hw-13,cache=TRUE}
## Number of vehicles data
hhveh.county <- read.csv(pd("Weight/Data_County/County_HH/HHVeh/hhvehicle.csv"),
                         header=TRUE)
head(hhveh.county)
hhveh.county$id <- 1:nrow(hhveh.county)
hhveh.data <- lapply(
  split(hhveh.county, hhveh.county$id),
  function(x){
    data.frame(HHVEHC = c(0:3),  Freq  = c(x$Veh0, x$Veh1,x$Veh2,x$Veh3))
  }
)
head(hhveh.data[2])
```

Finally, import the number of household workers:

```{r hw-14,cache=TRUE}
## Number of workers
hhwrk.county <- read.csv(pd("Weight/Data_County/County_HH/Worker/hhworkers.csv"),
                         header=TRUE)
head(hhwrk.county)
hhwrk.county$id <- 1:nrow(hhwrk.county)
hhwrk.data <- lapply(
  split(hhwrk.county, hhwrk.county$id),
  function(x){
    data.frame(HHWRK = c(0:3),  Freq  = c(x$W0, x$W1,x$W2,x$W3))
  }
)
head(hhwrk.data[3])
```

At this point, let us set the directory where we want to save our output:

```{r hw-15,cache=TRUE}
# Set directory where we want to save
setwd(pd("Weight/HHWeight_County"))
```


Our data are ready for raking. Next we create a loop to generate weights for each county. However, we first need to
install the library "survey" and design a survey using the `svydesign` command with sampling without replacement
(pps ="brewer"). The option 'fpc (finite population correction)' is the ratio of sample household number to
countywide total households. Once these steps have been completed, we perform the raking procedure. After raking, we
trim the weights so large weights are capped to be no more than five times the mean of weights. 
The code below also generates information about errors (the FIP code of counties for which the weight cannot
be generated). Finally, the generated expanded household weights for different
counties are saved in a selected directory.


```{r hw-16,cache=TRUE}
# Raking to create new expanded weight for all counties
#Need library "survey"
suppressMessages(library (survey))
# Loop function
# NOTE: length must be equal to 58=total number of county
for (i in 1:length(hhwrk.data)){ 
  tryCatch({
    # If there is any error it will give error message but still
    # continue the loop 
    f=seq(6001, 6115, by=2)
    s<-as.character(f[i])
    name<-paste('HHweight_',s, '.CSV',sep="")
    name<-pd(paste("Weight/HHWeight_County",name,sep="/"))

    # Survey design (Need library "survey")
    temp1<- svydesign(
      id=~1,
      weights=~HHWGT,
      data=hhcnty_list[[i]],
      fpc=~fpc,
      pps="brewer") 

    # Raking
    temp3<- rake(temp1,
                 list(~HHSIZE,~INCOME,~HHVEHC,~HHWRK),
                 list(hhsize.data[[i]],hhincome.data[[i]],
                      hhveh.data[[i]],hhwrk.data[[i]]),
                 control = list(maxit = 1000)) 

    # Trim the weights (max value no greater than 5 times the mean)
    temp4<-trimWeights(temp3,
                       lower=(min(weights(temp3))),
                       upper=(5*mean(weights(temp3))))

    # Combine weight with sample no and hh no
    temp5<-cbind(temp4$variables[1:2],weights(temp4))

    # change the column name of weight variable
    colnames(temp5)[3] <- "NEXPHHWGT" 

    # Save as csv file
    write.table(temp5, file=name,sep=",",
                row.names=FALSE, col.names=TRUE) 

  }, error=function(e){
    # Show the errors and the file names
    cat("ERROR :",conditionMessage(e), "\n",finally=print(name))
  }) 
}
```

After running the code above, we get errors for two counties
(`HCTFIP={3,43}`). The errors show that some `HHVEHC` (number of household vehicles) categories are absent
from the sample for these counties, so we don't include `HHVEHC` as a raking
variable for these two counties in the codes below.

```{r hw-17,cache=TRUE}
# Loop function for the rest of the counties
# only for counties which gave error for previous cases
# (HCTFIP=3[i=2],HCTFIP= 43[i=22])
e=c(2,22) 
for (i in e){ 
  tryCatch({
    f=seq(6001, 6115, by=2)
    s<-as.character(f[i])
    name=paste('HHweight_',s, '.CSV',sep="")
    name=pd(paste("Weight/HHWeight_County",name,sep="/"))

    # Survey design (Need library "survey")
    temp1<- svydesign(
      id=~1,
      weights=~EXPHHWGT,
      data=hhcnty_list[[i]],
      fpc=~fpc,pps="brewer") 

    # Raking
    temp3<- rake(temp1, list(~HHSIZE,~INCOME,~HHWRK),
                 list(hhsize.data[[i]],hhincome.data[[i]],
                      hhwrk.data[[i]]),
                 control = list(maxit = 1000)) 

    # Trim the weights (max value no greater than 5 times the mean)
    temp4<-trimWeights(temp3,lower=(min(weights(temp3))),
                       upper=(5*mean(weights(temp3))))

    # Combine weight with sample no and hh no
    temp5<-cbind(temp4$variables[1:2],weights(temp4))

    # change the column name of weight variable
    colnames(temp5)[3] <- "NEXPHHWGT" 

    # Save as csv file 
    write.table(temp5, file =name,sep=",", row.names=FALSE, col.names=TRUE) 

  }, error=function(e){
    # Show the errors and the file names
    cat("ERROR :",conditionMessage(e), "\n",finally=print(name))
  }) 
}

```

There is no more error now. We have generated expanded household weights for all
58 counties and these weights are saved in the directory specified
earlier. We now combine the weights for the 58 counties into one file:


```{r hw-18,cache=TRUE}
# Import multiple csv file of new weight and combine into one file
setwd(pd("Weight/HHWeight_County"))

# Import the files
filenames <- list.files(path = pd("Weight/HHWeight_County"))

# combined the files into one
nhhweight<-do.call("rbind", lapply(filenames, read.csv, header = TRUE)) 

head(nhhweight)
```

Now we have a file that contains new countywide expanded household
weights for all 58 counties. Let us generate new countywide household
weights (not expanded) from our expanded weights, using that `household
weight` = `Expanded household weight` * `fpc` (finite population
correction) and `fpc` = `sample household`/`county total household`.

First, we merge the `fpc` data with expanded new household weights:

```{r hw-19,cache=TRUE}
# fpc =sample household number/county total household number

# Subset the fpc
fpc2<-subset(hhdata3,select=c(SAMPN,fpc))
head(fpc2)

# Merge fpc with expanded weight
nhhweight<-merge(fpc2,nhhweight, by ="SAMPN", all = TRUE)

names(nhhweight)
```

Next, we generate unexpanded new household weighs:

```{r hw-20,cache=TRUE}
# Unexpanded household weight= Expanded weight* fpc
nhhweight$NHHWGT<-nhhweight$NEXPHHWGT*nhhweight$fpc
head(nhhweight)
```

Now, we subset the weight data and exclude the fpc variable since we
don't need it:

```{r hw-21,cache=TRUE}
# Subset the dataset (excluding the fpc variable)
nhhweight2<-subset(nhhweight,select=-c(2))
head(nhhweight2)

# Reorder the variables
nhhweight2 <- nhhweight2[,c(1,4,3,2)]
head(nhhweight2)
```

At this point we have a file with new countywide expanded and unexpanded
household weights. Let us save it in a selected directory:

```{r hw-22,cache=TRUE}
# Save 
setwd(pd("Weight/NewWeightData"))
write.csv(nhhweight2, file=pd("Weight/NewWeightData/nhhweight922015.csv"),
          row.names = FALSE) 
```

Let us now merge the new weights with household data and save the resulting dataset
as a CSV file:

```{r hw-23,cache=TRUE}
# Merge the new weight file with original household data file
nhhdataw<-merge(hhdata,nhhweight2, by =c("SAMPN"), all = TRUE)
head(nhhdataw)
write.csv(nhhdataw, file=pd("Weight/NewWeightData/nhhdataw922015.csv"),row.names=FALSE)
```

We have a household data file with new countywide household
weights.  As mentioned earlier we should check whether the new weights
represent the total number of households of each county, so we perform some
random checks for different variables. Let us take Alameda county as an example:

```{r hw-24,cache=TRUE}
# Random Check
setwd(pd("Weight/Check"))
# We do it for Alameda county (FIP code=1)
hhdata_01<-subset(nhhdataw,HCTFIP==1)
head(hhdata_01$HCTFIP)

# For household size
hsize<-as.data.frame(table(hhdata_01$HHSIZ))
hsize$Percent<-hsize$Freq*100/sum(hsize$Freq)
hsize
whsize<-aggregate(NEXPHHWGT~HHSIZ,FUN=sum,data=hhdata_01)
whsize$Percent<-whsize$NEXPHHWGT*100/sum(whsize$NEXPHHWGT)
whsize<- rbind(whsize, c("Total", colSums(whsize[,2:3],na.rm = TRUE)))
whsize
```

Let us perform a random check for the household home ownership (`OWN`) variable:

```{r hw-25,cache=TRUE}
# For home ownership  variable
hown<-as.data.frame(table(hhdata_01$OWN))
hown$Percent<-hown$Freq*100/sum(hown$Freq)
hown
wown<-aggregate(NEXPHHWGT~OWN,FUN=sum,data=hhdata_01)
wown$Percent<-wown$NEXPHHWGT*100/sum(wown$NEXPHHWGT)
wown<- rbind(wown, c("Total", colSums(wown[,2:3],na.rm = TRUE)))
wown
```

The total number of households in Alameda county is 53,2026 (based on
census data). Our random check shows that the new weights align with the
total number of households in Alameda county for both variables. This concludes this section.

\fi
