---
title: "4.3 Other Geographies (City of Los Angeles)"
author: "Suman Mitra"
date: "September 2, 2015"
output: html_document
---
## Overview 

This document explains how person level weights are developed for City of Los Angeles (LA) using the R statistical software for people of City of Los Angeles who participated in the 2012 CHTS. The purpose of these weights is to scale up information provided by respondents to the population of City of Los Angeles based on the socio-economic variables listed below.  Users of these weights are cautioned that these weights should not be applied to lower-level geographies such as zip codes. 

## Methodology

Weights are calculated by raking at the person level. These weights adjust the relative importance of responses to reflect the different probabilities of selection of respondents, and align the sample distributions to population distributions, based on the 2010 Census. In particular, the  weights are adjusted so that the sums of the adjusted weights are equal to known population totals for certain subgroups of the population of the City of Los Angeles. Since sampling information is not available for the City of Los Angeles, we assume equal probability here. Variables used for raking at the person level are:

  * Hispanic Status (Hispanic, Non-Hispanic)-LA City-wide distribution.
  * Ethnicity (White, African American, Asian, Other)-LA City-wide distribution.
  * Age (less than 20 years, 20 - 24 years, 25 - 34 years, 35 - 54 years, 55 - 64 years, 65 years or older)-LA City-wide distribution.
  * Employment Status (Part-time or full-time Employed, Not-employed)-LA City-wide distribution
  * Gender (Male, Female) - LA City-wide distribution

Following the raking procedure, very large weights are capped to be no more than five times the mean weight. To check whether the new weights represent the total population of the City of Los Angeles, we do some random checking for different variables and the results show that the new weights align with the population of the City of Los Angeles. We used this new person weight to calculate the final trip weight which is an outcome of Trip Correction Factor (TCF) * New Person Weight of LA City. 


## R Code and Implementation

The following commands are used to generate person weight for City of Los Angeles.


### Generate Person level weight for City of LA

First we set the working directory. Then we need to import the person data file and corresponding lookup file (containing location information) and then we merge them.

```{r og-1,echo=FALSE}
if ( !exists("rd") ) { rd <- "/home/crindt/chts-book" }
p <- function(ps) { return(paste(rd,ps,sep="/")) }
```

```{r og-2,cache=TRUE}
# Set the working directory
setwd(p("Weight/LA_Weight"))
# Import the person level variables csv file and give a new name
library(chts2011)
library(chts2011pvt)
#persondata <- read.csv("c:/Users/user/Documents/Suman_research/Caltrans Project/Data/deliv_per.csv", header=TRUE)
persondata <- chts_per
# Import the lookup variable csv file and give a new name (containing x, y coordinate data)
#lookup_hh <- read.csv("c:/Users/user/Documents/Suman_research/Caltrans Project/Data/lookup_hh.csv", header=TRUE)
lookup_hh <- chts_lu_hh
# Merging two data set to get the county information
perdata2<-merge(persondata,lookup_hh, by ="SAMPN", all = TRUE)
# Names of the new merged file
names(perdata2)
```

We have new file which contains information of respondents from the City of Los Angeles. We use this data file to generate person weight for City of Los Agneles. Now, we need to recode the raking variables to make them similar to the census data (Here we are using American Community Survey, 2010 five year estimate data). 

First we will assign zero for 'NA' observations of raking variables because the code we will use later for raking can't read 'NA'. 

```{r og-3,cache=TRUE}
# Assign 0 for all NA observations
perdata2$HISP[is.na(perdata2$HISP)] <- 0
perdata2$AGE[is.na(perdata2$AGE)] <- 0
perdata2$RACE1[is.na(perdata2$RACE1)] <- 0
perdata2$EMPLY[is.na(perdata2$EMPLY)] <- 0
```

Then, we change the value of hispanic variable to make it similar to census data. 

```{r og-4,cache=TRUE}
# For changing values of Hispanic
perdata2$HISPn<-ifelse(perdata2$HISP>2,2,perdata2$HISP)

# Show values
head(perdata2$HISPn,10)
```

Next, we recode the age and race variable.

```{r og-5,cache=TRUE}
# For recoding AGE
perdata2$AGEn<-ifelse(perdata2$AGE<20,1,ifelse((perdata2$AGE>=20 & perdata2$AGE<=24),2,ifelse((perdata2$AGE>=25 & perdata2$AGE<=34),3,ifelse((perdata2$AGE>=35 & perdata2$AGE<=54),4,ifelse((perdata2$AGE>=55 & perdata2$AGE<=64),5,6)))))
# Show values
head(perdata2$AGE,10)

# For recoding race
perdata2$RACEn<-ifelse(perdata2$RACE1==1,1,ifelse(perdata2$RACE1==2,2,ifelse(perdata2$RACE1==4,3,4)))
# Show values
head(perdata2$RACEn,10)
```

Here, we recode employmnet (EMPLY) and Gender variable.

```{r og-6,cache=TRUE}
# For recoding EMPLY
perdata2$EMPLYn<-ifelse(perdata2$EMPLY==1,1,ifelse(perdata2$EMPLY==2,2,ifelse(perdata2$EMPLY==8,1,2)))
# Show values
head(perdata2$EMPLYn,10)

# For recoding Gender
perdata2$GENDn<-ifelse(perdata2$GEND==1,1,2)
# Show the data
head(perdata2$GENDn,10)
```

Now we have data file (perdata2) which has similar coding  as census data for raking variables. Next, we need to subset the data for respondents who live in the City of Los Angeles. The variable named as HCITY contain the city name of respondents' residential  location.

```{r og-7,cache=TRUE}
# Create a subset for city of LA
sum(perdata2$HCITY=="LOS ANGELES")
perdata_LA<-subset(perdata2,HCITY=="LOS ANGELES")
dim(perdata_LA)
head(perdata_LA)
```

Now we have a new file named as perdata_LA which we use to generate person weight for the City of LA. We need census data of raking variables (Hispanic, Race, Employment status, Gender and Age). We collect this information from census and organize them as per our requriment and save as csv file. Then we import them and create seperate dataframe for City of LA (the csv file contain informaiton of all the counties).
 
First we import race variable for City of LA.

```{r og-8,cache=TRUE}
#Import the LA city population data of raking variables 
race.data_LA <- read.csv(p("Weight/LA_Weight/Raking variables/race_LA.csv"), header=TRUE)
race.data_LA
```

Next, we import hispanic variable for the City of LA.

```{r og-9,cache=TRUE}
hisp.data_LA<- read.csv(p("Weight/LA_Weight/Raking variables/hisp_LA.csv"), header=TRUE)
hisp.data_LA
```

Then, we import age variable.

```{r og-10,cache=TRUE}
age.data_LA <- read.csv(p("Weight/LA_Weight/Raking variables/age_LA.csv"), header=TRUE)
age.data_LA
```

Finally, we import geneder and employment status variable.

```{r og-11,cache=TRUE}
gend.data_LA <- read.csv(p("Weight/LA_Weight/Raking variables/gend_LA.csv"), header=TRUE)
gend.data_LA

emply.data_LA <- read.csv(p("Weight/LA_Weight/Raking variables/emply_LA.csv"), header=TRUE)
emply.data_LA
```

As our data is ready for raking, next we proceed for raking to generate person weight for the City of LA. To do this we need to install the library "survey". First we design a survey using 'svydesign' command and we use sampling without replacement (pps ="brewer") option. Since sampling information is not available for the City of Los Angeles, we assume equal probability here. So we don't need to specify the fpc option here. 


```{r og-12,cache=TRUE}
# Survey design with equal probability
# Need library "survey"
library (survey)
temp1<- svydesign(id=~1,data=perdata_LA,pps="brewer")
```

Then we perform the raking using 'rake' command and check the summary statistics of new weight.

```{r og-13,cache=TRUE}
# Raking for the City of LA to create new expanded weight
temp2<- rake(temp1, list(~RACEn,~HISPn,~AGEn,~EMPLYn,~GENDn), list(race.data_LA,hisp.data_LA,age.data_LA,emply.data_LA,gend.data_LA),control = list(maxit = 100))
# Summary statistics of new weight
summary(weights(temp2)) 
```

After raking, we trim the weight so that very large weights were capped to be no more than five times the mean of weight. The generated expanded person weights for City of LA are saved in our specified directory.

```{r og-14,cache=TRUE}
# Trim the weights (max value no greater than 5 times the mean)
temp3<-trimWeights(temp2,lower=(min(weights(temp2))),upper=(5*mean(weights(temp2))))
# Summary of trimmed weight
summary(weights(temp3))
# Combine expanded weight with sample no and person no
perweight_LA<-cbind(temp3$variables[1:2],weights(temp3))
names(perweight_LA)
# Change the column name of weight variable
colnames(perweight_LA)[3] <- "NEXPPERWGT_LA"
# show the new weights
head(perweight_LA)

```


Now we have on file which contains new expanded person weight for the City of LA. We want to generate new  person weight (not expanded) for the City of LA from our expanded weight. We know that person weight = Expanded weight * sample ratio and sample ratio = sample population /City of LA total population. 

First, we generate sample ratio for the City of LA.

```{r og-15,cache=TRUE}
# Create a variable for sampling ratio (sr)=sample population (4889) / City total population based on 2010 census(3792621)
sr<- (sum(perdata2$HCITY=="LOS ANGELES"))/3792621 
perdata_LA$SR <- sr
head(perdata_LA$sr)
```

Here, we generate the unexpanded new person weight for the City of LA.

```{r og-16,cache=TRUE}
# Unexpanded person weight= Expanded weight* sample ratio
perweight_LA$NPERWGT_LA<-perweight_LA$NEXPPERWGT_LA*perdata_LA$SR
head(perweight_LA)
```

We save the new weight as csv file by using following code.

```{r og-17,cache=TRUE}
# Save as csv file
write.csv(perweight_LA, file="PerWeight_LA.csv",row.names=FALSE) 
```

We want to merge the new weight with the City of LA person data file.

```{r og-18,cache=TRUE}
# Merge the new weight file with original person data file
nper_LA<-merge(perdata_LA,perweight_LA, by =c("SAMPN","PERNO"), all = TRUE)
head(nper_LA)
write.csv(nper_LA, file="npersondataLA.csv",row.names=FALSE)
```

Now, we have a person data file of the City of LA with new person weight.

As we mentioned earlier we want check whether the new weights represent the total population of the City of LA. So we perform some random checks for different variables using following codes. 

```{r og-19,cache=TRUE}
# Random Check
## For nativity
wntv<-aggregate(NEXPPERWGT_LA~NTVTY,FUN=sum,data=nper_LA)
wntv$Percent<-wntv$NEXPPERWGT_LA*100/sum(wntv$NEXPPERWGT_LA)
wntv
wntv1<- rbind(wntv, c("Total", colSums(wntv[,2:3])))
wntv1
```

Here, we do a random check for education variable.

```{r og-20,cache=TRUE}
# For education
weduc<-aggregate(NEXPPERWGT_LA~EDUCA,FUN=sum,data=nper_LA)
weduc$Percent<-weduc$NEXPPERWGT_LA*100/sum(weduc$NEXPPERWGT_LA)
weduc1<- rbind(weduc, c("Total", colSums(weduc[,2:3])))
weduc1
```

The total population of the City of LA is 3,792,621  (based on census). Our random check shows that the new weights align with the total population of the City of LA for both variables. 

Now we need to calculate the new final trip weight using the new person weight of the City of LA. To do this, first, we need to import the place level file and merge it with new weight file of the City of LA.

```{r og-21,cache=TRUE}
#Import the place level data  csv file and give a new name
nplacedata <- read.csv(p("Weight/NewWeightData/nplacedata932015.csv"), header=TRUE)

# Merge with LA weight
nplacedata_LA<-merge(nplacedata,perweight_LA, by =c("SAMPN","PERNO"), all = TRUE)
names(nplacedata_LA)
```

The final trip weight is an outcome of Trip Correction Factor (TCF) * New Person Weight of LA City. The follow codes are used to calculate the final trip weight and the expanded final trip weight for the City of LA.

```{r og-22,cache=TRUE}
# Generate Final trip weight (TCF*NPERWGT_LA) for LA
nplacedata_LA$NTCFPERWGT_LA<-nplacedata_LA$TCF*nplacedata_LA$NPERWGT_LA
# Generate Final expanded trip weight (TCF*NEXPPERWGT_LA) for LA
nplacedata_LA$NEXPTCFPERWGT_LA<-nplacedata_LA$TCF*nplacedata_LA$NEXPPERWGT_LA
# subset it for City of LA
placedata_LA<-subset(nplacedata_LA,PPrimaryCity=="LOS ANGELES")
# show 
head(placedata_LA)
```

If we want to save the new file with new trip weight as csv file, we can do it by using following codes.

```{r og-23,cache=TRUE}
write.csv(placedata_LA, file="nplacedata_LA.csv",row.names=FALSE)
```

Now we have new place data file with new trip weight for the City of LA.
