---
title: "4.3 Other Geographies (City of Los Angeles)"
author: "Suman Mitra"
date: "September 2, 2015"
output: html_document
---

```{r og-setup-0,echo=FALSE,cache=TRUE}
if ( !exists("rd") ) { rd <<- gsub("(.*?chts-book[^/]*).*$","\\1",getwd()) } 
source(paste(rd,"R/setup-hooks.R",sep="/"))
write(paste("Using Root Directory",rd,"\n"),stderr())
```


```{r results='asis',echo=FALSE,cache=TRUE}
############################################################ NESTABLE SECTION
SEC('#','Overview ')
```

This section explains how person-level weights are developed for the City
of Los Angeles (LA) using the R statistical software. The purpose of
these weights is to scale up information provided by 2012 CHTS respondents to
the population of City of Los Angeles based on the socio-economic
variables listed below.  Users of these weights are cautioned that
these weights should not be applied to lower-level geographies such as
zip codes.

```{r results='asis',echo=FALSE,cache=TRUE}
############################################################ NESTABLE SECTION
SEC('#','Methodology')
```

Weights are calculated by raking at the person level. These weights
adjust the relative importance of responses to reflect the different
probabilities of selection of respondents, and align the sample
distributions to population distributions, based on the 2010
Census. In particular, the weights are adjusted so that the sums of
adjusted weights are equal to known population totals for selected
subgroups of the population of the City of Los Angeles. Since detailed sampling
information is not available for the City of Los Angeles, we assume
equal probability of selection. Variables used for raking at the person level
are:

  * Hispanic Status (Hispanic, Non-Hispanic)-LA City-wide distribution.
  * Ethnicity (White, African American, Asian, Other)-LA City-wide distribution.
  * Age (less than 20 years, 20 to 24 years, 25 to 34 years, 35 to 54 years, 55 to 64 years, 65 years or older)-LA City-wide distribution.
  * Employment Status (Part-time or full-time Employed, Not-employed)-LA City-wide distribution
  * Gender (Male, Female) - LA City-wide distribution

Following the raking procedure, very large weights are capped to be no
more than five times the mean of weights. To check whether the new weights
represent the total population of the City of Los Angeles, we perform some
random checks for different variables and the results show that the
new weights align well with the population of the City of Los Angeles. We
use these new person weights to calculate the final trip weights from
`Trip Correction Factor (TCF)` * `New Person Weight of LA City`.


```{r results='asis',echo=FALSE,cache=TRUE}
############################################################ NESTABLE SECTION
SEC('#','R Code and Implementation')
```

To generate person weight for the City of Los Angeles, we first set the working directory. Then we import the person
data file and the corresponding lookup file (containing location information), and then we merge them.

```{r og-1,echo=FALSE}
if ( !exists("rd") ) { rd <- "/home/crindt/chts-book" }
p <- function(ps) { return(paste(rd,ps,sep="/")) }
```

```{r og-2,cache=TRUE}
# Set the working directory
setwd(p("Weight/LA_Weight"))

# Import the person level variables and give a new name
library(chts2011)
library(chts2011pvt)

persondata <- chts_per
lookup_hh <- chts_lu_hh

# Merging two data set to get the county information
perdata2<-merge(persondata,lookup_hh, by ="SAMPN", all = TRUE)

# Names of the new merged data frame
names(perdata2)
```

This new file contains information of 2012 CHTS respondents who live in the
City of Los Angeles. We use this file to generate person weights
for the City of Los Angeles. First, we need to recode the raking variables
so their definition matches with the corresponding Census variables (Here we use American
Community Survey, 2010 five year estimate data).

First we substitute zero to `NA` in raking variables
because the raking program cannot read `NA`.

```{r og-3,cache=TRUE}
# Assign 0 for all NA observations
perdata2$HISP[is.na(perdata2$HISP)] <- 0
perdata2$AGE[is.na(perdata2$AGE)] <- 0
perdata2$RACE1[is.na(perdata2$RACE1)] <- 0
perdata2$EMPLY[is.na(perdata2$EMPLY)] <- 0
```

Second, we change the value of the hispanic variable to make it similar to the corresponding census data variable.

```{r og-4,cache=TRUE}
# For changing values of Hispanic
perdata2$HISPn<-ifelse(perdata2$HISP>2,2,perdata2$HISP)

# Show values
head(perdata2$HISPn,10)
```

Next, we recode the age and race variables:

```{r og-5,cache=TRUE}
# For recoding AGE
perdata2$AGEn <-
  ifelse(perdata2$AGE<20,1,
         ifelse((perdata2$AGE>=20 & perdata2$AGE<=24),2,
                ifelse((perdata2$AGE>=25 & perdata2$AGE<=34),3,
                       ifelse((perdata2$AGE>=35 & perdata2$AGE<=54),4,
                              ifelse((perdata2$AGE>=55 &
                                        perdata2$AGE<=64),5,6)))))
# Show values
head(perdata2$AGE,10)

# For recoding race
perdata2$RACEn <-
  ifelse(perdata2$RACE1==1,1,
         ifelse(perdata2$RACE1==2,2,
                ifelse(perdata2$RACE1==4,3,4)))
# Show values
head(perdata2$RACEn,10)
```

Now let us recode the employment (`EMPLY`) and gender variables:

```{r og-6,cache=TRUE}
# For recoding EMPLY
perdata2$EMPLYn <-
  ifelse(perdata2$EMPLY==1,1,
         ifelse(perdata2$EMPLY==2,2,
                ifelse(perdata2$EMPLY==8,1,2)))

# Show values
head(perdata2$EMPLYn,10)

# For recoding Gender
perdata2$GENDn<-ifelse(perdata2$GEND==1,1,2)
# Show the data
head(perdata2$GENDn,10)
```

Raking variables in our data frame (`perdata2`) now are coded the same as the corresponding census
variables. Next, we need to subset the data for respondents who live in the City of Los Angeles. 
The variable named `HCITY` contain the city name of respondents' residential location.

```{r og-7,cache=TRUE}
# Create a subset for city of LA
sum(perdata2$HCITY=="LOS ANGELES")
perdata_LA<-subset(perdata2,HCITY=="LOS ANGELES")
dim(perdata_LA)
head(perdata_LA)
```

Now we have a new data frame named `perdata_LA` which we use to
generate person weights for the City of LA.  We need census data of
raking variables (Hispanic, Race, Employment status, Gender and
Age).  We collect this information from census and organize them as per
our requirement and save as CSV file. Then we import them and create
separate dataframe for City of LA (the CSV file contain information of
all the counties).
 
First, let us import the race data for residents of the city of LA:

```{r og-8,cache=TRUE}
#Import the LA city population data of raking variables 
race.data_LA <- read.csv(p("Weight/LA_Weight/Raking variables/race_LA.csv"),
                         header=TRUE)
race.data_LA
```

Next, we import the Hispanic variables for the City of LA.

```{r og-9,cache=TRUE}
hisp.data_LA<- read.csv(p("Weight/LA_Weight/Raking variables/hisp_LA.csv"),
                        header=TRUE)
hisp.data_LA
```

Now, let us import the age variables:

```{r og-10,cache=TRUE}
age.data_LA <- read.csv(p("Weight/LA_Weight/Raking variables/age_LA.csv"),
                        header=TRUE)
age.data_LA
```

Finally, let us import the gender and employment status variables:

```{r og-11,cache=TRUE}
gend.data_LA <- read.csv(p("Weight/LA_Weight/Raking variables/gend_LA.csv"),
                         header=TRUE)
gend.data_LA

emply.data_LA <- read.csv(p("Weight/LA_Weight/Raking variables/emply_LA.csv"),
                          header=TRUE)
emply.data_LA
```

Our data are ready for raking. We now need to install the library "survey". 
We design a survey using the `svydesign` command and we use sampling without replacement (pps
="brewer") option. Since sampling information is not available for the
City of Los Angeles, we assume equal probability for each LA resident, so we do not
need to specify the fpc option here.


```{r og-12,cache=TRUE}
# Survey design with equal probability
# Need library "survey"
library (survey)
temp1<- svydesign(id=~1,data=perdata_LA,pps="brewer")
```

Then we perform the raking using the `rake` command and check the summary
statistics of the new weighs:

```{r og-13,cache=TRUE}
# Raking for the City of LA to create new expanded weight
temp2<- rake(temp1,
             list(~RACEn,~HISPn,~AGEn,~EMPLYn,~GENDn),
             list(race.data_LA,hisp.data_LA,age.data_LA,
                  emply.data_LA,gend.data_LA),
             control = list(maxit = 100))

# Summary statistics of new weight
summary(weights(temp2)) 
```

After raking, we trim the weights so that large weights are
capped to no more than five times the mean of weights. The generated
expanded person weights for the City of LA are saved in a file located in a directory specified
above.

```{r og-14,cache=TRUE}
# Trim the weights (max value no greater than 5 times the mean)
temp3<-trimWeights(temp2,lower=(min(weights(temp2))),upper=(5*mean(weights(temp2))))

# Summary of trimmed weight
summary(weights(temp3))

# Combine expanded weight with sample no and person no
perweight_LA<-cbind(temp3$variables[1:2],weights(temp3))
names(perweight_LA)

# Change the column name of weight variable
colnames(perweight_LA)[3] <- "NEXPPERWGT_LA"

# show the new weights
head(perweight_LA)

```


We now have a file that contains new expanded person weights for the
City of LA. To generate new person weights (not expanded) for
the City of LA from our expanded weights, we use that `person weight` =
`Expanded weight` * `sample ratio` and `sample ratio` =
`sample population`/`City of LA total population`.

First, let us generate a sample ratio for the City of LA:

```{r og-15,cache=TRUE}
# Create a variable for sampling ratio (sr)=
# sample population (4889) / City total population
# based on 2010 census(3792621)
sr<- (sum(perdata2$HCITY=="LOS ANGELES"))/3792621 
perdata_LA$SR <- sr
head(perdata_LA$sr)
```

Now, let us generate the unexpanded new person weights for the City of LA:

```{r og-16,cache=TRUE}
# Unexpanded person weight= Expanded weight* sample ratio
perweight_LA$NPERWGT_LA<-perweight_LA$NEXPPERWGT_LA*perdata_LA$SR
head(perweight_LA)
```

Let us save the new weights in a CSV file:

```{r og-17,cache=TRUE}
# Save as csv file
write.csv(perweight_LA, file="PerWeight_LA.csv",row.names=FALSE) 
```

At this point, We want to merge the new weights with the City of LA person data frame
and save the resulting dataset:

```{r og-18,cache=TRUE}
# Merge the new weight file with original person data file
nper_LA<-merge(perdata_LA,perweight_LA, by =c("SAMPN","PERNO"), all = TRUE)
head(nper_LA)
write.csv(nper_LA, file="npersondataLA.csv",row.names=FALSE)
```

This gives us a person data file of the City of LA with new person
weights.

As mentioned earlier, we need to check whether the new weights
represent the total population of the City of LA, so we perform some
random checks for different variables:

```{r og-19,cache=TRUE}
# Random Check
## For nativity
wntv<-aggregate(NEXPPERWGT_LA~NTVTY,FUN=sum,data=nper_LA)
wntv$Percent<-wntv$NEXPPERWGT_LA*100/sum(wntv$NEXPPERWGT_LA)
wntv
wntv1<- rbind(wntv, c("Total", colSums(wntv[,2:3])))
wntv1
```

Here, we perform a random check for the education variable:

```{r og-20,cache=TRUE}
# For education
weduc<-aggregate(NEXPPERWGT_LA~EDUCA,FUN=sum,data=nper_LA)
weduc$Percent<-weduc$NEXPPERWGT_LA*100/sum(weduc$NEXPPERWGT_LA)
weduc1<- rbind(weduc, c("Total", colSums(weduc[,2:3])))
weduc1
```

The total population of the City of LA is 3,792,621 people (based on
census). Our random check shows that the new weights align well with the
total population of the City of LA.

Now we need to calculate the new final trip weights using the new
person weights for the City of LA. To do this, we first import
the place level file and merge it with the new weights file.

```{r og-21,cache=TRUE}
#Import the place level data  csv file and give a new name
nplacedata <- read.csv(p("Weight/NewWeightData/nplacedata932015.csv"), header=TRUE)

# Merge with LA weight
nplacedata_LA<-merge(nplacedata,perweight_LA, by =c("SAMPN","PERNO"), all = TRUE)
names(nplacedata_LA)
```

The final trip weights are calculated from `Trip Correction Factor
(TCF)` * `New Person Weight of LA City`. The code below calculates
the final trip weights and the expanded final trip weights for the City
of LA:

```{r og-22,cache=TRUE}
# Generate Final trip weight (TCF*NPERWGT_LA) for LA
nplacedata_LA$NTCFPERWGT_LA<-nplacedata_LA$TCF*nplacedata_LA$NPERWGT_LA
# Generate Final expanded trip weight (TCF*NEXPPERWGT_LA) for LA
nplacedata_LA$NEXPTCFPERWGT_LA<-nplacedata_LA$TCF*nplacedata_LA$NEXPPERWGT_LA
# subset it for City of LA
placedata_LA<-subset(nplacedata_LA,PPrimaryCity=="LOS ANGELES")
# show 
head(placedata_LA)
```

Finally, we save the new data frame with the trip weights as a CSV file:

```{r og-23,cache=TRUE}
write.csv(placedata_LA, file="nplacedata_LA.csv",row.names=FALSE)
```

We have a new place datafile with new trip weights for the City of
LA. This concludes this sectiion.
