---
title: "Data Request for the City of Los Angeles"
author: "Suman Mitra"
date: 'This version: `r (function() {f<-dir(); format(max(file.mtime(f[grepl("^.*\\.Rmd$",f)])), format="%B %d, %Y")})()`'
output:
  pdf_document:
    keep_tex: yes
    template: custom.latex
fontfamily: inconsolata
fontsize: 11pt
geometry: margin=1in
params:
  documentclass: article
  documentation: '`r if(exists("documentation")){documentation}else{documentation<-"false";documentation}`'
---

```{r la-city-setup-0,echo=FALSE,cache=FALSE}
if ( !exists("rd") ) { rd <<- gsub("(.*?chts-book[^/]*).*$","\\1",getwd()) } 
source(paste(rd,"R/setup-hooks.R",sep="/"))
write(paste("Using Root Directory",rd,"\n"),stderr())
```

# Question

This question asked for the following information:

1. Total number of linked trips, by mode, whose origin was in the City
   of Los Angeles. This will include trips whose origin and
   destination were in the City of Los Angeles. If possible, we would
   be interested in also seeing the total duration and total distance
   of these associated trips.

2. Total number of linked trips, by mode, whose destination was in the
   City of Los Angeles, but whose origin was outside the City of Los
   Angeles. If possible, we would be interested in also seeing the
   total duration and total distance of these associated trips.

3. Total number of linked trips, by mode, whose origin and destination
   was outside the City of Los Angeles that passed through the City of
   Los Angeles. If possible, we would be interested in also seeing the
   total duration and total distance of these associated trips.

# Summary and Methodology


We have generated weighted results for the first two requests. First
we generated specific person level weights for the City of Los
Angeles. Weights were calculated by raking at the person level. These
weights adjust the relative importance of responses to reflect the
different probabilities of selection of respondents, and align the
sample distributions to population distributions, based on the 2010
Census. In particular, the aforementioned weights are adjusted so that
the sums of the adjusted weights are equal to known population totals
for certain subgroups of the population of the City of Los
Angeles. Since sampling information was not available for the City of
Los Angeles, we assumed equal probability here. Variables used for
raking at the person level are:

* Hispanic Status (Hispanic, Non-Hispanic) – LA City-wide distribution.
* Ethnicity (White, African American, Asian, Other) – LA City-wide distribution.
* Age (less than 20 years, 20 – 24 years, 25 – 34 years, 35 – 54
  years, 55 – 64 years, 65 years or older) – LA City -wide
  distribution.
* Employment Status (Part-time or full-time Employed, Not-employed) -
  LA City-wide distribution
* Gender (Male, Female) - LA City-wide distribution

Following the raking procedure, very large weights were capped to be
no more than five times the mean weight. To check whether the new
weights represent the total population of the City of Los Angeles, we
did some random checking for different variables and the results
showed that the new weights align with the population of the City of
Los Angeles.

To calculate the total number of linked trips we generated linked trip
table following the CHTS 2001 methodology (detail methodology is in
separate document). Since linked trip table does not contain the
origin and destination city name, we merged it with the place level
file which contains those information. We then merged this new data
set with new weights of the City of LA based on sample number and
person number. We used this merged data set to generate linked trips
whose origin was in the City of Los Angeles, destination was in the
City of Los Angeles, origin and destination both were in the City of
Los Angeles and finally destination was in the City of Los 2 Angeles
but origin was outside the City of Los Angeles. To do this, we
generated different sub sets of data based on different
origin-destination conditions. The variables used for sub-setting the
data are `PPrimaryCity` (origin) and `City` (destination) which
renamed later as `oCity` and `dCity` respectively. We used the new
weights of the City of Los Angeles to generate weighted results for
trips originated in the City of Los Angeles. We then used statewide
weights to generate weighted results for trips originated outside of
the City of Los Angeles. We generated tables by mode for the
following: total number of linked trips, total linked trip distance
and total linked trip duration assuming that all requests were by
mode. We used R programming (RStudio). The codes used for generating
weights and the requested tables are given below:

\ifmydocumentation

# R Code and Implementation

## Preparing the data for answering the questions

The first step in the process to answer the questions is to generate
person-weights for the City of Los Angeles.  We use methods similar to
those described in the chapter on computing weights.

```{r la-city-weights-1}

# We need the survey library to perform raking
library(survey)

# Import the person level variables and give a new name
perdata2 <- chts_per %>% left_join(chts_hh,by=c('SAMPN'))
names(perdata2)
sum(perdata2$HCITY=="LOS ANGELES")
sum(perdata2$HPrimaryCity=="LOS ANGELES")


# Assign 0 for all NA observations
# FIXME: This is not good
perdata2$HISP[is.na(perdata2$HISP)] <- 0
perdata2$AGE[is.na(perdata2$AGE)] <- 0
perdata2$RACE1[is.na(perdata2$RACE1)] <- 0
perdata2$EMPLY[is.na(perdata2$EMPLY)] <- 0
```

We need to recode the raking variables to make them similar to the
census data (we are using American Community Survey, 2010 five year
estimate data).

```{r la-city-weights-2}

# For changing values of Hispanic
perdata2$HISPn <-
  ifelse(perdata2$HISP > 2, 2,
         perdata2$HISP)

# Show values
tail(perdata2$HISPn)

perdata2$AGEn <-
  ifelse(perdata2$AGE < 20,
         1,
         ifelse(perdata2$AGE <= 24,
                2,
                ifelse(perdata2$AGE <= 34,
                       3,
                       ifelse(perdata2$AGE <= 54,
                              4,
                              ifelse(perdata2$AGE <= 64,
                                     5,
                                     6)))))


# For recoding race

perdata2$RACEn <-
  ifelse(perdata2$RACE1 == 1,
         1,
         ifelse(perdata2$RACE1 == 2,
                2,
                ifelse(perdata2$RACE1 == 4,
                       3,
                       4)))

# For recoding employment
perdata2$EMPLYn <-
  ifelse(perdata2$EMPLY == 1,
         1,
         ifelse(perdata2$EMPLY == 2,
                2,
                ifelse(perdata2$EMPLY == 8,
                       1,
                       2)))

# For recoding Gender
perdata2$GENDn <-
  ifelse(perdata2$GEND == 1,
         1,
         2)


# Generate a subset for city of LA
perdata_LA<-subset(perdata2,HCITY=="LOS ANGELES")
dim(perdata_LA)
names(perdata_LA)
head(perdata_LA)


# Import the LA city population data of raking variables
# These data are based on 2010 Census
race.data_LA <- read.csv(pd("Weight/LA_Weight/Raking variables/race_LA.csv"),
                         header=TRUE)
race.data_LA

hisp.data_LA<- read.csv(pd("Weight/LA_Weight/Raking variables/hisp_LA.csv"),
                        header=TRUE)
hisp.data_LA

age.data_LA <- read.csv(pd("Weight/LA_Weight/Raking variables/age_LA.csv"),
                        header=TRUE)
age.data_LA

gend.data_LA <- read.csv(pd("Weight/LA_Weight/Raking variables/gend_LA.csv"),
                         header=TRUE)
gend.data_LA

emply.data_LA <- read.csv(pd("Weight/LA_Weight/Raking variables/emply_LA.csv"),
                          header=TRUE)
emply.data_LA
```

With all the data prepped, we now generate the expansion weights via
raking.

```{r la-city-weights-3}

# Survey design with equal probability
temp1<- svydesign(id=~1,data=perdata_LA,pps="brewer")
temp2<- rake(temp1,
             list(~RACEn,~HISPn,~AGEn,~EMPLYn,~GENDn),
             list(race.data_LA,hisp.data_LA,age.data_LA,emply.data_LA,gend.data_LA),
             control = list(maxit = 100)) 

# Raking
summary(weights(temp2)) # Summary statistics of new weight

# Trim the weights (max value no greater than 5 times the mean)
temp3<-trimWeights(temp2,
                   lower=(min(weights(temp2))),
                   upper=(5*mean(weights(temp2))))
summary(weights(temp3)) # Summary of trimmed weight

# Combine expanded weight with sample no and person no
perweight_LA<-cbind(temp3$variables[1:2],weights(temp3))
head(perweight_LA)

# change the column name of weight variable
colnames(perweight_LA)[3] <- "NEXPPERWGT_LA" 
head(perweight_LA)

# Generate person weight
# Generate a variable for sampling weight
# (fpc) = sample population (4889) / City total population
# based on 2010 census(3792621)

fpc<- (sum(perdata2$HCITY=="LOS ANGELES"))/3792621
perdata_LA$fpc <- fpc
head(perdata_LA$fpc)

# Person weight= Expanded weight* fpc (sampling weight)
perweight_LA$NPERWGT_LA<-perweight_LA$NEXPPERWGT_LA*perdata_LA$fpc
head(perweight_LA)

# Save as csv file
write.csv(perweight_LA, file=pto("PerWeight_LA.csv"),row.names=FALSE) 
```

Now we have new weight for LA City which we will use to generate the
requested tables Random Check whether new weights represent total
population of the City of LA.  We run some checks to to make sure we
did things correctly.

```{r la-city-weights-4}

# Merge the new weight file with original person data file
nper_LA<-merge(perdata_LA,perweight_LA, by =c("SAMPN","PERNO"), all = TRUE)
head(nper_LA)

# For race
wrace<-aggregate(NEXPPERWGT_LA~RACE1,FUN=sum,data=nper_LA)
wrace$Percent<-wrace$NEXPPERWGT_LA*100/sum(wrace$NEXPPERWGT_LA)
wrace
wrace1<- rbind(wrace, c("Total", colSums(wrace[,2:3])))
wrace1

# For nativity
wntv<-aggregate(NEXPPERWGT_LA~NTVTY,FUN=sum,data=nper_LA)
wntv$Percent<-wntv$NEXPPERWGT_LA*100/sum(wntv$NEXPPERWGT_LA)
wntv
wntv1<- rbind(wntv, c("Total", colSums(wntv[,2:3])))
wntv1

# For education
weduc<-aggregate(NEXPPERWGT_LA~EDUCA,FUN=sum,data=nper_LA)
weduc$Percent<-weduc$NEXPPERWGT_LA*100/sum(weduc$NEXPPERWGT_LA)
weduc1<- rbind(weduc, c("Total", colSums(weduc[,2:3])))
weduc1
```

To answer the questions we need the place level data, the linked trip
table (see the chapter on generating the linked trip table for
details), and new weights for LA City.  Since the linked trip table
does not contain the origin and destination city name, we have to
import the place level data which contains those information and merge
it with linked trip table.


```{r la-city-2}
# Import the place level data and give a new name
nplacedata <- chts_pla

# create mode text for table
nplacedata$Mode <-
  ifelse(nplacedata$MODE==1,"Walk",
  ifelse(nplacedata$MODE==2,"Bike",
  ifelse(nplacedata$MODE==3,"Wheelchair/Mobility Scooter",
  ifelse(nplacedata$MODE==4,"Other Non-Motorized",
  ifelse(nplacedata$MODE==5,"Auto/Van/Truck Driver",
  ifelse(nplacedata$MODE==6,"Auto /Van/Truck Passenger",
  ifelse(nplacedata$MODE==7,"Carpool/Vanpool",
  ifelse(nplacedata$MODE==8,"Motorcycle/Scooter/Moped",
  ifelse(nplacedata$MODE==9,"Taxi/Hired Car/Limo",
  ifelse(nplacedata$MODE==10,"Rental Car/Vehicle",
  ifelse(nplacedata$MODE==11,"Private shuttle ",
  ifelse(nplacedata$MODE==12,"Greyhound Bus",
  ifelse(nplacedata$MODE==13,"Plane",
  ifelse(nplacedata$MODE==14,"Other Private Transit",
  ifelse(nplacedata$MODE==15,"Local Bus,Rapid Bus",
  ifelse(nplacedata$MODE==16,"Express Bus/Commuter Bus",
  ifelse(nplacedata$MODE==17,"Premium Bus",
  ifelse(nplacedata$MODE==18,"School Bus",
  ifelse(nplacedata$MODE==19,"Public Transit Shuttle",
  ifelse(nplacedata$MODE==20,"Air BART/LAX Fly Away",
  ifelse(nplacedata$MODE==21,"Dial-a-Ride/Paratransit", 
  ifelse(nplacedata$MODE==22,"Amtrak Bus", 
  ifelse(nplacedata$MODE==23,"Other Bus ", 
  ifelse(nplacedata$MODE==24,"BART,Metro Red/Purple Line",
  ifelse(nplacedata$MODE==25,"ACE, Amtrak, Cal train",
  ifelse(nplacedata$MODE==26,"Metro",
  ifelse(nplacedata$MODE==27,"Street Car/Cable Car",
  ifelse(nplacedata$MODE==28,"Other Rail",
  ifelse(nplacedata$MODE==29,"Ferry / Boat",
         "Don't Know/Refused"
         )))))))))))))))))))))))))))))


# Subset the dataset for origin city names
npl_or<-subset(nplacedata ,select= c('SAMPN','PERNO','PLANO','PPrimaryCity'))

# rename one variable
names(npl_or) [c(3,4)]<-c('oPLANO','oCity')
names(npl_or)
head(npl_or)

# Subset the dataset for destination city names
npl_dt<-subset(nplacedata ,select= c('SAMPN','PERNO','PLANO','CITY'))
names(npl_dt) [c(3,4)]<-c('dPLANO','dCity')
names(npl_dt)
head(npl_dt)

# Import the linked trip table
linked_trip <- chts_lt %>%
  mutate(oPLANO=as.integer(oplano)) %>%
  mutate(dPLANO=as.integer(dplano))

head(linked_trip)

# Merge the linked trip with origin and destination city name file by
# origin and destination place no
nplace_l<-merge(linked_trip,npl_or, by =c("SAMPN","PERNO","oPLANO"), all = FALSE)
nplace_t<-merge(nplace_l,npl_dt, by =c("SAMPN","PERNO","dPLANO"), all = FALSE)
names(nplace_t)
head(nplace_t)

# Import the LA City weight file (no, just use from above)
# perweight_LA <- read.csv(pd("Weight/LA_Weight/PerWeight_LA.csv"), header=TRUE)

# Import the person level variables
persondata <- chts_per
names(persondata)

# Subset the person file for state-wide weight
per_w<-subset(persondata ,select= c('SAMPN','PERNO','PERWGT','EXPPERWGT'))
names(per_w)

# Merge with LA weight
perw2<-merge(per_w,perweight_LA, by =c("SAMPN","PERNO"), all = TRUE)
head(perw2)

# Merge the linked trip file with LA weight
linkplace<-merge(nplace_t,perw2, by =c("SAMPN","PERNO"), all = FALSE)
names(linkplace)
head (linkplace,10)

```

Now we have a data frame (linkplace) with link trip table, origin and
destination place city name and with new LA City weight. We will use
this data frame to generate the requested tables.


## Linked Trips Originating in LA

We begin with linked trips originating in Los Angeles.  We start by
producing a table with the number and percent of linked trips by mode.

```{r la.city.lt.orig.la}

# Generate a subset for link trips originating in LA
place_LAO<-subset(linkplace,(oCity=="LOS ANGELES"|oCity=="LOS ÁNGELES"))
names(place_LAO)

## Total number of linked trips for all trips originating in LA, by mode
## Generate frequency for trip mode
temp1<-as.data.frame(table(place_LAO$ltmode))

## Generate percentage of frequency
temp1$Percent<-temp1$Freq*100/sum(temp1$Freq)

## Names
names(temp1)

## Rename the variables
colnames(temp1)[c(1:3)] <- c("Mode","Number of Linked-Trips(UW)","Percent(UW)")

## Show the results
temp1

## Generate weighted frequency (Use weight generated for LA city)
temp2<-aggregate(NEXPPERWGT_LA~ltmode,FUN=sum,data=place_LAO) 

## Generate weighted percentage
temp2$Percent<-temp2$NEXPPERWGT_LA*100/sum(temp2$NEXPPERWGT_LA)

## Names
names(temp2)

## Rename the variables
colnames(temp2)[c(1:3)] <- c("Mode","Number of Linked-Trips(W)","Percent(W)")

## Show
temp2

## Merge the two files and save
temp3<-merge(temp1,temp2, by ="Mode", all = TRUE,sort=FALSE)
temp3 <- temp3[,c(1,2,4,3,5)]
temp3

## Save as csv file
write.csv(temp3,file=pto("WlinkMode_LAO.csv"),row.names=FALSE)
la.city.lt.orig.la <- temp3
```

Next, we compute a table with the total linked trip duration for all
trips originating in LA, by mode.  We also compute the fraction of
total time spent in each mode by the population.


```{r la.city.lt.dur.orig.la}

temp1<-aggregate(tripdur~ltmode,FUN=sum,data=place_LAO)

## Generate percentage
temp1$Percent<-temp1$tripdur*100/sum(temp1$tripdur)

## Names
names(temp1)

## Rename the variables
colnames(temp1)[c(1:3)] <- c("Mode","Linked-Trip Duration (Min)(UW)","Percent(UW)")

## Show
temp1

## Generate weighted frequency
temp<-place_LAO$NEXPPERWGT_LA*place_LAO$tripdur
temp2<-aggregate(temp~ltmode,FUN=sum,data=place_LAO)

## Generate weighted percentage
temp2$Percent<-temp2$temp*100/sum(temp2$temp)

## Names
names(temp2)

## Rename the variables
colnames(temp2)[c(1:3)] <- c("Mode","Linked-Trip Duration(Min)(W)","Percent(W)")

## Show
temp2

## Merge the two files and save
temp3<-merge(temp1,temp2, by ="Mode", all = TRUE)
temp3 <- temp3[,c(1,2,4,3,5)]
temp3

## Save as csv file
write.csv(temp3,file=pto("WlinkTRIPDUR_LAO.csv"),row.names=FALSE)
la.city.lt.dur.orig.la <- temp3
```


Similarly, we compute a table with the total linked trip distance for
all trips originating in LA, by mode.  Again, we also compute the
fraction of total miles traveled in each mode by the population.


```{r la.city.lt.dist.orig.la}

## Total linked trip distance for all trips originating in LA, by mode
## Generate frequency for linked trip distance by trip mode
temp1<-aggregate(tripdist~ltmode,FUN=sum,data=place_LAO)

## Generate percentage
temp1$Percent<-temp1$tripdist*100/sum(temp1$tripdist)

## Names
names(temp1)

## Rename the variables
colnames(temp1)[c(1:3)] <- c("Mode","Linked-Trip Distance (Miles)(UW)","Percent(UW)")

## Show
temp1

## Generate weighted frequency
temp<-place_LAO$NEXPPERWGT_LA*place_LAO$tripdist
temp2<-aggregate(temp~ltmode,FUN=sum,data=place_LAO)

## Generate weighted percentage
temp2$Percent<-temp2$temp*100/sum(temp2$temp)

## Names
names(temp2)

## Rename the variables
colnames(temp2)[c(1:3)] <- c("Mode","Linked-Trip Distance (miles)(W)","Percent(W)")

## Show
temp2

## Merge the two files and save
temp3<-merge(temp1,temp2, by ="Mode", all = TRUE)
temp3 <- temp3[,c(1,2,4,3,5)]
temp3

## Save as csv file
write.csv(temp3,file=pto("WlinkTRIPDIST_LAO.csv"),row.names=FALSE)
la.city.lt.dist.orig.la <- temp3
```

## Linked Trips Originating in LA and ending in LA

Next, we turn to linked trips originating in LA and ending in LA.  As
before, we start with a count and frequency of linked trips by mode.

```{r la.city.lt.od.la}

# Generate a subset for trips originating in LA and ending in LA
place_LAOD<-subset(
    linkplace,
    ((oCity=="LOS ANGELES"|oCity=="LOS ÁNGELES")
        & (dCity=="LOS ANGELES"|dCity=="LOS ÁNGELES")))
names(place_LAOD)
head(place_LAOD)

## Total number of linked trips for all trips originating and ending in LA, by mode
## Generate frequency for trip mode
temp1<-as.data.frame(table(place_LAOD$ltmode))

## Generate percentage of frequency
temp1$Percent<-temp1$Freq*100/sum(temp1$Freq)

## Names
names(temp1)

## Rename the variables
colnames(temp1)[c(1:3)] <- c("Mode","Number of Linked-Trips(UW)","Percent(UW)")

## Show the results
temp1

## Generate weighted frequency
temp2<-aggregate(NEXPPERWGT_LA~ltmode,FUN=sum,data=place_LAOD)

## Generate weighted percentage
temp2$Percent<-temp2$NEXPPERWGT_LA*100/sum(temp2$NEXPPERWGT_LA)

## Names
names(temp2)

## Rename the variables
colnames(temp2)[c(1:3)] <- c("Mode","Number of Linked-Trips(W)","Percent(W)")

## Show
temp2

## Merge the two files and save
temp3<-merge(temp1,temp2, by ="Mode", all = TRUE,sort=FALSE)
temp3 <- temp3[,c(1,2,4,3,5)]
temp3

## Save as csv file
write.csv(temp3,file=pto("WlinkMode_LAOD.csv"),row.names=FALSE)
la.city.lt.od.la <- temp3
```

As before, we next compute a table with the total and fractional
linked trip duration for all trips originating in LA, by mode.

```{r la.city.lt.dur.od.la}

## Generate frequency for linked trip duration by trip mode
temp1<-aggregate(tripdur~ltmode,FUN=sum,data=place_LAOD)

## Generate percentage
temp1$Percent<-temp1$tripdur*100/sum(temp1$tripdur)

## Names
names(temp1)

## Rename the variables
colnames(temp1)[c(1:3)] <- c("Mode","Linked-Trip Duration (Min)(UW)","Percent(UW)")

## Show
temp1

## Generate weighted frequency
temp<-place_LAOD$NEXPPERWGT_LA*place_LAOD$tripdur
temp2<-aggregate(temp~ltmode,FUN=sum,data=place_LAOD)

## Generate weighted percentage
temp2$Percent<-temp2$temp*100/sum(temp2$temp)

## Names
names(temp2)

## Rename the variables
colnames(temp2)[c(1:3)] <- c("Mode","Linked-Trip Duration (Min)(W)","Percent(W)")

## Show
temp2

## Merge the two files and save
temp3<-merge(temp1,temp2, by ="Mode", all = TRUE,sort=FALSE)
temp3 <- temp3[,c(1,2,4,3,5)]
temp3

## Save as csv file
write.csv(temp3,file=pto("WlinkTRIPDUR_LAOD.csv"),row.names=FALSE)
la.city.lt.dur.od.la <- temp3
```

Continuing, we again compute a table with the total and fractional
linked trip distance for all trips originating in LA, by mode.

```{r la.city.lt.dist.od.la}

## Total linked trip distance for all trips originating and ending in LA, by mode
## Generate frequency for linked trip distance by trip mode
temp1<-aggregate(tripdist~ltmode,FUN=sum,data=place_LAOD)

## Generate percentage
temp1$Percent<-temp1$tripdist*100/sum(temp1$tripdist)

## Names
names(temp1)

## Rename the variables
colnames(temp1)[c(1:3)] <- c("Mode","Linked-Trip Distance (Miles)(UW)","Percent(UW)")

## Show
temp1

## Generate weighted frequency
temp<-place_LAOD$NEXPPERWGT_LA*place_LAOD$tripdist
temp2<-aggregate(temp~ltmode,FUN=sum,data=place_LAOD)

## Generate weighted percentage
temp2$Percent<-temp2$temp*100/sum(temp2$temp)

## Names
names(temp2)

## Rename the variables
colnames(temp2)[c(1:3)] <- c("Mode","Linked-Trip Distance (Miles)(W)","Percent(W)")

## Show
temp2

## Merge the two files and save
temp3<-merge(temp1,temp2, by ="Mode", all = TRUE,sort=FALSE)
temp3 <- temp3[,c(1,2,4,3,5)]
temp3

## Save as csv file
write.csv(temp3,file=pto("WlinkTRIPDIST_LAOD.csv"),row.names=FALSE)
la.city.lt.dist.od.la <- temp3
```

## Linked Trips ending in LA

Now we turn to linked trips ending in LA.  Following the same pattern,
we start with a count and frequency of linked trips by mode.

```{r la.city.lt.dest.la}

#########################
# Generate a subset for trips ending in LA
place_LAD<-subset(linkplace,(dCity=="LOS ANGELES"|dCity=="LOS ÁNGELES"))
names(place_LAD)
head(place_LAD)

# Change the weight for non LA origin (Assign state wide weight for non LA origin)
place_LAD$NW <- ifelse((
    place_LAD$oCity=="LOS ANGELES"|place_LAD$oCity=="LOS ÁNGELES"),
    place_LAD$NEXPPERWGT_LA, place_LAD$EXPPERWGT)
names(place_LAD)
head(place_LAD)

## Total number of linked trips for all trips ending in LA, by mode
## Generate frequency for trip mode

temp1<-as.data.frame(table(place_LAD$ltmode))

## Generate percentage of frequency
temp1$Percent<-temp1$Freq*100/sum(temp1$Freq)

## Names
names(temp1)

## Rename the variables
colnames(temp1)[c(1:3)] <- c("Mode","Number of Linked-Trips(UW)","Percent(UW)")

## Show the results
temp1

## Generate weighted frequency
temp2<-aggregate(NW~ltmode,FUN=sum,data=place_LAD)

## Generate weighted percentage
temp2$Percent<-temp2$NW*100/sum(temp2$NW)

## Names
names(temp2)

## Rename the variables
colnames(temp2)[c(1:3)] <- c("Mode","Number of Linked-Trips(W)","Percent(W)")

## Show
temp2

## Merge the two files and save
temp3<-merge(temp1,temp2, by ="Mode", all = TRUE,sort=FALSE)
temp3 <- temp3[,c(1,2,4,3,5)]
temp3

## Save as csv file
write.csv(temp3,file=pto("WlinkMode_LAD.csv"),row.names=FALSE)
la.city.lt.dest.la <-temp3

```

The next step is to compute a table with the total and fractional
linked trip duration for all trips originating in LA, by mode.

```{r la.city.lt.dur.dest.la}

## Total linked trip duration for all trips ending in LA, by mode
## Generate frequency for trip duration by trip mode
temp1<-aggregate(tripdur~ltmode,FUN=sum,data=place_LAD)

## Generate percentage
temp1$Percent<-temp1$tripdur*100/sum(temp1$tripdur)

## Names
names(temp1)

## Rename the variables
colnames(temp1)[c(1:3)] <- c("Mode","Linked-Trip Duration (Min)(UW)","Percent(UW)")

## Show
temp1

## Generate weighted frequency
temp<-place_LAD$NW*place_LAD$tripdur
temp2<-aggregate(temp~ltmode,FUN=sum,data=place_LAD)

## Generate weighted percentage
temp2$Percent<-temp2$temp*100/sum(temp2$temp)

## Names
names(temp2)

## Rename the variables
colnames(temp2)[c(1:3)] <- c("Mode","Linked-Trip Duration (Min)(W)","Percent(W)")

## Show
temp2

## Merge the two files and save
temp3<-merge(temp1,temp2, by ="Mode", all = TRUE,sort=FALSE)
temp3 <- temp3[,c(1,2,4,3,5)]
temp3

## Save as csv file
write.csv(temp3,file=pto("WlinkTRIPDUR_LAD.csv"),row.names=FALSE)
la.city.lt.dur.dest.la <- temp3
```

And the last step is to compute a table with the total and fractional
linked trip distance for all trips originating in LA, by mode.

```{r la.city.lt.dist.dest.la}
## Generate frequency for trip distance by mode
temp1<-aggregate(tripdist~ltmode,FUN=sum,data=place_LAD)

## Generate weighted percentage
temp1$Percent<-temp1$tripdist*100/sum(temp1$tripdist)

## Names
names(temp1)

## Rename the variables
colnames(temp1)[c(1:3)] <- c("Mode","Linked-Trip Distance (Miles)(UW)","Percent(UW)")

## Show
temp1

## Generate weighted frequency
temp<-place_LAD$NW*place_LAD$tripdist
temp2<-aggregate(temp~ltmode,FUN=sum,data=place_LAD)

## Generate weighted percentage
temp2$Percent<-temp2$temp*100/sum(temp2$temp)

## Names
names(temp2)

## Rename the variables
colnames(temp2)[c(1:3)] <- c("Mode","Linked-Trip Distance (miles)(W)","Percent(W)")

## Show
temp2

## Merge the two files and save
temp3<-merge(temp1,temp2, by ="Mode", all = TRUE,sort=FALSE)
temp3 <- temp3[,c(1,2,4,3,5)]
temp3

## Save as csv file
write.csv(temp3,file=pto("WlinkTRIPDIST_LAD.csv"),row.names=FALSE)
la.city.lt.dist.dest.la <- temp3
```

## Linked Trips Ending in LA but Originating Outside LA

The last group of trips we'll perform these computations for are those
ending in LA but originating outside LA.  Here is the code for
producing a table with the number and percent of linked trips by mode.

```{r la.city.lt.dest.la.o}


# Generate a subset for trips ending in LA but originating in outside LA
place_LAD2<-subset(linkplace,((dCity=="LOS ANGELES"|dCity=="LOS ÁNGELES") &
                              !(oCity=="LOS ANGELES"|oCity=="LOS ÁNGELES")))
head(place_LAD2,10)
names(place_LAD2)

## Total number of linked trips for all trips ending in LA but
## originating in outside LA, by mode Generate frequency for trip mode
temp1<-as.data.frame(table(place_LAD2$ltmode))

## Generate percentage of frequency
temp1$Percent<-temp1$Freq*100/sum(temp1$Freq)

## Names
names(temp1)

## Rename the variables
colnames(temp1)[c(1:3)] <- c("Mode","Number of Linked-Trips(UW)","Percent(UW)")

## Show the results
temp1

## Generate weighted frequency
temp2<-aggregate(EXPPERWGT~ltmode,FUN=sum,data=place_LAD2)

## Generate weighted percentage
temp2$Percent<-temp2$EXPPERWGT*100/sum(temp2$EXPPERWGT)

## Names
names(temp2)

## Rename the variables
colnames(temp2)[c(1:3)] <- c("Mode","Number of Linked-Trips(W)","Percent(W)")

## Show
temp2

## Merge the two files and save
temp3<-merge(temp1,temp2, by ="Mode", all = TRUE,sort=FALSE)
temp3 <- temp3[,c(1,2,4,3,5)]
temp3

## Save as csv file
write.csv(temp3,file=pto("WlinkMode_LAD2.csv"),row.names=FALSE)
la.city.lt.dest.la.o <- temp3

```

This is again followed by computing a table with the total and
fractional linked trip duration for all trips originating in LA, by
mode.

```{r la.city.lt.dur.dest.la.o}

## Total linked trip duration for all trips ending in LA but
## originating in outside LA, by mode Generate frequency for linked
## trip duration by trip mode
temp1<-aggregate(tripdur~ltmode,FUN=sum,data=place_LAD2)

## Generate percentage
temp1$Percent<-temp1$tripdur*100/sum(temp1$tripdur)

## Names
names(temp1)

## Rename the variables
colnames(temp1)[c(1:3)] <- c("Mode","Linked-Trip Duration (Min)(UW)","Percent(UW)")

## Show
temp1

## Generate weighted frequency
temp<-place_LAD2$EXPPERWGT*place_LAD2$tripdur
temp2<-aggregate(temp~ltmode,FUN=sum,data=place_LAD2)

## Generate weighted percentage
temp2$Percent<-temp2$temp*100/sum(temp2$temp)

## Names
names(temp2)

## Rename the variables
colnames(temp2)[c(1:3)] <- c("Mode","Linked-Trip Duration (Min)(W)","Percent(W)")

## Show
temp2

## Merge the two files and save
temp3<-merge(temp1,temp2, by ="Mode", all = TRUE,sort=FALSE)
temp3 <- temp3[,c(1,2,4,3,5)]
temp3

## Save as csv file
write.csv(temp3,file=pto("WlinkTRIPDUR_LAD2.csv"),row.names=FALSE)
la.city.lt.dur.dest.la.o <- temp3
```

Finally, we compute a table with the total and fractional linked trip distance for
all trips originating in LA, by mode.


```{r la.city.lt.dist.dest.la.o}

## Total linked trip distance for all trips ending in LA but
## originating in outside LA, by mode Generate frequency for linked
## trip distance by trip mode
temp1<-aggregate(tripdist~ltmode,FUN=sum,data=place_LAD2)

## Generate percentage
temp1$Percent<-temp1$tripdist*100/sum(temp1$tripdist)

## Names
names(temp1)

## Rename the variables
colnames(temp1)[c(1:3)] <- c("Mode","Linked-Trip Distance (Miles)(UW)","Percent(UW)")

## Show
temp1

## Generate weighted frequency
temp<-place_LAD2$EXPPERWGT*place_LAD2$tripdist
temp2<-aggregate(temp~ltmode,FUN=sum,data=place_LAD2)

## Generate weighted percentage
temp2$Percent<-temp2$temp*100/sum(temp2$temp)

## Names
names(temp2)

## Rename the variables
colnames(temp2)[c(1:3)] <- c("Mode","Linked-Trip Distance (Miles)(W)","Percent(W)")

## Show
temp2

## Merge the two files and save
temp3<-merge(temp1,temp2, by ="Mode", all = TRUE,sort=FALSE)
temp3 <- temp3[,c(1,2,4,3,5)]
temp3

## Save as csv file
write.csv(temp3,file=pto("WlinkTRIPDIST_LAD2.csv"),row.names=FALSE)
la.city.lt.dist.dest.la.o <- temp3
```


\fi

\newpage

# Results


```{r la-city-mode-code,echo=FALSE,results='asis'}
mmap <- nplacedata %>%
  transmute(MODE,Modestr=as.character(Mode)) %>%
  group_by(MODE,Modestr) %>% summarise(n=n())
pandoc.table.double(
    mmap %>%
    select(Code=MODE,Mode=Modestr),
    caption="Mode Code",
    justify=c('left','left'),
    split.cells=Inf,
    missing="&nbsp;")
```

\newpage

```{r la.city.lt.orig.la-res,results='asis',echo=FALSE}
pandoc.table( 
  la.city.lt.orig.la %>% mutate(Mode=as.integer(Mode)) %>%
    left_join(mmap, by=c('Mode'='MODE')) %>%
    mutate(Mode=Modestr) %>%
    select(-Modestr,-n) %>%
    arrange(desc(`Percent(W)`)) %>%
    mutate(`Number of Linked-Trips(UW)`=obsfmt(`Number of Linked-Trips(UW)`),
           `Number of Linked-Trips(W)`=obsfmt(`Number of Linked-Trips(W)`),
           `Percent(UW)`=pctfmt(`Percent(UW)`),
           `Percent(W)`=pctfmt(`Percent(W)`)),
  caption="Linked Trips Originating from the City of Los Angeles",
  justify=c('left',rep('right',4)),
  split.cells=c(50,rep(10,4)),
  split.table=Inf,
  missing="&nbsp;"
  )
```  


\newpage

```{r la.city.lt.dist.orig.la-res,results='asis',echo=FALSE}
pandoc.table( 
  la.city.lt.dist.orig.la %>% mutate(Mode=as.integer(Mode)) %>%
    left_join(mmap, by=c('Mode'='MODE')) %>%
    mutate(Mode=Modestr) %>%
    select(-Modestr,-n) %>%
    arrange(desc(`Percent(W)`)) %>%
    mutate(`Linked-Trip Distance (Miles)(UW)`=obsfmt(`Linked-Trip Distance (Miles)(UW)`),
           `Linked-Trip Distance (miles)(W)`=obsfmt(`Linked-Trip Distance (miles)(W)`),
           `Percent(UW)`=pctfmt(`Percent(UW)`),
           `Percent(W)`=pctfmt(`Percent(W)`)),
  caption="Total Linked-Trip Distance for Trips Originating from the City of Los Angeles",
  justify=c('left',rep('right',4)),
  split.cells=c(60,rep(8,4)),
  split.table=Inf,
  missing="&nbsp;"
  )
```  


\newpage

```{r la.city.lt.dur.orig.la-res,results='asis',echo=FALSE}
pandoc.table( 
  la.city.lt.dur.orig.la %>% mutate(Mode=as.integer(Mode)) %>%
    left_join(mmap, by=c('Mode'='MODE')) %>%
    mutate(Mode=Modestr) %>%
    select(-Modestr,-n) %>%
    arrange(desc(`Percent(W)`)) %>%
    mutate(`Linked-Trip Duration (Min)(UW)`=obsfmt(`Linked-Trip Duration (Min)(UW)`),
           `Linked-Trip Duration(Min)(W)`=obsfmt(`Linked-Trip Duration(Min)(W)`),
           `Percent(UW)`=pctfmt(`Percent(UW)`),
           `Percent(W)`=pctfmt(`Percent(W)`)),
  caption="Total Linked-Trip Distance for Trips Originating from the City of Los Angeles",
  justify=c('left',rep('right',4)),
  split.cells=c(50,rep(8,4)),
  split.table=Inf,
  missing="&nbsp;"
  )
```  



\newpage

```{r la.city.lt.dest.la-res,results='asis',echo=FALSE}
pandoc.table( 
  la.city.lt.dest.la %>% mutate(Mode=as.integer(Mode)) %>%
    left_join(mmap, by=c('Mode'='MODE')) %>%
    mutate(Mode=Modestr) %>%
    select(-Modestr,-n) %>%
    arrange(desc(`Percent(W)`)) %>%
    mutate(`Number of Linked-Trips(UW)`=obsfmt(`Number of Linked-Trips(UW)`),
           `Number of Linked-Trips(W)`=obsfmt(`Number of Linked-Trips(W)`),
           `Percent(UW)`=pctfmt(`Percent(UW)`),
           `Percent(W)`=pctfmt(`Percent(W)`)),
  caption="Linked Trips Terminating in the City of Los Angeles",
  justify=c('left',rep('right',4)),
  split.cells=c(50,rep(10,4)),
  split.table=Inf,
  missing="&nbsp;"
  )
```  


\newpage

```{r la.city.lt.dist.dest.la-res,results='asis',echo=FALSE}
pandoc.table( 
  la.city.lt.dist.dest.la %>% mutate(Mode=as.integer(Mode)) %>%
    left_join(mmap, by=c('Mode'='MODE')) %>%
    mutate(Mode=Modestr) %>%
    select(-Modestr,-n) %>%
    arrange(desc(`Percent(W)`)) %>%
    mutate(`Linked-Trip Distance (Miles)(UW)`=obsfmt(`Linked-Trip Distance (Miles)(UW)`),
           `Linked-Trip Distance (miles)(W)`=obsfmt(`Linked-Trip Distance (miles)(W)`),
           `Percent(UW)`=pctfmt(`Percent(UW)`),
           `Percent(W)`=pctfmt(`Percent(W)`)),
  caption="Total Linked-Trip Distance for Trips Terminating in the City of Los Angeles",
  justify=c('left',rep('right',4)),
  split.cells=c(60,rep(8,4)),
  split.table=Inf,
  missing="&nbsp;"
  )
```  


\newpage

```{r la.city.lt.dur.dest.la-res,results='asis',echo=FALSE}
pandoc.table(
  la.city.lt.dur.dest.la %>% mutate(Mode=as.integer(Mode)) %>%
    left_join(mmap, by=c('Mode'='MODE')) %>%
    mutate(Mode=Modestr) %>%
    select(-Modestr,-n) %>%
    arrange(desc(`Percent(W)`)) %>%
    transmute(
      Mode=Mode,
      `Linked-Trip Duration (Min)(UW)`=obsfmt(`Linked-Trip Duration (Min)(UW)`),
      `Linked-Trip Duration (Min)(W)`=obsfmt(`Linked-Trip Duration (Min)(W)`),
      `Percent(UW)`=pctfmt(`Percent(UW)`),
      `Percent(W)`=pctfmt(`Percent(W)`)),
  caption="Linked-Trip Duration for Trips Terminating from the City of Los Angeles",
  justify=c('left',rep('right',4)),
  split.cells=c(50,rep(8,4)),
  split.table=Inf,
  missing="&nbsp;"
  )
```  





\newpage

```{r la.city.lt.od.la-res,results='asis',echo=FALSE}
pandoc.table( 
  la.city.lt.od.la %>% mutate(Mode=as.integer(Mode)) %>%
    left_join(mmap, by=c('Mode'='MODE')) %>%
    mutate(Mode=Modestr) %>%
    select(-Modestr,-n) %>%
    arrange(desc(`Percent(W)`)) %>%
    mutate(`Number of Linked-Trips(UW)`=obsfmt(`Number of Linked-Trips(UW)`),
           `Number of Linked-Trips(W)`=obsfmt(`Number of Linked-Trips(W)`),
           `Percent(UW)`=pctfmt(`Percent(UW)`),
           `Percent(W)`=pctfmt(`Percent(W)`)),
  caption="Linked Trips Originating and Terminating in the City of Los Angeles",
  justify=c('left',rep('right',4)),
  split.cells=c(50,rep(10,4)),
  split.table=Inf,
  missing="&nbsp;"
  )
```  


\newpage

```{r la.city.lt.dist.od.la-res,results='asis',echo=FALSE}
pandoc.table( 
  la.city.lt.dist.od.la %>% mutate(Mode=as.integer(Mode)) %>%
    left_join(mmap, by=c('Mode'='MODE')) %>%
    mutate(Mode=Modestr) %>%
    select(-Modestr,-n) %>%
    arrange(desc(`Percent(W)`)) %>%
    mutate(`Linked-Trip Distance (Miles)(UW)`=obsfmt(`Linked-Trip Distance (Miles)(UW)`),
           `Linked-Trip Distance (Miles)(W)`=obsfmt(`Linked-Trip Distance (Miles)(W)`),
           `Percent(UW)`=pctfmt(`Percent(UW)`),
           `Percent(W)`=pctfmt(`Percent(W)`)),
  caption=paste("Total Linked-Trip Distance for Trips Originating ",
                "and Terminating in the City of Los Angeles"),
  justify=c('left',rep('right',4)),
  split.cells=c(60,rep(8,4)),
  split.table=Inf,
  missing="&nbsp;"
  )
```  


\newpage

```{r la.city.lt.dur.od.la-res,results='asis',echo=FALSE}
pandoc.table( 
  la.city.lt.dur.od.la %>% mutate(Mode=as.integer(Mode)) %>%
    left_join(mmap, by=c('Mode'='MODE')) %>%
    mutate(Mode=Modestr) %>%
    select(-Modestr,-n) %>%
    arrange(desc(`Percent(W)`)) %>%
    transmute(
      Mode=Mode,
      `Linked-Trip Duration (Min)(UW)`=obsfmt(`Linked-Trip Duration (Min)(UW)`),
      `Linked-Trip Duration (Min)(W)`=obsfmt(`Linked-Trip Duration (Min)(W)`),
      `Percent(UW)`=pctfmt(`Percent(UW)`),
      `Percent(W)`=pctfmt(`Percent(W)`)),
  caption=paste("Linked-Trip Duration for Trips Originating ",
                "and Terminating from the City of Los Angeles"),
  justify=c('left',rep('right',4)),
  split.cells=c(50,rep(8,4)),
  split.table=Inf,
  missing="&nbsp;"
  )
```  


\newpage

```{r la.city.lt.dest.la.o-res,results='asis',echo=FALSE}
pandoc.table( 
  la.city.lt.dest.la.o %>% mutate(Mode=as.integer(Mode)) %>%
    left_join(mmap, by=c('Mode'='MODE')) %>%
    mutate(Mode=Modestr) %>%
    select(-Modestr,-n) %>%
    arrange(desc(`Percent(W)`)) %>%
    mutate(`Number of Linked-Trips(UW)`=obsfmt(`Number of Linked-Trips(UW)`),
           `Number of Linked-Trips(W)`=obsfmt(`Number of Linked-Trips(W)`),
           `Percent(UW)`=pctfmt(`Percent(UW)`),
           `Percent(W)`=pctfmt(`Percent(W)`)),
  caption="Linked Trips Originating Elsewhere and Terminating in the City of Los Angeles",
  justify=c('left',rep('right',4)),
  split.cells=c(50,rep(10,4)),
  split.table=Inf,
  missing="&nbsp;"
  )
```  


\newpage

```{r la.city.lt.dist.dest.la.o-res,results='asis',echo=FALSE}
pandoc.table( 
  la.city.lt.dist.dest.la.o %>% mutate(Mode=as.integer(Mode)) %>%
    left_join(mmap, by=c('Mode'='MODE')) %>%
    mutate(Mode=Modestr) %>%
    select(-Modestr,-n) %>%
    arrange(desc(`Percent(W)`)) %>%
    mutate(`Linked-Trip Distance (Miles)(UW)`=obsfmt(`Linked-Trip Distance (Miles)(UW)`),
           `Linked-Trip Distance (Miles)(W)`=obsfmt(`Linked-Trip Distance (Miles)(W)`),
           `Percent(UW)`=pctfmt(`Percent(UW)`),
           `Percent(W)`=pctfmt(`Percent(W)`)),
  caption=paste("Total Linked-Trip Distance for Trips Originating ",
                "Elsewhere and Terminating in the City of Los Angeles"),
  justify=c('left',rep('right',4)),
  split.cells=c(60,rep(8,4)),
  split.table=Inf,
  missing="&nbsp;"
  )
```  


\newpage

```{r la.city.lt.dur.dest.la.o-res,results='asis',echo=FALSE}
pandoc.table( 
  la.city.lt.dur.dest.la.o %>% mutate(Mode=as.integer(Mode)) %>%
    left_join(mmap, by=c('Mode'='MODE')) %>%
    mutate(Mode=Modestr) %>%
    select(-Modestr,-n) %>%
    arrange(desc(`Percent(W)`)) %>%
    transmute(
      Mode=Mode,
      `Linked-Trip Duration (Min)(UW)`=obsfmt(`Linked-Trip Duration (Min)(UW)`),
      `Linked-Trip Duration (Min)(W)`=obsfmt(`Linked-Trip Duration (Min)(W)`),
      `Percent(UW)`=pctfmt(`Percent(UW)`),
      `Percent(W)`=pctfmt(`Percent(W)`)),
  caption=paste("Linked-Trip Duration for Trips Originating Elsewhere ",
                "and Terminating from the City of Los Angeles"),
  justify=c('left',rep('right',4)),
  split.cells=c(50,rep(8,4)),
  split.table=Inf,
  missing="&nbsp;"
  )
```  
