---
title: "Data Request for Santa Cruz County"
author: "Suman Mitra"
date: 'This version: `r (function() {f<-dir(); format(max(file.mtime(f[grepl("^.*\\.Rmd$",f)])), format="%B %d, %Y")})()`'
output:
  pdf_document:
    keep_tex: yes
    template: custom.latex
fontfamily: inconsolata
fontsize: 11pt
geometry: margin=1in
params:
  documentclass: article
  documentation: '`r if(exists("documentation")){documentation}else{documentation<-"false";documentation}`'
---

```{r scc-setup-0,echo=FALSE,cache=FALSE}
if ( !exists("rd") ) { rd <<- gsub("(.*?chts-book[^/]*).*$","\\1",getwd()) } 
source(paste(rd,"R/setup-hooks.R",sep="/"))
write(paste("Using Root Directory",rd,"\n"),stderr())
```

# Question

This question asked for the following information for Santa Cruz county:

1. Trips by Mode
2. Modal split for driving alone
3. Trip distance by Mode
4. Trip duration by Mode
5. Trips by Purpose
6. Work trips by mode

# Summary and Methodology

In this sub-section, we generate
tables by mode for the following: total number of trips, total trip
distance, average trip distance, total trip duration, average trip
duration, and total number of trips for driving alone.  To calculate
the "drive alone"" numbers we extract data for total travelers
\ifmydocumentation(`TOTTR=1`)\fi and drivers of cars, trucks, or
motorcycles\ifmydocumentation (`MODE={5, 6, 7, 8, 10}`)\fi. We also
generate tables for the total number of trips by trip purpose and for
work trips by mode. Here, we actually generate tables for the total
number of trips by activity purpose since the trip purpose is given in
the database as the activity purpose and although many trips are
associated with more than one activity, We consider each activity as a
trip. We only consider the work trips \ifmydocumentation
(`APURP=9`)\fi.  We generate every table for both trips originating in
and ending in Santa Cruz County. We use the new county wide weights of
Santa Cruz County to generate weighted results for trips that
originated in Santa Cruz County and statewide weights to generate
weighted results for trips that originated outside of Santa Cruz
County.

\ifmydocumentation

# R Code and Implementation


The following commands are used to generate the results.

## Trips to  Santa Cruz County


### Mode split calculations


To get started we need the place level table with weights
that we computed previously.

```{r scc-1 }
# Set the working directory
#setwd(p("Santa Cruz County"))


#Import the place level with new weight csv file and give a new name
nplace<- read.csv(pd("Weight/NewWeightData/nplacedata932015.csv"), header=TRUE)
names(nplace)
head(nplace)# Show first ten observations

```

Next, we look for the origin county and convert the `MODE` codes into 
text descriptions.

```{r scc-2}
# Create trip origin county 
# Need dplyr
library("dplyr")
nplacedata <- nplace %>%
    group_by(SAMPN,PERNO) %>%
    arrange(SAMPN,PERNO,PLANO) %>%
    mutate(OCTFIP=lag(CTFIP))

# For recoding Mode
nplacedata$Mode <-
  ifelse(nplacedata$MODE==1,"Walk",
  ifelse(nplacedata$MODE==2,"Bike",
  ifelse(nplacedata$MODE==3,"Wheelchair/Mobility Scooter",
  ifelse(nplacedata$MODE==4,"Other Non-Motorized",
  ifelse(nplacedata$MODE==5,"Auto/Van/Truck Driver",
  ifelse(nplacedata$MODE==6,"Auto /Van/Truck Passenger",
  ifelse(nplacedata$MODE==7,"Carpool/Vanpool",
  ifelse(nplacedata$MODE==8,"Motorcycle/Scooter/Moped",
  ifelse(nplacedata$MODE==9,"Taxi/Hired Car/Limo",
  ifelse(nplacedata$MODE==10,"Rental Car/Vehicle",
  ifelse(nplacedata$MODE==11,"Private shuttle ",
  ifelse(nplacedata$MODE==12,"Greyhound Bus",
  ifelse(nplacedata$MODE==13,"Plane",
  ifelse(nplacedata$MODE==14,"Other Private Transit",
  ifelse(nplacedata$MODE==15,"Local Bus,Rapid Bus",
  ifelse(nplacedata$MODE==16,"Express Bus/Commuter Bus",
  ifelse(nplacedata$MODE==17,"Premium Bus",
  ifelse(nplacedata$MODE==18,"School Bus",
  ifelse(nplacedata$MODE==19,"Public Transit Shuttle",
  ifelse(nplacedata$MODE==20,"Air BART/LAX Fly Away",
  ifelse(nplacedata$MODE==21,"Dial-a-Ride/Paratransit", 
  ifelse(nplacedata$MODE==22,"Amtrak Bus", 
  ifelse(nplacedata$MODE==23,"Other Bus ", 
  ifelse(nplacedata$MODE==24,"BART,Metro Red/Purple Line",
  ifelse(nplacedata$MODE==25,"ACE, Amtrak, Cal train",
  ifelse(nplacedata$MODE==26,"Metro",
  ifelse(nplacedata$MODE==27,"Street Car/Cable Car",
  ifelse(nplacedata$MODE==28,"Other Rail",
  ifelse(nplacedata$MODE==29,"Ferry / Boat",
         "Don't Know/Refused"
         )))))))))))))))))))))))))))))

```

Let us now focus on trips ending in Santa Cruz County.  We use
different weights depending on the origin of a trip.  Specifically,
when the origin is Santa Cruz County, we use county-specific
weights and when the origin is outside Santa Cruz County, we use the
statewide weights.

```{r scc-3}
# Trips to  Santa Cruz County
# Create a subset for trips to  Santa Cruz county 
# We are only interested in trips to  Santa Cruz County (CTFIP=87) 
# Create a subset for trips to  Santa Cruz County 
place_87<-subset(nplacedata,CTFIP==87)
names(place_87)

# Change the weight for non Santa Cruz County origin
# (Assign state wide weight for non  Santa Cruz origin)
place_87$NTCFW <-
    ifelse((place_87$OCTFIP==87),
           place_87$NEXPTCFPERWGT,
           place_87$EXPTCFPERWGT)
head(place_87)
```

At this point, `place_87$NTCFW` contains the trips ending in Santa Cruz County
with appropriate expansion weights appended.  Now we can generate
observations for modal split for trips to Santa Cruz based on these
weights.

Notes: In the tables: Ob = Observations, Per = Percent, UW =
Unweighted and W = Weighted.

```{r scc-4}
## Generate frequency for trip Mode
temp1<-as.data.frame(table(place_87$Mode))
## Create percentage of frequency
temp1$Percent<-temp1$Freq*100/sum(temp1$Freq)
## Names 
names(temp1)
## Rename the variables
colnames(temp1)[c(1:3)] <- c("Mode","Ob_UW","Per_UW")
## Show the results
temp1
```

Now we compute weighted frequencies:

```{r scc-5}
## Create weighted frequency
temp2<-aggregate(NTCFW~Mode,FUN=sum,data=place_87)
## Create weighted percentage
temp2$Percent<-temp2$NTCFW*100/sum(temp2$NTCFW)
## Names
names(temp2)
## Rename the variables
colnames(temp2)[c(1:3)] <- c("Mode","Ob_W","Per_W")
## Show
temp2
```

Let us save these results in a `csv` file for later use:

```{r scc-6}
## Merge the two files and save
temp3<-merge(temp1,temp2, by ="Mode", all = TRUE,sort=TRUE)
temp3 <- temp3[,c(1,2,4,3,5)]
temp3
Mode_87d<-temp3
temp3[,1]<-as.character(temp3[,1])
temp3<- rbind(temp3, c("Total", colSums(temp3[,2:5],na.rm = TRUE)))
temp3
## Save as csv file
write.csv(temp3,file=pto("Mode_87d.csv"),row.names=FALSE)
mode.scc.dest <- temp3
rm(temp1)
rm(temp2)
rm(temp3)
```

### Trip duration by mode


Next, we perform similar computations for Total Trip duration for all
trips ending in Santa Cruz, by mode.

```{r scc-7}
## Generate frequency for trip Mode
temp1<-aggregate(TRIPDUR~Mode,FUN=sum,data=place_87)
## Create  percentage
temp1$Percent<-temp1$TRIPDUR*100/sum(temp1$TRIPDUR)
## Names
names(temp1)
## Rename the variables
colnames(temp1)[c(1:3)] <- c("Mode","TDu_UW","Per_UW")
## Show
temp1
```

Let us compute weighted frequencies:

```{r scc-8}
## Create weighted frequency
temp<-place_87$NTCFW*place_87$TRIPDUR
temp2<-aggregate(temp~Mode,FUN=sum,data=place_87)
## Create weighted percentage
temp2$Percent<-temp2$temp*100/sum(temp2$temp)
## Names
names(temp2)
## Rename the variables
colnames(temp2)[c(1:3)] <- c("Mode","TDu_W","Per_W")
## Show
temp2
```


Now save the results for later use:

```{r scc-9}
## Merge the two files and save
temp3<-merge(temp1,temp2, by ="Mode", all = TRUE)
temp3 <- temp3[,c(1,2,4,3,5)]
# Average trip duration
temp3$avg_uw<-temp3$TDu_UW/Mode_87d$Ob_UW
temp3
temp3$avg_w<-temp3$TDu_W/Mode_87d$Ob_W
temp3
temp3[,1]<-as.character(temp3[,1])
temp3<- rbind(temp3, c("Total", colSums(temp3[,2:7],na.rm = TRUE)))
temp3
## Save as csv file
write.csv(temp3,file=pto("TripDUrm_D87.csv"),row.names=FALSE)
dur.scc.dest <- temp3
rm(temp1)
rm(temp2)
rm(temp3)
rm(temp)
```

### Trip distance by mode


Moving on, let us calculate Total Trip distance for all trips ending in
Santa Cruz, by mode:

```{r scc-10}
## Generate frequency for trip Mode
temp1<-aggregate(TripDistance~Mode,FUN=sum,data=place_87)
## Create  percentage
temp1$Percent<-temp1$TripDistance*100/sum(temp1$TripDistance)
## Names
names(temp1)
## Rename the variables
colnames(temp1)[c(1:3)] <- c("Mode","TDis_UW","Per_UW")
## Show
temp1
```

Next, let us compute weighted trip distances by mode:

```{r scc-11}
## Create weighted frequency
temp<-place_87$NTCFW*place_87$TripDistance
temp2<-aggregate(temp~Mode,FUN=sum,data=place_87)
## Create weighted percentage
temp2$Percent<-temp2$temp*100/sum(temp2$temp)
## Names
names(temp2)
## Rename the variables
colnames(temp2)[c(1:3)] <- c("Mode","TDis_W","Per_W")
## Show
temp2
```

Save the results for later use in a csv file:

```{r scc-12}
## Merge the two files and save
temp3<-merge(temp1,temp2, by ="Mode", all = TRUE)
temp3 <- temp3[,c(1,2,4,3,5)]
# Average trip distance
temp3$avg_uw<-temp3$TDis_UW/Mode_87d$Ob_UW
temp3
temp3$avg_w<-temp3$TDis_W/Mode_87d$Ob_W
temp3
temp3[,1]<-as.character(temp3[,1])
temp3<- rbind(temp3, c("Total", colSums(temp3[,2:7],na.rm = TRUE)))
temp3
## Save as csv file
write.csv(temp3,file=pto("TripDistanceM_D87.csv"),row.names=FALSE)
dist.scc.dest <- temp3
rm(temp1)
rm(temp2)
rm(temp3)
rm(temp)
```

### Modal split for driving alone


The next question was the modal split for Drive Alone.  To estimate
this, we subset the data based on `TOTTR` (Total travelers) and
`MODE={5, 6, 7, 8, 10}` (basically, the driver of a car, truck, or
motorcycle):

```{r scc-13}
place_87d <-
    subset(
      place_87,
      (TOTTR==1 & (MODE==5|MODE==6|MODE==7|MODE==8|MODE==10)))
head(place_87d)
```

Now let us find mode frequency for this subset of the data:

```{r scc-14}
## Create frequency for different Mode category for drive alone
temp1<-as.data.frame(table(place_87d$Mode))
## Create percentage of frequency
temp1$Percent<-temp1$Freq*100/sum(temp1$Freq)
## Names 
names(temp1)
## Rename the variables
colnames(temp1)[c(1:3)] <- c("Mode","Ob_UW","Per_UW")
## Show the results
temp1
```

Following the same approach as above, let us find weighted frequencies:

```{r scc-15}
## Create weighted frequency
temp2<-aggregate(NTCFW~Mode,FUN=sum,data=place_87d)
## Create weighted percentage
temp2$Percent<-temp2$NTCFW*100/sum(temp2$NTCFW)
## Names
names(temp2)
## Rename the variables
colnames(temp2)[c(1:3)] <- c("Mode","Ob_W","Per_W")
## Show
temp2
```

Let us now save the results in a csv file for later use:

```{r scc-16}

## Merge the two files and save
temp3<-merge(temp1,temp2, by ="Mode", all = TRUE,sort=FALSE)
temp3 <- temp3[,c(1,2,4,3,5)]
temp3
temp3[,1]<-as.character(temp3[,1])
temp3<- rbind(temp3, c("Total", colSums(temp3[,2:5],na.rm = TRUE)))
temp3
## Save as csv file
write.csv(temp3,file=pto("Mode_87Ddrialn.csv"),row.names=FALSE)
mode.da.scc.dest <- temp3
rm(temp1)
rm(temp2)
rm(temp3)
```


## Trips from  Santa Cruz County


We proceed the same way for trips *from* Santa Cruz County.  
As before, we subset the place data, but this time for trips
originating in Santa Cruz County.

```{r scc-17}

# Let us estimate trips from  Santa Cruz county (OCTFIP=87) 
# Create a subset for trips from  Santa Cruz county 
place_87O<-subset(nplacedata,OCTFIP==87)
names(place_87O)
head(place_87O)
```

### Trips by mode


Now we generate the mode frequency table for all trips from Santa Cruz
County:

```{r scc-18}
# Generate observations for modal split for trips from  Santa Cruz based on new weights
## Generate frequency for trip Mode
temp1<-as.data.frame(table(place_87O$Mode))
## Create percentage of frequency
temp1$Percent<-temp1$Freq*100/sum(temp1$Freq)
## Names 
names(temp1)
## Rename the variables
colnames(temp1)[c(1:3)] <- c("Mode","Ob_UW","Per_UW")
## Show the results
temp1
```

This should look familiar now. Here are the weighted frequencies:

```{r scc-19}
## Create weighted frequency
temp2<-aggregate(NEXPTCFPERWGT~Mode,FUN=sum,data=place_87O)
## Create weighted percentage
temp2$Percent<-temp2$NEXPTCFPERWGT*100/sum(temp2$NEXPTCFPERWGT)
## Names
names(temp2)
## Rename the variables
colnames(temp2)[c(1:3)] <- c("Mode","Ob_W","Per_W")
## Show
temp2
```

And again, we save the results in a csv file for later use:

```{r scc-20}
## Merge the two files and save
temp3<-merge(temp1,temp2, by ="Mode", all = TRUE,sort=FALSE)
temp3 <- temp3[,c(1,2,4,3,5)]
temp3
Mode_87O<-temp3
temp3[,1]<-as.character(temp3[,1])
temp3<- rbind(temp3, c("Total", colSums(temp3[,2:5],na.rm = TRUE)))
temp3
## Save as csv file
write.csv(temp3,file=pto("Mode_87O.csv"),row.names=FALSE)
mode.scc.orig <- temp3
rm(temp1)
rm(temp2)
rm(temp3)
```

#### Trip duration by Mode

Let us repeat the trip duration calculation for Santa Cruz Origins:

```{r scc-21}
## Total Trip duration for all trips from  Santa Cruz, by Mode
## Generate frequency for trip Mode
temp1<-aggregate(TRIPDUR~Mode,FUN=sum,data=place_87O)
## Create  percentage
temp1$Percent<-temp1$TRIPDUR*100/sum(temp1$TRIPDUR)
## Names
names(temp1)
## Rename the variables
colnames(temp1)[c(1:3)] <- c("Mode","TDu_UW","Per_UW")
## Show
temp1
```

Here are the calculations for weighted frequencies:

```{r scc-22}
## Create weighted frequency
temp<-place_87O$NEXPTCFPERWGT*place_87O$TRIPDUR
temp2<-aggregate(temp~Mode,FUN=sum,data=place_87O)
## Create weighted percentage
temp2$Percent<-temp2$temp*100/sum(temp2$temp)
## Names
names(temp2)
## Rename the variables
colnames(temp2)[c(1:3)] <- c("Mode","TDu_W","Per_W")
## Show
temp2
```

Save the results in "TripDUrm_O.csv":"

```{r scc-23}
## Merge the two files and save
temp3<-merge(temp1,temp2, by ="Mode", all = TRUE)
temp3 <- temp3[,c(1,2,4,3,5)]
temp3
temp3$avg_uw<-temp3$TDu_UW/Mode_87O$Ob_UW
temp3
temp3$avg_w<-temp3$TDu_W/Mode_87O$Ob_W
temp3
temp3[,1]<-as.character(temp3[,1])
temp3<- rbind(temp3, c("Total", colSums(temp3[,2:7],na.rm = TRUE)))
temp3
## Save as csv file
write.csv(temp3,file=pto("TripDUrm_O.csv"),row.names=FALSE)
dur.scc.orig <- temp3
rm(temp1)
rm(temp2)
rm(temp3)
rm(temp)
```

### Trip distance by Mode

Here are the trip distance calculations for all trips from Santa Cruz,
by Mode:

```{r scc-24}
## Generate frequency for trip Mode
temp1<-aggregate(TripDistance~Mode,FUN=sum,data=place_87O)
## Create  percentage
temp1$Percent<-temp1$TripDistance*100/sum(temp1$TripDistance)
## Names
names(temp1)
## Rename the variables
colnames(temp1)[c(1:3)] <- c("Mode","TDis_UW","Per_UW")
## Show
temp1
```

This is followed by the weighted frequency calculations:

```{r scc-25}
## Create weighted frequency
temp<-place_87O$NEXPTCFPERWGT*place_87O$TripDistance
temp2<-aggregate(temp~Mode,FUN=sum,data=place_87O)
## Create weighted percentage
temp2$Percent<-temp2$temp*100/sum(temp2$temp)
## Names
names(temp2)
## Rename the variables
colnames(temp2)[c(1:3)] <- c("Mode","TDis_W","Per_W")
## Show
temp2
```

and by commands to save the data in "TripDistanceM_O.csv":

```{r scc-26}
## Merge the two files and save
temp3<-merge(temp1,temp2, by ="Mode", all = TRUE)
temp3 <- temp3[,c(1,2,4,3,5)]
temp3
temp3$avg_uw<-temp3$TDis_UW/Mode_87O$Ob_UW
temp3
temp3$avg_w<-temp3$TDis_W/Mode_87O$Ob_W
temp3
temp3[,1]<-as.character(temp3[,1])
temp3<- rbind(temp3, c("Total", colSums(temp3[,2:7],na.rm = TRUE)))
temp3
## Save as csv file
write.csv(temp3,file=pto("TripDistanceM_O.csv"),row.names=FALSE)
dist.scc.orig <- temp3
rm(temp1)
rm(temp2)
rm(temp3)
rm(temp)
```

### Modal split for driving alone


Lastly for Santa Cruz origins, let us generate the drive alone modal
split.  As before, we subset the data based on `TOTTR` (Total
travelers) and `MODE={5, 6, 7, 8, 10}` (the driver of a car, truck, or
motorcycle):

```{r scc-27} 
place_87dO <-
  subset(
    place_87O,
    (TOTTR==1 & (MODE==5|MODE==6|MODE==7|MODE==8|MODE==10)))
head(place_87dO)
```

The raw frequency table is created below:

```{r scc-28}
## Create frequency for different mode category for drive alone
temp1<-as.data.frame(table(place_87dO$Mode))
## Create percentage of frequency
temp1$Percent<-temp1$Freq*100/sum(temp1$Freq)
## Names 
names(temp1)
## Rename the variables
colnames(temp1)[c(1:3)] <- c("Mode","Observations_UW","Percent_UW")
## Show the results
temp1
```

We now generate the weighted frequency table:

```{r scc-29}
## Create weighted frequency
temp2<-aggregate(NEXPTCFPERWGT~Mode,FUN=sum,data=place_87dO)
## Create weighted percentage
temp2$Percent<-temp2$NEXPTCFPERWGT*100/sum(temp2$NEXPTCFPERWGT)
## Names
names(temp2)
## Rename the variables
colnames(temp2)[c(1:3)] <- c("Mode","Observations_W","Percent_W")
## Show
temp2
```

and save the results in "Mode_87Odrialn.csv":

```{r scc-30}
## Merge the two files and save
temp3<-merge(temp1,temp2, by ="Mode", all = TRUE,sort=FALSE)
temp3 <- temp3[,c(1,2,4,3,5)]
temp3
temp3[,1]<-as.character(temp3[,1])
temp3<- rbind(temp3, c("Total", colSums(temp3[,2:5],na.rm = TRUE)))
temp3
## Save as csv file
write.csv(temp3,file=pto("Mode_87Odrialn.csv"),row.names=FALSE)
mode.da.scc.orig <- temp3
rm(temp1)
rm(temp2)
rm(temp3)
```

## Trip Purpose Calculations


Now we need the trip purpose breakdown, so we import the activity level data with the 
appropriate weights. Let us load and merge tthese two datasets:

```{r scc-31}
library(chts2011)
#Import the activity level data frame and give a new name
act_data <- chts_act
names(act_data)

# Merge the person weight file with place level data file
act_place<-merge(act_data,nplacedata, by =c("SAMPN","PERNO","PLANO"), all = FALSE)
names(act_place)
```

For display purposes, we convert the purpose codes into readable strings:

```{r scc-32 }
act_place$Actpur<-
  ifelse(act_place$APURP==1," Personal Activities",
  ifelse(act_place$APURP==2," Preparing Meals/Eating",
  ifelse(act_place$APURP==3," Hosting Visitors/Entertaining Guests",
  ifelse(act_place$APURP==4," Exercise ",
  ifelse(act_place$APURP==5," Study/Schoolwork ",
  ifelse(act_place$APURP==6," Work For Pay At Home Using ",
  ifelse(act_place$APURP==7," Using Computer/Telephone/Cell Or Smart Phone",
  ifelse(act_place$APURP==8," All Other Activities At My Home ",
  ifelse(act_place$APURP==9," Work/Job Duties ",
  ifelse(act_place$APURP==10," Training",
  ifelse(act_place$APURP==11," Meals At Work ",
  ifelse(act_place$APURP==12," Work-Sponsored Social Activities ",
  ifelse(act_place$APURP==13," Non-Work Related Activities",
  ifelse(act_place$APURP==14," Exercise/Sports ",
  ifelse(act_place$APURP==15," Volunteer Work/Activities",
  ifelse(act_place$APURP==16," All Other Work-Related Activities At My Work ",
  ifelse(act_place$APURP==17," In School/Classroom/Laboratory ",
  ifelse(act_place$APURP==18," Meals At School/College ",
  ifelse(act_place$APURP==19," After School Sports/Physical Activity ",
  ifelse(act_place$APURP==20," All Other After School Related Activities ",
  ifelse(act_place$APURP==21," Change Type Of Transportation/Transfer ", 
  ifelse(act_place$APURP==22," Pickup/Drop Off Passenger(S)", 
  ifelse(act_place$APURP==23," Drive Through Meals ", 
  ifelse(act_place$APURP==24," Drive Through Other (ATM, Bank) ",
  ifelse(act_place$APURP==25," Work-Related",
  ifelse(act_place$APURP==26," Service Private Vehicle",
  ifelse(act_place$APURP==27," Routine Shopping ",
  ifelse(act_place$APURP==28," Shopping For Major Purchases",
  ifelse(act_place$APURP==29," Household Errands", 
  ifelse(act_place$APURP==30," Personal Business ", 
  ifelse(act_place$APURP==31," Eat Meal at Restaurant/Diner", 
  ifelse(act_place$APURP==32," Health Care", 
  ifelse(act_place$APURP==33," Civic/Religious Activities", 
  ifelse(act_place$APURP==34," Outdoor Exercise", 
  ifelse(act_place$APURP==35," Indoor Exercise", 
  ifelse(act_place$APURP==36," Entertainment ", 
  ifelse(act_place$APURP==37," Social/Visit Friends/Relatives", 
  ifelse(act_place$APURP==38," Other ", 
  ifelse(act_place$APURP==39," Loop Trip", 
         " Don't Know/Refused"
         )))))))))))))))))))))))))))))))))))))))
head(act_place)
```

## Trip Purposes from  Santa Cruz County

We start with trips from Santa Cruz County, subsetting based on origin county `OCTFIP=87`.

```{r scc-33 }
# Now we are interested in trips from  Santa Cruz county (OCTFIP=87) 
# Create a subset for trips from  Santa Cruz County 
act_87O<-subset(act_place,OCTFIP==87)
names(act_87O)
head(act_87O)
```

Next, we compute the observations for trip purpose for trips from
Santa Cruz and use new weights:

```{r scc-34}
## Generate frequency for trip Mode
temp1<-as.data.frame(table(act_87O$Actpur))
## Create percentage of frequency
temp1$Percent<-temp1$Freq*100/sum(temp1$Freq)
## Names 
names(temp1)
## Rename the variables
colnames(temp1)[c(1:3)] <- c("Actpur","Ob_UW","Per_UW")
## Show the results
temp1
```

Next let us generate weighted trip purpose frequencies:

```{r scc-35}
## Create weighted frequency
temp2<-aggregate(NEXPTCFPERWGT~Actpur,FUN=sum,data=act_87O)
## Create weighted percentage
temp2$Percent<-temp2$NEXPTCFPERWGT*100/sum(temp2$NEXPTCFPERWGT)
## Names
names(temp2)
## Rename the variables
colnames(temp2)[c(1:3)] <- c("Actpur","Ob_W","Per_W")
## Show
temp2
```

And now let us save the results in "Actpur_87O.csv":

```{r scc-36}
## Merge the two files and save
temp3<-merge(temp1,temp2, by ="Actpur", all = TRUE,sort=FALSE)
temp3 <- temp3[,c(1,2,4,3,5)]
temp3
temp3[,1]<-as.character(temp3[,1])
temp3<- rbind(temp3, c("Total", colSums(temp3[,2:5],na.rm = TRUE)))
temp3
## Save as csv file
write.csv(temp3,file=pto("Actpur_87O.csv"),row.names=FALSE)
apurp.scc.orig <- temp3
rm(temp1)
rm(temp2)
rm(temp3)
```


## Work Trips by Mode


In the remainder of this sub question, let us generate work trips by Mode for Santa Cruz County:


### Work Trips Originating in Santa Cruz County


First, we reuse the merged activity table from above for trips
activities (and associated trips) originating in Santa Cruz County,
and subset it for work trips only:

```{r scc-37}
# Create a subset for work trips
actw_87O<-subset(act_87O,APURP==9)
```

From this information we can find observations of each Mode used for work trips:

```{r scc-38}
## Generate frequency for trip Mode
temp1<-as.data.frame(table(actw_87O$Mode))
## Create percentage of frequency
temp1$Percent<-temp1$Freq*100/sum(temp1$Freq)
## Names 
names(temp1)
## Rename the variables
colnames(temp1)[c(1:3)] <- c("Mode","Ob_UW","Per_UW")
## Show the results
temp1
```

Given those, we can expand them with the weights for Santa Cruz County:

```{r scc-39}
## Create weighted frequency
temp2<-aggregate(NEXPTCFPERWGT~Mode,FUN=sum,data=actw_87O)
## Create weighted percentage
temp2$Percent<-temp2$NEXPTCFPERWGT*100/sum(temp2$NEXPTCFPERWGT)
## Names
names(temp2)
## Rename the variables
colnames(temp2)[c(1:3)] <- c("Mode","Ob_W","Per_W")
## Show
temp2
```

And again, we finish by saving the results in a csv file:

```{r scc-40}
## Merge the two files and save
temp3<-merge(temp1,temp2, by ="Mode", all = TRUE,sort=FALSE)
temp3 <- temp3[,c(1,2,4,3,5)]
temp3
temp3[,1]<-as.character(temp3[,1])
temp3<- rbind(temp3, c("Total", colSums(temp3[,2:5],na.rm = TRUE)))
temp3
## Save as csv file
write.csv(temp3,file=pto("Mode_wact87O.csv"),row.names=FALSE)
scc.mode.wact.orig <- temp3
rm(temp1)
rm(temp2)
rm(temp3)
```

### Work Trips Ending in  Santa Cruz County


Now let us find trips ending in Santa Cruz county `(CTFIP=87)`.
Again, we subset the place table and use Santa
Cruz County weights if trip origins are in Santa Cruz, and statewide
weights if origins are outside of Santa Cruz County.


```{r scc-41}
# Create a subset for trips from  Santa Cruz County 
act_87D<-subset(act_place,CTFIP==87)
names(act_87D)
head(act_87D)
# Change the weight for non  Santa Cruz county origin
# (Assign state wide weight for non  Santa Cruz origin)
act_87D$NTCFW <-
  ifelse((act_87D$OCTFIP==87),
         act_87D$NEXPTCFPERWGT,
         act_87D$EXPTCFPERWGT.y)
names(act_87D)
head(act_87D)
```

Below, we generate a frequency table by purpose:

```{r scc-42}
# Generate observations by trip purpose from  Santa Cruz based on new weights
## Generate frequency for trip Mode
temp1<-as.data.frame(table(act_87D$Actpur))
## Create percentage of frequency
temp1$Percent<-temp1$Freq*100/sum(temp1$Freq)
## Names 
names(temp1)
## Rename the variables
colnames(temp1)[c(1:3)] <- c("Actpur","Ob_UW","Per_UW")
## Show the results
temp1
```

Now we create a weighted frequency table:

```{r scc-43}
## Create weighted frequency
temp2<-aggregate(NTCFW~Actpur,FUN=sum,data=act_87D)
## Create weighted percentage
temp2$Percent<-temp2$NTCFW*100/sum(temp2$NTCFW)
## Names
names(temp2)
## Rename the variables
colnames(temp2)[c(1:3)] <- c("Actpur","Ob_W","Per_W")
## Show
temp2
```

And now we save the results in a csv file:

```{r scc-44}
## Merge the two files and save
temp3<-merge(temp1,temp2, by ="Actpur", all = TRUE,sort=FALSE)
temp3 <- temp3[,c(1,2,4,3,5)]
temp3
temp3[,1]<-as.character(temp3[,1])
temp3<- rbind(temp3, c("Total", colSums(temp3[,2:5],na.rm = TRUE)))
temp3
## Save as csv file
write.csv(temp3,file=pto("Actpur_87D.csv"),row.names=FALSE)
apurp.scc.dest <- temp3
rm(temp1)
rm(temp2)
rm(temp3)
```

Finally, we compute the frequencies of various modes used for work
trips ending in Santa Cruz County:

```{r scc-45}
# We want only work trips (APURP=9)
# Create a subset for work trips
actw_87D<-subset(act_87D,APURP==9)
## Generate frequency for trip Mode
temp1<-as.data.frame(table(actw_87D$Mode))
## Create percentage of frequency
temp1$Percent<-temp1$Freq*100/sum(temp1$Freq)
## Names 
names(temp1)
## Rename the variables
colnames(temp1)[c(1:3)] <- c("Mode","Ob_UW","Per_UW")
## Show the results
temp1
```

Here is the weighted frequency table:

```{r scc-46}
## Create weighted frequency
temp2<-aggregate(NTCFW~Mode,FUN=sum,data=actw_87D)
## Create weighted percentage
temp2$Percent<-temp2$NTCFW*100/sum(temp2$NTCFW)
## Names
names(temp2)
## Rename the variables
colnames(temp2)[c(1:3)] <- c("Mode","Ob_W","Per_W")
## Show
temp2
## Merge the two files and save
temp3<-merge(temp1,temp2, by ="Mode", all = TRUE,sort=FALSE)
temp3 <- temp3[,c(1,2,4,3,5)]
temp3
temp3[,1]<-as.character(temp3[,1])
temp3<- rbind(temp3, c("Total", colSums(temp3[,2:5],na.rm = TRUE)))
temp3
```

...and here we save it:

```{r scc-47}
## Save as csv file
write.csv(temp3,file=pto("Mode_wact87D.csv"),row.names=FALSE)
scc.mode.wact.dest <- temp3
rm(temp1)
rm(temp2)
rm(temp3)
```

We are done generating tables by mode for Santa Cruz County.

\fi

\newpage

# Results

```{r scc-mode-code,echo=FALSE,results='asis'}
mmap <- nplacedata %>%
    group_by(MODE,Mode) %>% summarise(n=n())
pandoc.table.double(
    mmap %>%
    select(Code=MODE,Mode),
    caption="Mode Code",
    justify=c('left','left'),
    split.cells=Inf,
    missing="&nbsp;")
```

\newpage

```{r scc-act-purp-code,echo=FALSE,results='asis'}
amap <- act_place %>%
    group_by(APURP,Actpur) %>% summarise(n=n())
pandoc.table.double(
    amap %>% select(Code=APURP,Purpose=Actpur),
    caption="Activity Purpose Code",
    justify=c('left','left'),
    split.cells=c('10%','90%'),
    split.table=100,
  missing='&nbsp;')
```

\newpage

```{r mode.scc.orig,echo=FALSE,results='asis'}
mode.scc.orig <- mode.scc.orig %>%
  arrange(ifelse(Mode=="Total",1,0),desc(as.numeric(Ob_W))) %>%
  transmute(Mode=Mode,
            'Observed Unweighted'=format(round(as.numeric(Ob_UW),0),big.mark=','),
            'Observed Weighted'=format(round(as.numeric(Ob_W),0),big.mark=','),
            'Percent Unweighted'=format(round(as.numeric(Per_UW),2),big.mark=','),
            'Percent Weighted'=format(round(as.numeric(Per_W),2),big.mark=','))
#scc.mode.orig2[,4:5] <- format(round(as.numeric(scc.mode.orig2[,4:5]),2))
pandoc.table(
    mode.scc.orig,
    caption="Total number of trips originating in Santa Cruz County",
    justify=c('left','right','right','right','right'),
    split.tables=100,
    split.cells=c('40%','15%','15%','15%','15%'),
    big.mark=",",
    emphasize.strong.rows=which(mode.scc.orig$Mode=='Total',arr.ind=TRUE)
)
```

```{r mode.da.scc.orig,echo=FALSE,results='asis'}
mode.da.scc.orig <- mode.da.scc.orig %>%
  arrange(ifelse(Mode=="Total",1,0),desc(as.numeric(Observations_W))) %>%
  transmute(Mode=Mode,
            'Observed Unweighted'=format(round(as.numeric(Observations_UW),0),big.mark=','),
            'Observed Weighted'=format(round(as.numeric(Observations_W),0),big.mark=','),
            'Percent Unweighted'=format(round(as.numeric(Percent_UW),2),big.mark=','),
            'Percent Weighted'=format(round(as.numeric(Percent_W),2),big.mark=','))
pandoc.table(
    mode.da.scc.orig,
    caption="Modal split (drive alone) of trips originating in Santa Cruz County",
    justify=c('left','right','right','right','right'),
    split.tables=100,
    split.cells=c('40%','15%','15%','15%','15%'),
    big.mark=",",
    emphasize.strong.rows=which(mode.da.scc.orig$Mode=='Total',arr.ind=TRUE)
)
```

\newpage

```{r dist.scc.orig,echo=FALSE,results='asis'}
dist.scc.orig <- dist.scc.orig %>%
  arrange(ifelse(Mode=="Total",1,0),desc(as.numeric(TDis_W))) %>%
  transmute('Mode'=Mode,
            'Trip&nbsp;Dist UW'=format(round(as.numeric(TDis_UW),0),big.mark=','),
            'Trip&nbsp;Dist W'=format(round(as.numeric(TDis_W),0),big.mark=','),
            'Percent UW'=format(round(as.numeric(Per_UW),2),big.mark=','),
            'Percent W'=format(round(as.numeric(Per_W),2),big.mark=','),
            'Avg Dist UW'=format(round(as.numeric(avg_uw),2),big.mark=','),
            'Avg Dist W'=format(round(as.numeric(avg_w),2),big.mark=',')
            )
pandoc.table(
    dist.scc.orig,
    caption="Trip distance (mi) by mode for trips originating in Santa Cruz County",
    justify=c('left','right','right','right','right','right','right'),
    split.tables=120,
    split.cells=c('20%','15%','25%','10%','10%','10%','10%'),
    big.mark=",",
    emphasize.strong.rows=which(dist.scc.orig$Mode=='Total',arr.ind=TRUE)
)
```

\newpage

```{r dur.scc.orig,echo=FALSE,results='asis'}
dur.scc.orig <- dur.scc.orig %>%
  arrange(ifelse(Mode=="Total",1,0),desc(as.numeric(TDu_W))) %>%
  transmute(Mode=Mode,
            'Trip&nbsp;Dur UW'=format(round(as.numeric(TDu_UW),0),big.mark=','),
            'Trip&nbsp;Dur W'=format(round(as.numeric(TDu_W),0),big.mark=','),
            'Percent UW'=format(round(as.numeric(Per_UW),2),big.mark=','),
            'Percent W'=format(round(as.numeric(Per_W),2),big.mark=','),
            'Avg Dur UW'=format(round(as.numeric(avg_uw),2),big.mark=','),
            'Avg Dur W'=format(round(as.numeric(avg_w),2),big.mark=',')
            )
pandoc.table(
    dur.scc.orig,
    caption="Trip duration (min) by mode for trips originating in Santa Cruz County",
    justify=c('left','right','right','right','right','right','right'),
    split.tables=120,
    split.cells=c('20%','15%','25%','10%','10%','10%','10%'),
    big.mark=",",
    emphasize.strong.rows=which(dur.scc.orig$Mode=='Total',arr.ind=TRUE)
)
```

\newpage

```{r apurp.scc.orig,echo=FALSE,results='asis'}
apurp.scc.orig <- apurp.scc.orig %>%
  arrange(ifelse(Actpur=="Total",1,0),desc(as.numeric(Ob_W))) %>%
  transmute('Activity Purpose'=Actpur,
            'Obs UW'=format(round(as.numeric(Ob_UW),0),big.mark=','),
            'Obs W'=format(round(as.numeric(Ob_W),0),big.mark=','),
            'Pct UW'=format(round(as.numeric(Per_UW),2),big.mark=','),
            'Pct W'=format(round(as.numeric(Per_W),2),big.mark=',')
            )
pandoc.table(
  apurp.scc.orig,
  caption="Total number of work trips (APURP=9) for trips originating in Santa Cruz County",
  justify=c('left','right','right','right','right'),
  split.tables=120,
  split.cells=c('60%','10%','10%','10%','10%'),
  big.mark=",",
  emphasize.strong.rows=which(apurp.scc.orig$'Activity Purpose'=='Total',arr.ind=TRUE)
)
```

```{r scc.mode.wact.orig,echo=FALSE,results='asis'}
scc.mode.wact.orig <- scc.mode.wact.orig %>%
  arrange(ifelse(Mode=="Total",1,0),desc(as.numeric(Ob_W))) %>%
  transmute(Mode=Mode,
            'Obs UW'=format(round(as.numeric(Ob_UW),0),big.mark=','),
            'Obs W'=format(round(as.numeric(Ob_W),0),big.mark=','),
            'Pct UW'=format(round(as.numeric(Per_UW),2),big.mark=','),
            'Pct W'=format(round(as.numeric(Per_W),2),big.mark=',')
            )
pandoc.table(
  scc.mode.wact.orig,
  caption="Total number of trips by activity purpose for trips originating in Santa Cruz County",
  justify=c('left','right','right','right','right'),
  split.tables=120,
  split.cells=c('60%','10%','10%','10%','10%'),
  big.mark=",",
  emphasize.strong.rows=which(scc.mode.wact.orig$Mode=='Total',arr.ind=TRUE)
)
```


<!-- Destinations --->

\newpage

```{r mode.scc.dest,echo=FALSE,results='asis'}
mode.scc.dest <- mode.scc.dest %>%
  arrange(ifelse(Mode=="Total",1,0),desc(as.numeric(Ob_W))) %>%
  transmute(Mode=Mode,
            'Observed Unweighted'=format(round(as.numeric(Ob_UW),0),big.mark=','),
            'Observed Weighted'=format(round(as.numeric(Ob_W),0),big.mark=','),
            'Percent Unweighted'=format(round(as.numeric(Per_UW),2),big.mark=','),
            'Percent Weighted'=format(round(as.numeric(Per_W),2),big.mark=','))
#scc.mode.dest2[,4:5] <- format(round(as.numeric(scc.mode.dest2[,4:5]),2))
pandoc.table(
    mode.scc.dest,
    caption="Total number of trips ending in Santa Cruz County",
    justify=c('left','right','right','right','right'),
    split.tables=100,
    split.cells=c('40%','15%','15%','15%','15%'),
    big.mark=",",
    emphasize.strong.rows=which(mode.scc.dest$Mode=='Total',arr.ind=TRUE)
)
```

```{r mode.da.scc.dest,echo=FALSE,results='asis'}
mode.da.scc.dest <- mode.da.scc.dest %>%
  arrange(ifelse(Mode=="Total",1,0),desc(as.numeric(Ob_W))) %>%
  transmute(Mode=Mode,
            'Observed Unweighted'=format(round(as.numeric(Ob_UW),0),big.mark=','),
            'Observed Weighted'=format(round(as.numeric(Ob_W),0),big.mark=','),
            'Percent Unweighted'=format(round(as.numeric(Per_UW),2),big.mark=','),
            'Percent Weighted'=format(round(as.numeric(Per_W),2),big.mark=','))
pandoc.table(
    mode.da.scc.dest,
    caption="Modal split (drive alone) of trips ending in Santa Cruz County",
    justify=c('left','right','right','right','right'),
    split.tables=100,
    split.cells=c('40%','15%','15%','15%','15%'),
    big.mark=",",
    emphasize.strong.rows=which(mode.da.scc.dest$Mode=='Total',arr.ind=TRUE)
)
```

\newpage

```{r dist.scc.dest,echo=FALSE,results='asis'}
dist.scc.dest <- dist.scc.dest %>%
  arrange(ifelse(Mode=="Total",1,0),desc(as.numeric(TDis_W))) %>%
  transmute('Mode'=Mode,
            'Trip&nbsp;Dist UW'=format(round(as.numeric(TDis_UW),0),big.mark=','),
            'Trip&nbsp;Dist W'=format(round(as.numeric(TDis_W),0),big.mark=','),
            'Percent UW'=format(round(as.numeric(Per_UW),2),big.mark=','),
            'Percent W'=format(round(as.numeric(Per_W),2),big.mark=','),
            'Avg Dist UW'=format(round(as.numeric(avg_uw),2),big.mark=','),
            'Avg Dist W'=format(round(as.numeric(avg_w),2),big.mark=',')
            )
pandoc.table(
    dist.scc.dest,
    caption="Trip distance (mi) by mode for trips ending in Santa Cruz County",
    justify=c('left','right','right','right','right','right','right'),
    split.tables=120,
    split.cells=c('20%','15%','25%','10%','10%','10%','10%'),
    big.mark=",",
    emphasize.strong.rows=which(dist.scc.dest$Mode=='Total',arr.ind=TRUE)
)
```

\newpage

```{r dur.scc.dest,echo=FALSE,results='asis'}
dur.scc.dest <- dur.scc.dest %>%
  arrange(ifelse(Mode=="Total",1,0),desc(as.numeric(TDu_W))) %>%
  transmute(Mode=Mode,
            'Trip&nbsp;Dur UW'=format(round(as.numeric(TDu_UW),0),big.mark=','),
            'Trip&nbsp;Dur W'=format(round(as.numeric(TDu_W),0),big.mark=','),
            'Percent UW'=format(round(as.numeric(Per_UW),2),big.mark=','),
            'Percent W'=format(round(as.numeric(Per_W),2),big.mark=','),
            'Avg Dur UW'=format(round(as.numeric(avg_uw),2),big.mark=','),
            'Avg Dur W'=format(round(as.numeric(avg_w),2),big.mark=',')
            )
pandoc.table(
    dur.scc.dest,
    caption="Trip duration (min) by mode for trips ending in Santa Cruz County",
    justify=c('left','right','right','right','right','right','right'),
    split.tables=120,
    split.cells=c('20%','15%','25%','10%','10%','10%','10%'),
    big.mark=",",
    emphasize.strong.rows=which(dur.scc.dest$Mode=='Total',arr.ind=TRUE)
)
```

\newpage

```{r apurp.scc.dest,echo=FALSE,results='asis'}
apurp.scc.dest <- apurp.scc.dest %>%
  arrange(ifelse(Actpur=="Total",1,0),desc(as.numeric(Ob_W))) %>%
  transmute('Activity Purpose'=Actpur,
            'Obs UW'=format(round(as.numeric(Ob_UW),0),big.mark=','),
            'Obs W'=format(round(as.numeric(Ob_W),0),big.mark=','),
            'Pct UW'=format(round(as.numeric(Per_UW),2),big.mark=','),
            'Pct W'=format(round(as.numeric(Per_W),2),big.mark=',')
            )
pandoc.table(
  apurp.scc.dest,
  caption="Total number of work trips (APURP=9) for trips ending in Santa Cruz County",
  justify=c('left','right','right','right','right'),
  split.tables=120,
  split.cells=c('60%','10%','10%','10%','10%'),
  big.mark=",",
  emphasize.strong.rows=which(apurp.scc.dest$'Activity Purpose'=='Total',arr.ind=TRUE)
)
```

```{r scc.mode.wact.dest,echo=FALSE,results='asis'}
scc.mode.wact.dest <- scc.mode.wact.dest %>%
  arrange(ifelse(Mode=="Total",1,0),desc(as.numeric(Ob_W))) %>%
  transmute(Mode=Mode,
            'Obs UW'=format(round(as.numeric(Ob_UW),0),big.mark=','),
            'Obs W'=format(round(as.numeric(Ob_W),0),big.mark=','),
            'Pct UW'=format(round(as.numeric(Per_UW),2),big.mark=','),
            'Pct W'=format(round(as.numeric(Per_W),2),big.mark=',')
            )
pandoc.table(
  scc.mode.wact.dest,
  caption="Total number of trips by activity purpose for trips ending in Santa Cruz County",
  justify=c('left','right','right','right','right'),
  split.tables=120,
  split.cells=c('60%','10%','10%','10%','10%'),
  big.mark=",",
  emphasize.strong.rows=which(scc.mode.wact.dest$Mode=='Total',arr.ind=TRUE)
)
```
