---
title: "6.4 Data Request for Lake County"
author: "Suman Mitra"
date: "September 3, 2015"
output: html_document
---

```{r lc-setup-0,echo=FALSE,cache=FALSE}
if ( !exists("rd") ) { rd <- gsub("(.*?chts-book).*","\\1",getwd()) }
source(paste(rd,'R/setup-hooks.R',sep="/"))
```

1. Trips by Mode
2. Modal split for driving alone
3. Trip distance by Mode
4. Trip duration by Mode
5. Trips by Purpose
6. Work trips by mode

```{r results='asis',echo=FALSE,cache=FALSE}
############################################################ NESTABLE SECTION
SEC('#','Summary and Methodology')
```


We generate tables by mode for the following: total number of trips, total trip distance, average trip distance, total trip duration, average trip duration, and total number of trips for driving alone.  To calculate the drive alone numbers we extract data for total travelers (`TOTTR=1`) and `MODE={5, 6, 7, 8, 10}` (basically, drivers of cars, trucks, or motorcycles). We also generate tables for the total number of trips by trip purpose and for work trips by mode. Here, we actually generate tables for total number of trips by activity purpose since the trip purpose is given in the database as activity purpose and there is more than one activity in many single trips. We consider each activity as a trip. We only consider the purpose code (`APURP=9`) for work trips.  We generate every table for both trips originating in and ending in Lake County. We use the new county wide weights of the Lake County to generate weighted results for trips that originated in Lake County and statewide weights to generate weighted results for trips that originated outside of Lake County. 


```{r results='asis',echo=FALSE,cache=FALSE}
############################################################ NESTABLE SECTION
SEC('#','R Code and Implementation')
```


The following commands are used to generate the results.

```{r results='asis',echo=FALSE,cache=FALSE}
############################################################ NESTABLE SECTION
SEC('##','Trips to Lake County')
```


```{r results='asis',echo=FALSE,cache=FALSE}
############################################################ NESTABLE SECTION
SEC('###','Mode split calculations')
```


To answer the questions we need the place level table (with weights) that we computed in the section 4.

```{r lc-0,echo=FALSE}
if ( !exists("rd") ) { rd <- "/home/crindt/chts-book" }
p <- function(ps) { return(paste(rd,ps,sep="/")) }
```

```{r lc-1, cache=TRUE}
# Set the working directory
setwd(p("District 1/Lake County"))


#Import the place level with new weight csv file and give a new name
nplace<- read.csv(p("Weight/NewWeightData/nplacedata3032015.csv"), header=TRUE)
names(nplace)
head(nplace)# Show first ten observations

```

Next, we do some manipulations to determine origin county and convert the `MODE` codes into 
text descriptions.

```{r lc-2,cache=TRUE}
# Create trip origin county 
# Need dplyr
library("dplyr")
nplacedata <- nplace %>% group_by(SAMPN,PERNO)%>% arrange(SAMPN,PERNO,PLANO) %>% mutate(OCTFIP=lag(CTFIP))

# For recoding Mode
nplacedata$Mode<-ifelse(nplacedata$MODE==1,"Walk",ifelse(nplacedata$MODE==2,"Bike",ifelse(nplacedata$MODE==3,"Wheelchair/Mobility Scooter",ifelse(nplacedata$MODE==4,"Other Non-Motorized",ifelse(nplacedata$MODE==5,"Auto/Van/Truck Driver",ifelse(nplacedata$MODE==6,"Auto /Van/Truck Passenger",ifelse(nplacedata$MODE==7,"Carpool/Vanpool",ifelse(nplacedata$MODE==8,"Motorcycle/Scooter/Moped",ifelse(nplacedata$MODE==9,"Taxi/Hired Car/Limo",ifelse(nplacedata$MODE==10,"Rental Car/Vehicle",ifelse(nplacedata$MODE==11,"Private shuttle ",ifelse(nplacedata$MODE==12,"Greyhound Bus",ifelse(nplacedata$MODE==13,"Plane",ifelse(nplacedata$MODE==14,"Other Private Transit",ifelse(nplacedata$MODE==15,"Local Bus,Rapid Bus",ifelse(nplacedata$MODE==16,"Express Bus/Commuter Bus",ifelse(nplacedata$MODE==17,"Premium Bus",ifelse(nplacedata$MODE==18,"School Bus",ifelse(nplacedata$MODE==19,"Public Transit Shuttle",ifelse(nplacedata$MODE==20,"Air BART/LAX Fly Away",ifelse(nplacedata$MODE==21,"Dial-a-Ride/Paratransit", ifelse(nplacedata$MODE==22,"Amtrak Bus", ifelse(nplacedata$MODE==23,"Other Bus ", ifelse(nplacedata$MODE==24,"BART,Metro Red/Purple Line",ifelse(nplacedata$MODE==25,"ACE, Amtrak, Cal train",ifelse(nplacedata$MODE==26,"Metro",ifelse(nplacedata$MODE==27,"Street Car/Cable Car",ifelse(nplacedata$MODE==28,"Other Rail",ifelse(nplacedata$MODE==29,"Ferry / Boat","Don't Know/Refused")))))))))))))))))))))))))))))
```


Now, we first focus on trips ending in Lake County.  Here we use different 
weights depending on the origin of the trip.  Specifically, when the origin is Lake 
County, we use the county-specific weights.  When the origin is outside 
Lake County, we use the statewide weights.

```{r lc-3,cache=TRUE}
# Trips to Lake County
# Create a subset for trips to Lake county 
# We are only interested in trips to Lake County (CTFIP=33) 
# Create a subset for trips to Lake County 
place_33<-subset(nplacedata,CTFIP==33)
names(place_33)
# Change the weight for non Lake County origin (Assign state wide weight for non Lake origin)
place_33$NTCFW <- ifelse((place_33$OCTFIP==33), place_33$NEXPTCFPERWGT, place_33$EXPTCFPERWGT)
head(place_33)
```

So `place_33$NTCFW` now contains the trips ending in Lake county with appropriate 
expansion weights appended.  Now we can generate observations for modal split for trips 
to Lake based on these weights.

Notes: In the tables: Ob = Observations, Per = Percent, UW = Unweighted and W = Weighted.

```{r lc-4,cache=TRUE}
## Generate frequency for trip Mode
temp1<-as.data.frame(table(place_33$Mode))
## Create percentage of frequency
temp1$Percent<-temp1$Freq*100/sum(temp1$Freq)
## Names 
names(temp1)
## Rename the variables
colnames(temp1)[c(1:3)] <- c("Mode","Ob_UW","Per_UW")
## Show the results
temp1
```

Now we compute the weighted frequency.

```{r lc-5,cache=TRUE}
## Create weighted frequency
temp2<-aggregate(NTCFW~Mode,FUN=sum,data=place_33)
## Create weighted percentage
temp2$Percent<-temp2$NTCFW*100/sum(temp2$NTCFW)
## Names
names(temp2)
## Rename the variables
colnames(temp2)[c(1:3)] <- c("Mode","Ob_W","Per_W")
## Show
temp2
```

Now we can save these results in a `csv` file for later use.

```{r lc-6,cache=TRUE}
## Merge the two files and save
temp3<-merge(temp1,temp2, by ="Mode", all = TRUE,sort=TRUE)
temp3 <- temp3[,c(1,2,4,3,5)]
temp3
Mode_33d<-temp3
temp3[,1]<-as.character(temp3[,1])
temp3<- rbind(temp3, c("Total", colSums(temp3[,2:5],na.rm = TRUE)))
temp3
## Save as csv file
write.csv(temp3,file="Mode_33d.csv",row.names=FALSE)
rm(temp1)
rm(temp2)
rm(temp3)
```

```{r results='asis',echo=FALSE,cache=FALSE}
############################################################ NESTABLE SECTION
SEC('###','Trip duration by mode')
```


Next, we perform a similar computation for Total Trip duration for all trips ending 
in Lake, by mode.

```{r lc-7,cache=TRUE}
## Generate frequency for trip Mode
temp1<-aggregate(TRIPDUR~Mode,FUN=sum,data=place_33)
## Create  percentage
temp1$Percent<-temp1$TRIPDUR*100/sum(temp1$TRIPDUR)
## Names
names(temp1)
## Rename the variables
colnames(temp1)[c(1:3)] <- c("Mode","TDu_UW","Per_UW")
## Show
temp1
```

Again, we compute the weighted frequency.

```{r lc-8,cache=TRUE}
## Create weighted frequency
temp<-place_33$NTCFW*place_33$TRIPDUR
temp2<-aggregate(temp~Mode,FUN=sum,data=place_33)
## Create weighted percentage
temp2$Percent<-temp2$temp*100/sum(temp2$temp)
## Names
names(temp2)
## Rename the variables
colnames(temp2)[c(1:3)] <- c("Mode","TDu_W","Per_W")
## Show
temp2
```


And as before, we save the results for later use. 

```{r lc-9,cache=TRUE}
## Merge the two files and save
temp3<-merge(temp1,temp2, by ="Mode", all = TRUE)
temp3 <- temp3[,c(1,2,4,3,5)]
# Average trip duration
temp3$avg_uw<-temp3$TDu_UW/Mode_33d$Ob_UW
temp3
temp3$avg_w<-temp3$TDu_W/Mode_33d$Ob_W
temp3
temp3[,1]<-as.character(temp3[,1])
temp3<- rbind(temp3, c("Total", colSums(temp3[,2:7],na.rm = TRUE)))
temp3
## Save as csv file
write.csv(temp3,file="TripDUrm_D33.csv",row.names=FALSE)
rm(temp1)
rm(temp2)
rm(temp3)
rm(temp)
```

```{r results='asis',echo=FALSE,cache=FALSE}
############################################################ NESTABLE SECTION
SEC('###','Trip distance by mode')
```


Moving on, we determine Total Trip distance for all trips ending in Lake, by mode.

```{r lc-10,cache=TRUE}
## Generate frequency for trip Mode
temp1<-aggregate(TripDistance~Mode,FUN=sum,data=place_33)
## Create  percentage
temp1$Percent<-temp1$TripDistance*100/sum(temp1$TripDistance)
## Names
names(temp1)
## Rename the variables
colnames(temp1)[c(1:3)] <- c("Mode","TDis_UW","Per_UW")
## Show
temp1
```

Next, we compute the weighted trip distance by mode.

```{r lc-11,cache=TRUE}
## Create weighted frequency
temp<-place_33$NTCFW*place_33$TripDistance
temp2<-aggregate(temp~Mode,FUN=sum,data=place_33)
## Create weighted percentage
temp2$Percent<-temp2$temp*100/sum(temp2$temp)
## Names
names(temp2)
## Rename the variables
colnames(temp2)[c(1:3)] <- c("Mode","TDis_W","Per_W")
## Show
temp2
```

And, we save the results for later use.

```{r lc-12,cache=TRUE}
## Merge the two files and save
temp3<-merge(temp1,temp2, by ="Mode", all = TRUE)
temp3 <- temp3[,c(1,2,4,3,5)]
# Average trip distance
temp3$avg_uw<-temp3$TDis_UW/Mode_33d$Ob_UW
temp3
temp3$avg_w<-temp3$TDis_W/Mode_33d$Ob_W
temp3
temp3[,1]<-as.character(temp3[,1])
temp3<- rbind(temp3, c("Total", colSums(temp3[,2:7],na.rm = TRUE)))
temp3
## Save as csv file
write.csv(temp3,file="TripDistanceM_D33.csv",row.names=FALSE)
rm(temp1)
rm(temp2)
rm(temp3)
rm(temp)
```

```{r results='asis',echo=FALSE,cache=FALSE}
############################################################ NESTABLE SECTION
SEC('###','Modal split for driving alone')
```


The next question was the modal split for Drive Alone.  To compute this, we
subset the data based on `TOTTR` (Total travelers) and `MODE={5, 6, 7, 8, 10}` 
(basically, the driver of a car, truck, or motorcycle)

```{r lc-13}
place_33d<-subset(place_33,(TOTTR==1 & (MODE==5|MODE==6|MODE==7|MODE==8|MODE==10)))
head(place_33d)
```

Now we compute the frequency of modes for this subset.

```{r lc-14}
## Create frequency for different Mode category for drive alone
temp1<-as.data.frame(table(place_33d$Mode))
## Create percentage of frequency
temp1$Percent<-temp1$Freq*100/sum(temp1$Freq)
## Names 
names(temp1)
## Rename the variables
colnames(temp1)[c(1:3)] <- c("Mode","Ob_UW","Per_UW")
## Show the results
temp1
```

Following the pattern, we compute the weighted frequency.

```{r lc-15}
## Create weighted frequency
temp2<-aggregate(NTCFW~Mode,FUN=sum,data=place_33d)
## Create weighted percentage
temp2$Percent<-temp2$NTCFW*100/sum(temp2$NTCFW)
## Names
names(temp2)
## Rename the variables
colnames(temp2)[c(1:3)] <- c("Mode","Ob_W","Per_W")
## Show
temp2
```

Then, once again, we save the results for later use.

```{r lc-16}

## Merge the two files and save
temp3<-merge(temp1,temp2, by ="Mode", all = TRUE,sort=FALSE)
temp3 <- temp3[,c(1,2,4,3,5)]
temp3
temp3[,1]<-as.character(temp3[,1])
temp3<- rbind(temp3, c("Total", colSums(temp3[,2:5],na.rm = TRUE)))
temp3
## Save as csv file
write.csv(temp3,file="Mode_33Ddrialn.csv",row.names=FALSE)
rm(temp1)
rm(temp2)
rm(temp3)
```


```{r results='asis',echo=FALSE,cache=FALSE}
############################################################ NESTABLE SECTION
SEC('##','Trips from Lake County')
```


Now we perform a similar computation for trips *from* Lake County.  As before, we
subset the place data, but this time for trips originating in Lake County.

```{r lc-17,cache=TRUE}

# Let us estimate trips from Lake county (OCTFIP=33) 
# Create a subset for trips from Lake county 
place_33O<-subset(nplacedata,OCTFIP==33)
names(place_33O)
head(place_33O)
```

```{r results='asis',echo=FALSE,cache=FALSE}
############################################################ NESTABLE SECTION
SEC('###','Trips by mode')
```


Now we generate the mode frequency table for all trips from Lake County.

```{r lc-18,cache=TRUE}
# Generate observations for modal split for trips from Lake based on new weights
## Generate frequency for trip Mode
temp1<-as.data.frame(table(place_33O$Mode))
## Create percentage of frequency
temp1$Percent<-temp1$Freq*100/sum(temp1$Freq)
## Names 
names(temp1)
## Rename the variables
colnames(temp1)[c(1:3)] <- c("Mode","Ob_UW","Per_UW")
## Show the results
temp1
```

This should be looking familiar now.  Here is the weighted frequency.

```{r lc-19,cache=TRUE}
## Create weighted frequency
temp2<-aggregate(NEXPTCFPERWGT~Mode,FUN=sum,data=place_33O)
## Create weighted percentage
temp2$Percent<-temp2$NEXPTCFPERWGT*100/sum(temp2$NEXPTCFPERWGT)
## Names
names(temp2)
## Rename the variables
colnames(temp2)[c(1:3)] <- c("Mode","Ob_W","Per_W")
## Show
temp2
```

And again, we save for later use.

```{r lc-20,cache=TRUE}
## Merge the two files and save
temp3<-merge(temp1,temp2, by ="Mode", all = TRUE,sort=FALSE)
temp3 <- temp3[,c(1,2,4,3,5)]
temp3
Mode_33O<-temp3
temp3[,1]<-as.character(temp3[,1])
temp3<- rbind(temp3, c("Total", colSums(temp3[,2:5],na.rm = TRUE)))
temp3
## Save as csv file
write.csv(temp3,file="Mode_33O.csv",row.names=FALSE)
rm(temp1)
rm(temp2)
rm(temp3)
```

```{r results='asis',echo=FALSE,cache=FALSE}
############################################################ NESTABLE SECTION
SEC('###','Trip duration by Mode')
```


Now we repeat the trip duration calculation for Lake Origins.

```{r lc-21,cache=TRUE}
## Total Trip duration for all trips from Lake, by Mode
## Generate frequency for trip Mode
temp1<-aggregate(TRIPDUR~Mode,FUN=sum,data=place_33O)
## Create  percentage
temp1$Percent<-temp1$TRIPDUR*100/sum(temp1$TRIPDUR)
## Names
names(temp1)
## Rename the variables
colnames(temp1)[c(1:3)] <- c("Mode","TDu_UW","Per_UW")
## Show
temp1
```

Here is the weighted frequency calculation.

```{r lc-22,cache=TRUE}
## Create weighted frequency
temp<-place_33O$NEXPTCFPERWGT*place_33O$TRIPDUR
temp2<-aggregate(temp~Mode,FUN=sum,data=place_33O)
## Create weighted percentage
temp2$Percent<-temp2$temp*100/sum(temp2$temp)
## Names
names(temp2)
## Rename the variables
colnames(temp2)[c(1:3)] <- c("Mode","TDu_W","Per_W")
## Show
temp2
```

And we save again.

```{r lc-23,cache=TRUE}
## Merge the two files and save
temp3<-merge(temp1,temp2, by ="Mode", all = TRUE)
temp3 <- temp3[,c(1,2,4,3,5)]
temp3
temp3$avg_uw<-temp3$TDu_UW/Mode_33O$Ob_UW
temp3
temp3$avg_w<-temp3$TDu_W/Mode_33O$Ob_W
temp3
temp3[,1]<-as.character(temp3[,1])
temp3<- rbind(temp3, c("Total", colSums(temp3[,2:7],na.rm = TRUE)))
temp3
## Save as csv file
write.csv(temp3,file="TripDUrm_O.csv",row.names=FALSE)
rm(temp1)
rm(temp2)
rm(temp3)
rm(temp)
```

```{r results='asis',echo=FALSE,cache=FALSE}
############################################################ NESTABLE SECTION
SEC('###','Trip distance by Mode')
```


Here is the trip distance calculation for all trips from Lake, by Mode.

```{r lc-24,cache=TRUE}
## Generate frequency for trip Mode
temp1<-aggregate(TripDistance~Mode,FUN=sum,data=place_33O)
## Create  percentage
temp1$Percent<-temp1$TripDistance*100/sum(temp1$TripDistance)
## Names
names(temp1)
## Rename the variables
colnames(temp1)[c(1:3)] <- c("Mode","TDis_UW","Per_UW")
## Show
temp1
```

Followed again by the weighted frequency calculation.

```{r lc-25,cache=TRUE}
## Create weighted frequency
temp<-place_33O$NEXPTCFPERWGT*place_33O$TripDistance
temp2<-aggregate(temp~Mode,FUN=sum,data=place_33O)
## Create weighted percentage
temp2$Percent<-temp2$temp*100/sum(temp2$temp)
## Names
names(temp2)
## Rename the variables
colnames(temp2)[c(1:3)] <- c("Mode","TDis_W","Per_W")
## Show
temp2
```

And our now familiar save of the data.

```{r lc-26,cache=TRUE}
## Merge the two files and save
temp3<-merge(temp1,temp2, by ="Mode", all = TRUE)
temp3 <- temp3[,c(1,2,4,3,5)]
temp3
temp3$avg_uw<-temp3$TDis_UW/Mode_33O$Ob_UW
temp3
temp3$avg_w<-temp3$TDis_W/Mode_33O$Ob_W
temp3
temp3[,1]<-as.character(temp3[,1])
temp3<- rbind(temp3, c("Total", colSums(temp3[,2:7],na.rm = TRUE)))
temp3
## Save as csv file
write.csv(temp3,file="TripDistanceM_O.csv",row.names=FALSE)
rm(temp1)
rm(temp2)
rm(temp3)
rm(temp)
```

```{r results='asis',echo=FALSE,cache=FALSE}
############################################################ NESTABLE SECTION
SEC('###','Modal split for driving alone')
```


And lastly for the Lake origins, here is the drive alone modal split.  As before, 
we subset the data based on `TOTTR` (Total travelers) and `MODE={5, 6, 7, 8, 10}`
(the driver of a car, truck, or motorcycle) 

```{r lc-27,cache=TRUE} 
place_33dO<-subset(place_33O,(TOTTR==1 & (MODE==5|MODE==6|MODE==7|MODE==8|MODE==10)))
head(place_33dO)
```

The raw frequency table is here.

```{r lc-28,cache=TRUE}
## Create frequency for different mode category for drive alone
temp1<-as.data.frame(table(place_33dO$Mode))
## Create percentage of frequency
temp1$Percent<-temp1$Freq*100/sum(temp1$Freq)
## Names 
names(temp1)
## Rename the variables
colnames(temp1)[c(1:3)] <- c("Mode","Observations_UW","Percent_UW")
## Show the results
temp1
```

This is followed by the weighted frequency table.

```{r lc-29,cache=TRUE}
## Create weighted frequency
temp2<-aggregate(NEXPTCFPERWGT~Mode,FUN=sum,data=place_33dO)
## Create weighted percentage
temp2$Percent<-temp2$NEXPTCFPERWGT*100/sum(temp2$NEXPTCFPERWGT)
## Names
names(temp2)
## Rename the variables
colnames(temp2)[c(1:3)] <- c("Mode","Observations_W","Percent_W")
## Show
temp2
```

And here we again save the results.

```{r lc-30,cache=TRUE}
## Merge the two files and save
temp3<-merge(temp1,temp2, by ="Mode", all = TRUE,sort=FALSE)
temp3 <- temp3[,c(1,2,4,3,5)]
temp3
temp3[,1]<-as.character(temp3[,1])
temp3<- rbind(temp3, c("Total", colSums(temp3[,2:5],na.rm = TRUE)))
temp3
## Save as csv file
write.csv(temp3,file="Mode_33Odrialn.csv",row.names=FALSE)
rm(temp1)
rm(temp2)
rm(temp3)
```

```{r results='asis',echo=FALSE,cache=FALSE}
############################################################ NESTABLE SECTION
SEC('##','Trip Purpose Calculations')
```


Now we need trip purpose breakdown, so we need to import activity level data with the 
appropriate weights.  Here we load and merge them.

```{r lc-31,cache=TRUE}
#Import the activity level variables csv file and give a new name
library(chts2011)
#act_data <- read.csv("c:/Users/user/Documents/Suman_research/Caltrans Project/Data/deliv_activity.csv", header=TRUE)
act_data <- chts_act
names(act_data)

# Merge the person weight file with place level data file
act_place<-merge(act_data,nplacedata, by =c("SAMPN","PERNO","PLANO"), all = FALSE)
names(act_place)
```

For display purposes, we convert the purpose codes into readable strings.

```{r lc-32, cache=TRUE}
act_place$Actpur<-ifelse(
  act_place$APURP==1," Personal Activities",
  ifelse
  (act_place$APURP==2," Preparing Meals/Eating",
   ifelse
   (act_place$APURP==3," Hosting Visitors/Entertaining Guests",
    ifelse
    (act_place$APURP==4," Exercise ",
     ifelse
     (act_place$APURP==5," Study/Schoolwork ",
      ifelse
      (act_place$APURP==6," Work For Pay At Home Using ",
       ifelse
       (act_place$APURP==7," Using Computer/Telephone/Cell Or Smart Phone",
        ifelse
        (act_place$APURP==8," All Other Activities At My Home ",
         ifelse
         (act_place$APURP==9," Work/Job Duties ",
          ifelse
          (act_place$APURP==10," Training",
           ifelse
           (act_place$APURP==11," Meals At Work ",
            ifelse
            (act_place$APURP==12," Work-Sponsored Social Activities ",
                                                                                                                                                      ifelse
             (act_place$APURP==13," Non-Work Related Activities",
              ifelse
              (act_place$APURP==14," Exercise/Sports ",
               ifelse
               (act_place$APURP==15," Volunteer Work/Activities",
                ifelse
                (act_place$APURP==16," All Other Work-Related Activities At My Work ",
                 ifelse
                 (act_place$APURP==17," In School/Classroom/Laboratory ",
                  ifelse
                  (act_place$APURP==18," Meals At School/College ",
                   ifelse
                   (act_place$APURP==19," After School Sports/Physical Activity ",
                    ifelse
                    (act_place$APURP==20," All Other After School Related Activities ",
                     ifelse
                     (act_place$APURP==21," Change Type Of Transportation/Transfer ", 
                      ifelse
                      (act_place$APURP==22," Pickup/Drop Off Passenger(S)", 
                       ifelse
                       (act_place$APURP==23," Drive Through Meals ", 
                        ifelse
                        (act_place$APURP==24," Drive Through Other (ATM, Bank) ",
                         ifelse
                         (act_place$APURP==25," Work-Related",
                          ifelse
                          (act_place$APURP==26," Service Private Vehicle",
                           ifelse
                           (act_place$APURP==27," Routine Shopping ",
                            ifelse
                            (act_place$APURP==28," Shopping For Major Purchases",
                             ifelse
                             (act_place$APURP==29," Household Errands", 
                              ifelse
                              (act_place$APURP==30," Personal Business ", 
                               ifelse
                               (act_place$APURP==31," Eat Meal at Restaurant/Diner", 
                                ifelse
                                (act_place$APURP==32," Health Care", 
                                 ifelse
                                 (act_place$APURP==33," Civic/Religious Activities", 
                                  ifelse
                                  (act_place$APURP==34," Outdoor Exercise", 
                                   ifelse
                                   (act_place$APURP==35," Indoor Exercise", 
                                    ifelse
                                    (act_place$APURP==36," Entertainment ", 
                                     ifelse
                                     (act_place$APURP==37," Social/Visit Friends/Relatives", 
                                      ifelse
                                      (act_place$APURP==38," Other ", 
                                       ifelse
                                       (act_place$APURP==39," Loop Trip", 
                                        " Don't Know/Refused")))))))))))))))))))))))))))))))))))))))
head(act_place)
```

```{r results='asis',echo=FALSE,cache=FALSE}
############################################################ NESTABLE SECTION
SEC('###','Trip Purposes from Lake County')
```


We start with trips from Lake County, subsetting based open the origin county `OCTFIP=33`.

```{r lc-33, cache=TRUE}
# Now we are interested in trips from Lake county (OCTFIP=33) 
# Create a subset for trips from Lake County 
act_33O<-subset(act_place,OCTFIP==33)
names(act_33O)
head(act_33O)
```

Next, we compute the observations for trip purpose for trips from Lake based on new weights.

```{r lc-34,cache=TRUE}
## Generate frequency for trip Mode
temp1<-as.data.frame(table(act_33O$Actpur))
## Create percentage of frequency
temp1$Percent<-temp1$Freq*100/sum(temp1$Freq)
## Names 
names(temp1)
## Rename the variables
colnames(temp1)[c(1:3)] <- c("Actpur","Ob_UW","Per_UW")
## Show the results
temp1
```

This is followed by the weighted trip purpose frequencies.

```{r lc-35,cache=TRUE}
## Create weighted frequency
temp2<-aggregate(NEXPTCFPERWGT~Actpur,FUN=sum,data=act_33O)
## Create weighted percentage
temp2$Percent<-temp2$NEXPTCFPERWGT*100/sum(temp2$NEXPTCFPERWGT)
## Names
names(temp2)
## Rename the variables
colnames(temp2)[c(1:3)] <- c("Actpur","Ob_W","Per_W")
## Show
temp2
```

And here we save the results.

```{r lc-36}
## Merge the two files and save
temp3<-merge(temp1,temp2, by ="Actpur", all = TRUE,sort=FALSE)
temp3 <- temp3[,c(1,2,4,3,5)]
temp3
temp3[,1]<-as.character(temp3[,1])
temp3<- rbind(temp3, c("Total", colSums(temp3[,2:5],na.rm = TRUE)))
temp3
## Save as csv file
write.csv(temp3,file="Actpur_33O.csv",row.names=FALSE)
rm(temp1)
rm(temp2)
rm(temp3)
```


```{r results='asis',echo=FALSE,cache=FALSE}
############################################################ NESTABLE SECTION
SEC('##','Work Trips by Mode')
```


The last sub question wanted work trips by Mode for Lake County.  


```{r results='asis',echo=FALSE,cache=FALSE}
############################################################ NESTABLE SECTION
SEC('###','Work Trips Originating in Lake County')
```


First, we reuse the merged activity table from above for trips activities (and associated trips) originating in Lake 
County, and subset it for work trips only.

```{r lc-37}
# Create a subset for work trips
actw_33O<-subset(act_33O,APURP==9)
```

Now, we can determine the observations of each Mode used for work trips.

```{r lc-38,cache=TRUE}
## Generate frequency for trip Mode
temp1<-as.data.frame(table(actw_33O$Mode))
## Create percentage of frequency
temp1$Percent<-temp1$Freq*100/sum(temp1$Freq)
## Names 
names(temp1)
## Rename the variables
colnames(temp1)[c(1:3)] <- c("Mode","Ob_UW","Per_UW")
## Show the results
temp1
```

Given those, we can expand them with the weights for Lake.

```{r lc-39,cache=TRUE}
## Create weighted frequency
temp2<-aggregate(NEXPTCFPERWGT~Mode,FUN=sum,data=actw_33O)
## Create weighted percentage
temp2$Percent<-temp2$NEXPTCFPERWGT*100/sum(temp2$NEXPTCFPERWGT)
## Names
names(temp2)
## Rename the variables
colnames(temp2)[c(1:3)] <- c("Mode","Ob_W","Per_W")
## Show
temp2
```

And again, we finish by saving the results.

```{r lc-40,cache=TRUE}
## Merge the two files and save
temp3<-merge(temp1,temp2, by ="Mode", all = TRUE,sort=FALSE)
temp3 <- temp3[,c(1,2,4,3,5)]
temp3
temp3[,1]<-as.character(temp3[,1])
temp3<- rbind(temp3, c("Total", colSums(temp3[,2:5],na.rm = TRUE)))
temp3
## Save as csv file
write.csv(temp3,file="Mode_wact33O.csv",row.names=FALSE)
rm(temp1)
rm(temp2)
rm(temp3)
```

```{r results='asis',echo=FALSE,cache=FALSE}
############################################################ NESTABLE SECTION
SEC('###','Work Trips Ending in Lake County')
```


Now we want to find trips ending in Lake county `(CTFIP=33)`.  Again, we subset
the place table and adjust the weights to be Lake weights if the origins are in
Lake, and to use statewide weights if the origins are outside Lake.


```{r lc-41,cache=TRUE}
# Create a subset for trips from Lake County 
act_33D<-subset(act_place,CTFIP==33)
names(act_33D)
head(act_33D)
# Change the weight for non Lake county origin (Assign state wide weight for non Lake origin)
act_33D$NTCFW <- ifelse((act_33D$OCTFIP==33), act_33D$NEXPTCFPERWGT, act_33D$EXPTCFPERWGT.y)
names(act_33D)
head(act_33D)
```

Again, we generate a frequency table by purpose.

```{r lc-42}
# Generate observations by trip purpose from Lake based on new weights
## Generate frequency for trip Mode
temp1<-as.data.frame(table(act_33D$Actpur))
## Create percentage of frequency
temp1$Percent<-temp1$Freq*100/sum(temp1$Freq)
## Names 
names(temp1)
## Rename the variables
colnames(temp1)[c(1:3)] <- c("Actpur","Ob_UW","Per_UW")
## Show the results
temp1
```

We follow this with the weighted frequency table.

```{r lc-43,cache=TRUE}
## Create weighted frequency
temp2<-aggregate(NTCFW~Actpur,FUN=sum,data=act_33D)
## Create weighted percentage
temp2$Percent<-temp2$NTCFW*100/sum(temp2$NTCFW)
## Names
names(temp2)
## Rename the variables
colnames(temp2)[c(1:3)] <- c("Actpur","Ob_W","Per_W")
## Show
temp2
```

And again we save the result.

```{r lc-44,cache=TRUE}
## Merge the two files and save
temp3<-merge(temp1,temp2, by ="Actpur", all = TRUE,sort=FALSE)
temp3 <- temp3[,c(1,2,4,3,5)]
temp3
temp3[,1]<-as.character(temp3[,1])
temp3<- rbind(temp3, c("Total", colSums(temp3[,2:5],na.rm = TRUE)))
temp3
## Save as csv file
write.csv(temp3,file="Actpur_33D.csv",row.names=FALSE)
rm(temp1)
rm(temp2)
rm(temp3)
```

Finally, we compute the frequencies of various Modes used for work trips ending
in Lake County.

```{r lc-45,cache=TRUE}
# We want only work trips (APURP=9)
# Create a subset for work trips
actw_33D<-subset(act_33D,APURP==9)
## Generate frequency for trip Mode
temp1<-as.data.frame(table(actw_33D$Mode))
## Create percentage of frequency
temp1$Percent<-temp1$Freq*100/sum(temp1$Freq)
## Names 
names(temp1)
## Rename the variables
colnames(temp1)[c(1:3)] <- c("Mode","Ob_UW","Per_UW")
## Show the results
temp1
```

Here is the weighted frequency table.

```{r lc-46,cache=TRUE}
## Create weighted frequency
temp2<-aggregate(NTCFW~Mode,FUN=sum,data=actw_33D)
## Create weighted percentage
temp2$Percent<-temp2$NTCFW*100/sum(temp2$NTCFW)
## Names
names(temp2)
## Rename the variables
colnames(temp2)[c(1:3)] <- c("Mode","Ob_W","Per_W")
## Show
temp2
## Merge the two files and save
temp3<-merge(temp1,temp2, by ="Mode", all = TRUE,sort=FALSE)
temp3 <- temp3[,c(1,2,4,3,5)]
temp3
temp3[,1]<-as.character(temp3[,1])
temp3<- rbind(temp3, c("Total", colSums(temp3[,2:5],na.rm = TRUE)))
temp3
```

...and one last time we save it.

```{r lc-47,cache=TRUE}
## Save as csv file
write.csv(temp3,file="Mode_wact33D.csv",row.names=FALSE)
rm(temp1)
rm(temp2)
rm(temp3)
```

