---
title: "Long Distance Trip Questions"
author: "Suman Mitra"
date: "September 3, 2015"
output: html_document
---

```{r ldtq-setup-0,echo=FALSE,cache=FALSE}
if ( !exists("rd") ) { rd <<- gsub("(.*?chts-book[^/]*).*$","\\1",getwd()) } 
source(paste(rd,"R/setup-hooks.R",sep="/"))
write(paste("Using Root Directory",rd,"\n"),stderr())
```

In this subsection, we generate tables for number of non-commute, long
distance trips by trip mode for counties in the following geographic
areas:

  *	District 1 - Del Norte, Humboldt, Lake, Mendocino
  * District 2 - Lassen, Modoc, Plumas, Shasta, Siskiyou, Tehama,
       Trinity
  * District 3 - Butte, Colusa, El Dorado, Glenn, Nevada, Placer,
       Sacramento, Sierra, Sutter, Yolo, Yuba
  * District 4 - Alameda, Contra Costa, Marin, Napa, San Mateo, San
       Francisco, Santa Clara, Solano, Sonoma
  * District 5 - Monterey, San Benito, San Luis Obispo, Santa Barbara,
       Santa Cruz
  * Southern California - Fresno, Madera, Kings, Tulare, Kern, Los
       Angeles, Orange, San Diego, Imperial, Ventura, Riverside and a
       portion of San Bernardino
  *	District 9 - Inyo, Mono and the rest of San Bernardino County.


# Summary and Methodology

We generate weighted and unweighted tables for trips originating in
and ending in the above districts, using the long distance recall
survey data. Note that in the long distance recall survey, trips 50
miles or longer are classified as long distance trips.  We exclude
work trips to calculate non-commute trips and use the new countywide
long distance weights described previously to generate weighted
results.

\ifmydocumentation

# R Code and Implementation


After setting our working directory, let us import the long distance data with the corresponding weights:

```{r ldtq-1, }
# Set the working directory
setwd(pd("LD data request"))

#Import the long distance  (with new weight) csv file and give a new name
wLDdata <- read.csv(pd("Weight/LD_Weight2/wLD_data922015.csv"), header=TRUE)
names(wLDdata)
```

Next, we convert the `LDMODE1` codes into text descriptions:

```{r ldtq-2}
# For recoding Mode
wLDdata$Mode<-
  ifelse(wLDdata$LDMODE1==1,"Walk",
  ifelse(
    wLDdata$LDMODE1==2,"Bike",
  ifelse(
    wLDdata$LDMODE1==3,"Wheelchair / Mobility Scooter",
  ifelse(
    wLDdata$LDMODE1==4,"Other Non-Motorized",
  ifelse(
    wLDdata$LDMODE1==5,"Auto/Van/Truck Driver",
  ifelse(
    wLDdata$LDMODE1==6,"Auto /Van/Truck Passenger",
  ifelse(
    wLDdata$LDMODE1==7,"Carpool/Vanpool",
  ifelse(
    wLDdata$LDMODE1==8,"Motorcycle/Scooter/Moped",
  ifelse(
    wLDdata$LDMODE1==9,"Taxi/Hired Car/Limo",
  ifelse(
    wLDdata$LDMODE1==10,"Rental Car/Vehicle",
  ifelse(
    wLDdata$LDMODE1==11,"Private shuttle",
  ifelse(
    wLDdata$LDMODE1==12,"Greyhound Bus",
  ifelse(
    wLDdata$LDMODE1==13,"Plane",
  ifelse(
    wLDdata$LDMODE1==14,"Other Private Transit",
  ifelse(
    wLDdata$LDMODE1==15,"Local Bus,Rapid Bus",
  ifelse(
    wLDdata$LDMODE1==16,"Express Bus/Commuter Bus",
  ifelse(
    wLDdata$LDMODE1==17,"Premium Bus",
  ifelse(
    wLDdata$LDMODE1==18,"School Bus",
  ifelse(
    wLDdata$LDMODE1==19,"Public Transit Shuttle",
  ifelse(
    wLDdata$LDMODE1==20,"Air BART/LAX Fly Away",
  ifelse(
    wLDdata$LDMODE1==21,"Dial-a-Ride/Paratransit", 
  ifelse(
    wLDdata$LDMODE1==22,"Amtrak Bus", 
  ifelse(
    wLDdata$LDMODE1==23,"Other Bus ", 
  ifelse(
    wLDdata$LDMODE1==24,"BART,Metro Red/Purple Line",
  ifelse(
    wLDdata$LDMODE1==25,"ACE, Amtrak,  etc",
  ifelse(
    wLDdata$LDMODE1==26,"Metro Blue / Green / Gold Line etc",
  ifelse(
    wLDdata$LDMODE1==27,"Street Car / Cable Car",
  ifelse(
    wLDdata$LDMODE1==28,"Other Rail",
  ifelse(
    wLDdata$LDMODE1==29,"Ferry / Boat","Don't Know/Refused"
  )))))))))))))))))))))))))))))
```


## Modal Split of Trips from District 1


Now, let us focus on long distance trips originating from District 1. we
subset the long distance data for trips originating in District 1:

```{r ldtq-3}
# Create a subset of variables for districts (Here District==1)
Odist1<-subset(wLDdata,LDOCTFIP==15|LDOCTFIP==23|LDOCTFIP==33|LDOCTFIP==45)
table(Odist1$LDOCTFIP)
```

Then we subset the data for non-commute trips:

```{r ldtq-4}
# Creat subset for non-commute trips
Odistrict1<-subset(Odist1,!(LDTPURP==1))
table(Odistrict1$LDTPURP)
```

The variable `Odistrict1` now contains the non-commute long distance trips
originiating in District 1 with appropriate expansion weights. 
At this point, we can generate observations for modal split for
non-commute long distance trips from District 1 based on these
weights.

Notes: In the tables:  UW = Unweighted and W = Weighted.

```{r ldtq-5, }
## Generate frequency for trip mode
temp1<-as.data.frame(table(Odistrict1$Mode))
## Create percentage of frequency
temp1$Percent<-temp1$Freq*100/sum(temp1$Freq)
## Names 
names(temp1)
## Rename the variables
colnames(temp1)[c(1:3)] <- c("LD_Mode","Observations(UW)","Percent(UW)")
## Show the results
temp1
```

Let us compute the weighted frequencies:

```{r ldtq-6, }
## Create weighted frequency
temp2<-aggregate(EXPLDWGT~Mode,FUN=sum,data=Odistrict1)
## Create weighted percentage
temp2$Percent<-temp2$EXPLDWGT*100/sum(temp2$EXPLDWGT)
## Names
names(temp2)
## Rename the variables
colnames(temp2)[c(1:3)] <- c("LD_Mode","Observations(W)","Percent(W)")
## Show
temp2
```

Save the results in a `csv` file for later use:

```{r ldtq-7, }
## Merge the two files and save
temp3<-merge(temp1,temp2, by ="LD_Mode", all = TRUE,sort=FALSE)
temp3 <- temp3[,c(1,2,4,3,5)]
temp3
write.csv(temp3,file=pto("O_Mode.csv"),row.names=FALSE)
```

Now we have modal split data based on new weights for non-commute long
distance trips from District 1.


## Modal Split of Trips from District 2

Let us use the same procedure to find the total number of non-commute
long distance trips from District 2, by mode.

First, subset the long distance data for trips originating in
District 2.

```{r ldtq-8, }
# For origin in District2
## Creat subset
Odist2<-subset(wLDdata,
               LDOCTFIP==35|LDOCTFIP==49|LDOCTFIP==63|LDOCTFIP==89|
                 LDOCTFIP==93|LDOCTFIP==103|LDOCTFIP==105)
table(Odist2$LDOCTFIP)
```

Then, subset the non-commute long distance trips originating in District 2:

```{r ldtq-9, }
## create subset for non-commute trips
Odistrict2<-subset(Odist2,!(LDTPURP==1))
table(Odistrict2$LDTPURP)
```

Let us generate trip frequency:

```{r ldtq-10, }
## Generate frequency and percentage for trip mode
O_mode2<-as.data.frame(table(Odistrict2$Mode))
O_mode2$Percent<-O_mode2$Freq*100/sum(O_mode2$Freq)
names(O_mode2)
colnames(O_mode2)[c(1:3)] <- c("LD_Mode","Observations(UW)","Percent(UW)")
O_mode2
```

Now, compute weighted frequencies:

```{r ldtq-11, }
temp2<-aggregate(EXPLDWGT~Mode,FUN=sum,data=Odistrict2)
temp2$Percent<-temp2$EXPLDWGT*100/sum(temp2$EXPLDWGT)
names(temp2)
colnames(temp2)[c(1:3)] <- c("LD_Mode","Observations(W)","Percent(W)")
temp2
```

Save the results for later use:

```{r ldtq-12, }
temp3<-merge(O_mode2,temp2, by ="LD_Mode", all = TRUE,sort=FALSE)
temp3 <- temp3[,c(1,2,4,3,5)]
temp3
write.csv(temp3,file=pto("O_LDMode2.csv"),row.names=FALSE)
```



## Modal Split of Trips to District 1


We perform similar procedures for trips *to* District 1.  As
before, we subset the long distance data, but this time for trips
ending in District 1 and keep only non-commute trips.


```{r ldtq-33, }
# Create a subset of variables for districts (Here District==1)
ddist1<-subset(wLDdata,LDDCTFIP==15|LDDCTFIP==23|LDDCTFIP==33|LDDCTFIP==45)
table(ddist1$LDDCTFIP)
```

Then we subset the data for non-commute trips:

```{r ldtq-34, }
# Creat subset for non-commute trips
ddistrict1<-subset(ddist1,!(LDTPURP==1))
table(ddistrict1$LDTPURP)
```

Th evariable `ddistrict1` contains the non-commute long distance trips
ending in District 1 with appropriate expansion weights
appended. Let us generate observations for modal split for
non-commute long distance trips to District 1 based on these weights:


```{r ldtq-35, }
## Generate frequency for trip mode
temp1<-as.data.frame(table(ddistrict1$Mode))
## Create percentage of frequency
temp1$Percent<-temp1$Freq*100/sum(temp1$Freq)
## Names 
names(temp1)
## Rename the variables
colnames(temp1)[c(1:3)] <- c("LD_Mode","Observations(UW)","Percent(UW)")
## Show the results
temp1
```

Now compute the weighted frequencies:

```{r ldtq-36, }
## Create weighted frequency
temp2<-aggregate(EXPLDWGT~Mode,FUN=sum,data=ddistrict1)
## Create weighted percentage
temp2$Percent<-temp2$EXPLDWGT*100/sum(temp2$EXPLDWGT)
## Names
names(temp2)
## Rename the variables
colnames(temp2)[c(1:3)] <- c("LD_Mode","Observations(W)","Percent(W)")
## Show
temp2
```

Save these results in a `csv` file for later use.

```{r ldtq-37, }
## Merge the two files and save
temp3<-merge(temp1,temp2, by ="LD_Mode", all = TRUE,sort=FALSE)
temp3 <- temp3[,c(1,2,4,3,5)]
temp3<- rbind(temp3, c("Total", colSums(temp3[,2:5],na.rm = TRUE)))
temp3

write.csv(temp3,file=pto("D_OMode.csv"),row.names=FALSE)
mode.o.d1 <- temp3
```

Now we have modal split data based on new weight for non-commute long
distance trips to district 1.


## Modal Split of Trips to District 2

Next, we follow similar steps to find the total number of non-commute
long distance trips to District 2, by mode.

First we subset the long distance data for trips ending in District 2.

```{r ldtq-38, }
# For destination in District2
## Creat subset
ddist2<-subset(wLDdata,LDDCTFIP==35|LDDCTFIP==49|LDDCTFIP==63|LDDCTFIP==89|LDDCTFIP==93|LDDCTFIP==103|LDDCTFIP==105)
table(ddist2$LDDCTFIP)
```

Second, we subset the non-commute long distance trips ending in District 2.

```{r ldtq-39, }
## create subset for non-commute trips
ddistrict2<-subset(ddist2,!(LDTPURP==1))
table(ddistrict2$LDTPURP)
```

Let us generate trip frequencies:

```{r ldtq-40, }
## Generate frequency and percentage for trip mode
O_mode2<-as.data.frame(table(ddistrict2$Mode))
O_mode2$Percent<-O_mode2$Freq*100/sum(O_mode2$Freq)
names(O_mode2)
colnames(O_mode2)[c(1:3)] <- c("LD_Mode","Observations(UW)","Percent(UW)")
O_mode2
```

Now compute weighted frequencies:

```{r ldtq-41, }
temp2<-aggregate(EXPLDWGT~Mode,FUN=sum,data=ddistrict2)
temp2$Percent<-temp2$EXPLDWGT*100/sum(temp2$EXPLDWGT)
names(temp2)
colnames(temp2)[c(1:3)] <- c("LD_Mode","Observations(W)","Percent(W)")
temp2
```

As before, save the results for later use:

```{r ldtq-42, }
temp3<-merge(O_mode2,temp2, by ="LD_Mode", all = TRUE,sort=FALSE)
temp3 <- temp3[,c(1,2,4,3,5)]
temp3<- rbind(temp3, c("Total", colSums(temp3[,2:5],na.rm = TRUE)))
temp3
write.csv(temp3,file=pto("D_OLDMode2.csv"),row.names=FALSE)
mode.o.d2 <- temp3
```


## Generate similar statistics for other districts

We can generate similar tables for other districts (District 3, 4, 5, 9 and Southern California) using the above codes. To do that we need to subset the data using county fip numbers of all counties of resepective district.

For example, we could generate statistics for trip originating from District 3 which includes the following counties:
Butte, Colusa, El Dorado, Glenn, Nevada, Placer, Sacramento, Sierra, Sutter, Yolo, Yuba. 
In that case, we would subset the data based on the county fip nubmer of these counties using following codes:

```{r echo=TRUE,eval=FALSE}
Odist3 <- subset(
    wLDdata,
    LDOCTFIP==7|LDOCTFIP==11|LDOCTFIP==17|LDOCTFIP==21|
    LDOCTFIP==57|LDOCTFIP==61|LDOCTFIP==67|LDOCTFIP==91|
    LDOCTFIP==101|LDOCTFIP==113|LDOCTFIP==115
)
```

For generating tables, we would use the 'Odist3' file, which contains
all trips originating from District 3.  To generate frequency for trip
mode we would use the following command:
`temp1<-as.data.frame(table(Odistrict3$Mode))`, where we just changed
the `Odistrict2` file to `Odistrict3`.  To produce similar results for
the other districts, we would subset the data following the same
procedure and we would use the new datafile to generate the tables.

To generate trips ending in District 3, we would subset the data using the same steps. 
For example: We would use

```{r echo=TRUE,eval=FALSE}
ddist3 <- subset(
    wLDdata,
    LDDCTFIP==7|LDDCTFIP==11|LDDCTFIP==17|LDDCTFIP==21|
    LDDCTFIP==57|LDDCTFIP==61| LDDCTFIP==67|LDDCTFIP==91|
    LDDCTFIP==101|LDDCTFIP==113|LDDCTFIP==115
)
```

We would need to use `ddist3` to generate tables for district 3. For
example, to generate frequency for trip mode for trips ending in
district 3, we would use the command
`temp1<-as.data.frame(table(ddist3$Mode))`, i.e., we would just change
the `ddist2` data file to `ddist3`.  We need to apply same procedure
to generate similar results for other districts.

Note: We used ArcGis 10.2 to select the households located within the
selected portion of San Bernardino county (For Southern
California). We imported the x-y coordinates of the location of CHTS
household respondents to ArcMap and converted it to a shapefile. We
then performed intersect operation in ArcGis to subset the sample of
respondents who reside within selected portion of San Bernardino
county.  We followed the same procedure for subsetting the sample of
respondents who reside within the rest of San Bernardino county (For
District 9).

\fi

\newpage

# Results

```{r ldtq-res-1,echo=FALSE,results='asis'}
mode.o.d1 <- mode.o.d1 %>%
    arrange(ifelse(LD_Mode=="Total",1,0),desc(as.numeric(mode.o.d1$`Observations(W)`))) %>%
    transmute(`Long Distance Mode` = LD_Mode,
              `Observed Unweighted` = obsfmt(`Observations(UW)`),
              `Observed Weighted` = obsfmt(`Observations(W)`),
              `Percent Unweighted` = pctfmt(`Percent(UW)`),
              `Percent Weighted` = pctfmt(`Percent(W)`)
              )
pandoc.table(
    mode.o.d1,
    caption="Modal split of trips originating in District 1",
    justify=c('left','right','right','right','right'),
    split.tables=100,
    split.cells=c('40%','15%','15%','15%','15%'),
    big.mark=",",
    emphasize.strong.rows=which(mode.scc.orig$Mode=='Total',arr.ind=TRUE)
)

```

```{r ldtq-res-2,echo=FALSE,results='asis'}
mode.o.d2 <- mode.o.d2 %>%
    arrange(ifelse(LD_Mode=="Total",1,0),desc(as.numeric(mode.o.d2$`Observations(W)`))) %>%
    transmute(`Long Distance Mode` = LD_Mode,
              `Observed Unweighted` = obsfmt(`Observations(UW)`),
              `Observed Weighted` = obsfmt(`Observations(W)`),
              `Percent Unweighted` = pctfmt(`Percent(UW)`),
              `Percent Weighted` = pctfmt(`Percent(W)`)
              )
pandoc.table(
    mode.o.d2,
    caption="Modal split of trips originating in District 2",
    justify=c('left','right','right','right','right'),
    split.tables=100,
    split.cells=c('40%','15%','15%','15%','15%'),
    big.mark=",",
    emphasize.strong.rows=which(mode.scc.orig$Mode=='Total',arr.ind=TRUE)
)

```

