---
title: "Travel Statistics along the SR-1 Corridor"
author: "Suman Mitra"
date: "September 4, 2015"
output: pdf_document
---

```{r sr1-setup-0,echo=FALSE,cache=FALSE}
if ( !exists("rd") ) { rd <<- gsub("(.*?chts-book[^/]*).*$","\\1",getwd()) } 
source(paste(rd,"R/setup-hooks.R",sep="/"))
write(paste("Using Root Directory",rd,"\n"),stderr())
```

```{r results='asis',echo=FALSE,cache=FALSE}
############################################################ NESTABLE SECTION
SEC('#','Summary and Methodology')
```


The purpose of this subsection is to show how 2012 CHTS data can be used
to generate useful statistics in a corridor provided enough CHTS observations are available in that coridor.
For the SR-1 corridor that extends from the Golden Gate in San Francisco to Pacifica, we found 269 household respondents (622 person respondents) including 98 GPS respondents within 1 km on each side of the SR-1.
The size of this sample i ssufficient to generate tables by mode for 1) total number of trips; 
2) total trip duration; 3) average trip duration; 4) total trip distance; and 5) average trip distance.
We also created tables for the total number of trips by trip purpose (task 6) and for
work trips by mode (task 7). We generated tables for total number of trips by activity purpose since 
the trip purpose is given in the database as activity purpose and there is more than one activity
in many single trips. We considered each activity as a trip. We only considered the purpose code (APURP) 9 
for work trips.  We generated every table for both unweighted and weighted results using countywide
weights for the latter.


```{r results='asis',echo=FALSE,cache=FALSE}
############################################################ NESTABLE SECTION
SEC('#','R Code and Implementation')
```


```{r sr1-0,echo=FALSE}
if ( !exists("rd") ) { rd <- "/home/crindt/chts-book" }
p <- function(ps) { return(paste(rd,ps,sep="/")) }
```


To create these tables, we need the place file with the new countywide
person weights that were computed in the section on
[County Level Person Weights].


```{r sr1-1, }
#Import the place level with new weight csv file and give a new name
nplace<- read.csv(p("Weight/NewWeightData/nplacedata3032015.csv"), header=TRUE)
```

Next, we convert the `MODE` codes into text descriptions:

```{r sr1-2, }
# For recoding Mode
nplace$Mode<-
  ifelse(
    nplace$MODE==1,"Walk",
  ifelse(
      nplace$MODE==2,"Bike",
  ifelse(
    nplace$MODE==3,"Wheelchair/Mobility Scooter",
  ifelse(
    nplace$MODE==4,"Other Non-Motorized",
  ifelse(
    nplace$MODE==5,"Auto/Van/Truck Driver",
  ifelse(
    nplace$MODE==6,"Auto /Van/Truck Passenger",
  ifelse(
    nplace$MODE==7,"Carpool/Vanpool",
  ifelse(
    nplace$MODE==8,"Motorcycle",
  ifelse(
    nplace$MODE==9,"Taxi/Hired Car/Limo",
  ifelse(
    nplace$MODE==10,"Rental Car/Vehicle",
  ifelse(
    nplace$MODE==11,"Private shuttle ",
  ifelse(
    nplace$MODE==12,"Greyhound Bus",
  ifelse(
    nplace$MODE==13,"Plane",
  ifelse(
    nplace$MODE==14,"Other Private Transit",
  ifelse(
    nplace$MODE==15,"Local Bus,Rapid Bus",
  ifelse(
    nplace$MODE==16,"Express Bus",
  ifelse(
    nplace$MODE==17,"Premium Bus",
  ifelse(
    nplace$MODE==18,"School Bus",
  ifelse(
    nplace$MODE==19,"Public Transit Shuttle",
  ifelse(
    nplace$MODE==20,"Air BART/LAX Fly Away",
  ifelse(
    nplace$MODE==21,"Dial-a-Ride/Paratransit", 
  ifelse(
    nplace$MODE==22,"Amtrak Bus", 
  ifelse(
    nplace$MODE==23,"Other Bus ", 
  ifelse(
    nplace$MODE==24,"BART,Metro",
  ifelse(
    nplace$MODE==25,"ACE, Amtrak etc",
  ifelse(
    nplace$MODE==26,"Metro",
  ifelse(
    nplace$MODE==27,"Street Car",
  ifelse(
    nplace$MODE==28,"Other Rail",
  ifelse(
    nplace$MODE==29,"Ferry / Boat","Don't Know/Refused"
  )))))))))))))))))))))))))))))
```

We found 269 household respondents (622 person respondents) including
98 GPS respondents within 1 km on each side of the stretch of the SR1 that extends from San Francisco
Golden Gate Bridge to Pacifica using ArcGis and we saved it in a
'csv' file. Let us import that file:

```{r sr1-3, }
# Import SR1 sample households (obtained from GIS) 
sr1 <- read.csv(p("District 4/SR1/SR1_block1km.csv"), header=TRUE)
```

Now let us merge this file with the place file:

```{r sr1-4, }
# Merge the SR1 sample data with place level data
sr1_place<-merge(nplace,sr1, by ="SAMPN", all = FALSE)

# Check the Length of the sampn number of new file (it should be 269)
length(unique(sr1_place$SAMPN))
names(sr1_place)
head(sr1_place)
```


```{r results='asis',echo=FALSE,cache=FALSE}
############################################################ NESTABLE SECTION
SEC('##','Modal split for SR1 corridor')
```


Task 1: Now, let us generate trip frequencies by mode:

```{r sr1-5, }
## Generate frequency for trip Mode
temp1<-as.data.frame(table(sr1_place$Mode))
temp1
## Create percentage of frequency
temp1$Percent<-temp1$Freq*100/sum(temp1$Freq)
temp1
## Names 
names(temp1)
## Rename the variables
colnames(temp1)[c(1:3)] <- c("Mode","Ob_UW","Per_UW")
## Show the results
temp1
```

Second, let us compute weighted trip frequencies:

```{r sr1-6, }
## Create weighted frequency
temp2<-aggregate(NEXPTCFPERWGT~Mode,FUN=sum,data=sr1_place)
## Create weighted percentage
temp2$Percent<-temp2$NEXPTCFPERWGT*100/sum(temp2$NEXPTCFPERWGT)
## Names
names(temp2)
## Rename the variables
colnames(temp2)[c(1:3)] <- c("Mode","Ob_W","Per_W")
## Show
temp2
```

Let us save these results in a `csv` file for later use:

```{r sr1-7, }
## Merge the two files and save
temp3<-merge(temp1,temp2, by ="Mode", all = TRUE,sort=FALSE)
temp3 <- temp3[,c(1,2,4,3,5)]
temp3
Mode_sr1<-temp3
temp3[,1]<-as.character(temp3[,1])
temp3<- rbind(temp3, c("Total", colSums(temp3[,2:5],na.rm = TRUE)))
temp3
## Save as csv file
write.csv(temp3,file="Mode_sr1.csv",row.names=FALSE)
rm(temp1)
rm(temp2)
rm(temp3)
```


```{r results='asis',echo=FALSE,cache=FALSE}
############################################################ NESTABLE SECTION
SEC('##','Total Trip duration for all trips by mode for SR1 corridor')
```


Task 2: Next, we apply a similar procedure to computate Total Trip durations for all
trips in the SR1 corridor, by mode:

```{r sr1-8, }
## Generate frequency for trip Mode
temp1<-aggregate(TRIPDUR~Mode,FUN=sum,data=sr1_place)
## Create  percentage
temp1$Percent<-temp1$TRIPDUR*100/sum(temp1$TRIPDUR)
## Names
names(temp1)
## Rename the variables
colnames(temp1)[c(1:3)] <- c("Mode","TDu_UW","Per_UW")
## Show
temp1
```

Let us find the weighted frequency:

```{r sr1-9, }
## Create weighted frequency
temp<-sr1_place$NEXPTCFPERWGT*sr1_place$TRIPDUR
temp2<-aggregate(temp~Mode,FUN=sum,data=sr1_place)
## Create weighted percentage
temp2$Percent<-temp2$temp*100/sum(temp2$temp)
## Names
names(temp2)
## Rename the variables
colnames(temp2)[c(1:3)] <- c("Mode","TDu_W","Per_W")
## Show
temp2
```

Task 3: Now calculate average trip duration:

```{r sr1-10,}
## Merge the two files and save
temp3<-merge(temp1,temp2, by ="Mode", all = TRUE)
temp3 <- temp3[,c(1,2,4,3,5)]
temp3
# Average trip duration
temp3$avg_uw<-temp3$TDu_UW/Mode_sr1$Ob_UW
temp3
temp3$avg_w<-temp3$TDu_W/Mode_sr1$Ob_W
temp3
temp3[,1]<-as.character(temp3[,1])
temp3<- rbind(temp3, c("Total", colSums(temp3[,2:7],na.rm = TRUE)))
temp3
```

Let us save results for later use:

```{r sr1-11,}
## Save as csv file
write.csv(temp3,file="TripDUrm_sr1.csv",row.names=FALSE)
rm(temp1)
rm(temp2)
rm(temp3)
rm(temp)
```


```{r results='asis',echo=FALSE,cache=FALSE}
############################################################ NESTABLE SECTION
SEC('###','Trip distance by mode')
```


Task 4: Next, let us calculate Total Trip distance for SR1 Corridor trips, by mode:


```{r sr1-12,}
## Generate frequency for trip Mode
temp1<-aggregate(TripDistance~Mode,FUN=sum,data=sr1_place)
## Create  percentage
temp1$Percent<-temp1$TripDistance*100/sum(temp1$TripDistance)
## Names
names(temp1)
## Rename the variables
colnames(temp1)[c(1:3)] <- c("Mode","TDis_UW","Per_UW")
## Show
temp1
```

Calculate weighted frequencies:

```{r sr1-13,}
## Create weighted frequency
temp<-sr1_place$NEXPTCFPERWGT*sr1_place$TripDistance
temp2<-aggregate(temp~Mode,FUN=sum,data=sr1_place)
## Create weighted percentage
temp2$Percent<-temp2$temp*100/sum(temp2$temp)
## Names
names(temp2)
## Rename the variables
colnames(temp2)[c(1:3)] <- c("Mode","TDis_W","Per_W")
## Show
temp2
```

Task 5: Now, find average trip distances:

```{r sr1-14,}
## Merge the two files and save
temp3<-merge(temp1,temp2, by ="Mode", all = TRUE)
temp3 <- temp3[,c(1,2,4,3,5)]
temp3
# Average trip distance
temp3$avg_uw<-temp3$TDis_UW/Mode_sr1$Ob_UW
temp3
temp3$avg_w<-temp3$TDis_W/Mode_sr1$Ob_W
temp3
temp3[,1]<-as.character(temp3[,1])
temp3<- rbind(temp3, c("Total", colSums(temp3[,2:7],na.rm = TRUE)))
temp3
```

And let us save results for later use in "TripDistanceM_sr1.csv":

```{r sr1-15,}
## Save as csv file
write.csv(temp3,file="TripDistanceM_sr1.csv",row.names=FALSE)
rm(temp1)
rm(temp2)
rm(temp3)
rm(temp)
```

```{r results='asis',echo=FALSE,cache=FALSE}
############################################################ NESTABLE SECTION
SEC('##','Trip Purpose Calculations')
```



NTask 6: for trip purpose data, we need to import activity level data with the 
appropriate weights.  Here we load and merge them:


```{r sr1-16,}
#Import the activity level data frame and give a new name
library(chts2011)
act_data <- chts_act
names(act_data)
# Merge the person weight file with place level data file
act_place<-merge(act_data,nplace, by =c("SAMPN","PERNO","PLANO"), all = FALSE)
names(act_place)
```

For display purposes, we convert the purpose codes into readable strings:

```{r sr1-17,}
# Recoding activity purpose
act_place$Actpur<-
  ifelse(
    act_place$APURP==1," Personal Activities",
  ifelse(
    act_place$APURP==2," Preparing Meals/Eating ",
  ifelse(
    act_place$APURP==3," Hosting Visitors",
  ifelse(
    act_place$APURP==4," Exercise ",
  ifelse(
    act_place$APURP==5," Study/Schoolwork ",
  ifelse(
    act_place$APURP==6," Work For Pay At Home",
  ifelse(
    act_place$APURP==7,"Using Computer/Telephone/Cell",
  ifelse(
    act_place$APURP==8," All Other Activities At My Home ",
  ifelse(
    act_place$APURP==9," Work/Job Duties ",
  ifelse(
    act_place$APURP==10," Training",
  ifelse(
    act_place$APURP==11," Meals At Work ",
  ifelse(
    act_place$APURP==12," Work-Sponsored Social Activities ",
  ifelse(
    act_place$APURP==13," Non-Work Related Activities",
  ifelse(
    act_place$APURP==14," Exercise/Sports ",
  ifelse(
    act_place$APURP==15," Volunteer Work/Activities ",
  ifelse(
    act_place$APURP==16," All Other Work-Related Activities",
  ifelse(
    act_place$APURP==17," In School/Classroom/Laboratory ",
  ifelse(
    act_place$APURP==18," Meals At School/College ",
  ifelse(
    act_place$APURP==19," After School Sports/Physical Activity ",
  ifelse(
    act_place$APURP==20," All Other After SchoolActivities ",
  ifelse(
    act_place$APURP==21," Change Type Of Transportation/Transfer ", 
  ifelse(
    act_place$APURP==22," Pickup/Drop Off Passenger(S) ", 
  ifelse(
    act_place$APURP==23," Drive Through Meals", 
  ifelse(
    act_place$APURP==24," Drive Through Other (ATM, Bank) ",
  ifelse(
    act_place$APURP==25," Work-Related",
  ifelse(
    act_place$APURP==26," Service Private Vehicle",
  ifelse(
    act_place$APURP==27," Routine Shopping ",
  ifelse(
    act_place$APURP==28," Shopping For Major Purchases",
  ifelse(
    act_place$APURP==29," Household Errands", 
  ifelse(
    act_place$APURP==30," Personal Business", 
  ifelse(
    act_place$APURP==31," Eat Meal at Restaurant/Diner", 
  ifelse(
    act_place$APURP==32," Health Care", 
  ifelse(
    act_place$APURP==33," Civic/Religious Activities", 
  ifelse(
    act_place$APURP==34," Outdoor Exercise", 
  ifelse(
    act_place$APURP==35," Indoor Exercise ", 
  ifelse(
    act_place$APURP==36," Entertainment ", 
  ifelse(
    act_place$APURP==37," Social/Visit Friends/Relatives", 
  ifelse(
    act_place$APURP==38," Other ", 
  ifelse(
    act_place$APURP==39," Loop Trip", " Don't Know/Refused"
  )))))))))))))))))))))))))))))))))))))))
```

Now, merge the these data with the SR1 data table:

```{r sr1-18,}
# Merge the SR1 sample data with activity level data
sr1_act<-merge(act_place,sr1, by ="SAMPN", all = FALSE)

# Check the Length of the sampn number of the new data frame
# (it should be 269)
length(unique(sr1_act$SAMPN))
names(sr1_act)
head(sr1_act)
```

```{r results='asis',echo=FALSE,cache=FALSE}
############################################################ NESTABLE SECTION
SEC('##','Trip Purposes for trips along SR1 Corridor')
```


Let us find count observations by trip purpose for trips in the SR1 corridor
using weights:

```{r sr1-19,}
## Generate frequency for trip purpose
temp1<-as.data.frame(table(sr1_act$Actpur))
## Create percentage of frequency
temp1$Percent<-temp1$Freq*100/sum(temp1$Freq)
## Names 
names(temp1)
## Rename the variables
colnames(temp1)[c(1:3)] <- c("Apurp","Ob_UW","Per_UW")
## Show the results
temp1
```

Calculate weighted trip purpose frequencies:

```{r sr1-20,}
## Create weighted frequency
temp2<-aggregate(NEXPTCFPERWGT~Actpur,FUN=sum,data=sr1_act)
## Create weighted percentage
temp2$Percent<-temp2$NEXPTCFPERWGT*100/sum(temp2$NEXPTCFPERWGT)
## Names
names(temp2)
## Rename the variables
colnames(temp2)[c(1:3)] <- c("Apurp","Ob_W","Per_W")
## Show
temp2
```

And save the results:

```{r sr1-21,}
## Merge the two files and save
temp3<-merge(temp1,temp2, by ="Apurp", all = TRUE,sort=FALSE)
temp3 <- temp3[,c(1,2,4,3,5)]
temp3
temp3[,1]<-as.character(temp3[,1])
temp3<- rbind(temp3, c("Total", colSums(temp3[,2:5],na.rm = TRUE)))
temp3
## Save as csv file
write.csv(temp3,file="Apurp_sr1.csv",row.names=FALSE)
rm(temp1)
rm(temp2)
rm(temp3)
```


```{r results='asis',echo=FALSE,cache=FALSE}
############################################################ NESTABLE SECTION
SEC('##','Work Trips by Mode')
```


Task 7: Work trips by Mode for the SR1 Corridor.

First, let us reuse the merged activity table from above for trips
activities (and associated trips), and subset it for work trips only.

```{r sr1-22,}
# Create a subset for work trips
actw_sr1<-subset(sr1_act,APURP==9)
head(actw_sr1$Mode)
```

Now, we can find the number of work trips for each mode:

```{r sr1-23,}
## Generate frequency for trip Mode
temp1<-as.data.frame(table(actw_sr1$Mode))
## Create percentage of frequency
temp1$Percent<-temp1$Freq*100/sum(temp1$Freq)
## Names 
names(temp1)
## Rename the variables
colnames(temp1)[c(1:3)] <- c("Mode","Ob_UW","Per_UW")
## Show the results
temp1
```

Let us expand these numbers with the countywide weights:

```{r sr1-24,}
## Create weighted frequency
temp2<-aggregate(NEXPTCFPERWGT~Mode,FUN=sum,data=actw_sr1)
## Create weighted percentage
temp2$Percent<-temp2$NEXPTCFPERWGT*100/sum(temp2$NEXPTCFPERWGT)
## Names
names(temp2)
## Rename the variables
colnames(temp2)[c(1:3)] <- c("Mode","Ob_W","Per_W")
## Show
temp2
```

Finally, save the results.

```{r sr1-25,}
## Merge the two files and save
temp3<-merge(temp1,temp2, by ="Mode", all = TRUE,sort=FALSE)
temp3 <- temp3[,c(1,2,4,3,5)]
temp3
temp3[,1]<-as.character(temp3[,1])
temp3<- rbind(temp3, c("Total", colSums(temp3[,2:5],na.rm = TRUE)))
temp3
## Save as csv file
write.csv(temp3,file="Mode_wact_SR1.csv",row.names=FALSE)
rm(temp1)
rm(temp2)
rm(temp3)
```

This completes this sub-section.