---
title: "Travel Statistics along the SR-17 Corridor"
author: "Suman Mitra"
date: "September 4, 2015"
output:
  html_document: default
  pdf_document:
    highlight: tango
---

```{r sr17-setup-0,echo=FALSE,cache=FALSE}
if ( !exists("rd") ) { rd <<- gsub("(.*?chts-book[^/]*).*$","\\1",getwd()) } 
source(paste(rd,"R/setup-hooks.R",sep="/"))
write(paste("Using Root Directory",rd,"\n"),stderr())
```


# Summary and Methodology

In this subsection, we show how 2012 CHTS data can be used to generate
useful statistics in the SR-17 corridor. This is another illustration
of how to put the 2012 CHTS data for Caltrans. Steps of this analysis
are similar to those for the analysis of the SR-1. We found 74
household respondents (177 person respondents) including 8 GPS
respondents within a 1 km buffer on each side of the SR-17. To
delineate the SR-17 corridor, we relied on Caltrans’ definition. We
generated tables by mode for: total number of trips, total trip
duration, average trip duration, total trip distance, and average trip
distance.  We also generated tables for the total number of trips by
trip purpose and for work trips by mode. For these, we generated
tables for total number of trips by activity purpose since trip
purpose is given in the data as activity purpose and there is more
than one activity for many single trips. We considered each activity
as a trip. For work trips, we only considered work/job
duties\ifmydocumentation (purpose code `APURP=9`)\fi. We generated
every table for both weighted and unweighted results. We used
countywide weights to generate weighted results.


\ifmydocumentation

# R Code and Implementation

To compute each of the seven tables outlined above, we need the place
file with the new countywide person weights that were computed
previously.

```{r sr17-1, }
#Import the place level with new weight csv file and give a new name
#nplace<- read.csv(pd("Weight/NewWeightData/nplacedata3032015.csv"), header=TRUE)
nplace<- read.csv(pd("Weight/NewWeightData/nplacedata932015.csv"), header=TRUE)
```

Next, we convert the `MODE` codes into text descriptions:

```{r sr17-2, } 
# For recoding Mode 
nplace$Mode <-
  ifelse(nplacedata$MODE==1,"Walk",
  ifelse(nplacedata$MODE==2,"Bike",
  ifelse(nplacedata$MODE==3,"Wheelchair/Mobility Scooter",
  ifelse(nplacedata$MODE==4,"Other Non-Motorized",
  ifelse(nplacedata$MODE==5,"Auto/Van/Truck Driver",
  ifelse(nplacedata$MODE==6,"Auto /Van/Truck Passenger",
  ifelse(nplacedata$MODE==7,"Carpool/Vanpool",
  ifelse(nplacedata$MODE==8,"Motorcycle/Scooter/Moped",
  ifelse(nplacedata$MODE==9,"Taxi/Hired Car/Limo",
  ifelse(nplacedata$MODE==10,"Rental Car/Vehicle",
  ifelse(nplacedata$MODE==11,"Private shuttle ",
  ifelse(nplacedata$MODE==12,"Greyhound Bus",
  ifelse(nplacedata$MODE==13,"Plane",
  ifelse(nplacedata$MODE==14,"Other Private Transit",
  ifelse(nplacedata$MODE==15,"Local Bus,Rapid Bus",
  ifelse(nplacedata$MODE==16,"Express Bus/Commuter Bus",
  ifelse(nplacedata$MODE==17,"Premium Bus",
  ifelse(nplacedata$MODE==18,"School Bus",
  ifelse(nplacedata$MODE==19,"Public Transit Shuttle",
  ifelse(nplacedata$MODE==20,"Air BART/LAX Fly Away",
  ifelse(nplacedata$MODE==21,"Dial-a-Ride/Paratransit",
  ifelse(nplacedata$MODE==22,"Amtrak Bus",
  ifelse(nplacedata$MODE==23,"Other Bus ",
  ifelse(nplacedata$MODE==24,"BART,Metro Red/Purple Line",
  ifelse(nplacedata$MODE==25,"ACE, Amtrak, Cal train",
  ifelse(nplacedata$MODE==26,"Metro",
  ifelse(nplacedata$MODE==27,"Street Car/Cable Car",
  ifelse(nplacedata$MODE==28,"Other Rail",
  ifelse(nplacedata$MODE==29,"Ferry / Boat",
         "Don't Know/Refused"
         )))))))))))))))))))))))))))))
```

We found 74 household respondents (177 person respondents) including 8
GPS respondents within 1 km on each side of the SR 17 using ArcGis and
we saved it in a 'csv' file. Let us import that file:

```{r sr17-3, }
#Import SR17 sample households (obtained from GIS) 
sr17 <- read.csv(pd("District 4/SR 17/SR17_block1km.csv"), header=TRUE)
```

And, let us merge it with the place file:

```{r sr17-4, }
# Merge the SR1 sample data with place level data
sr17_place<-merge(nplace,sr17, by ="SAMPN", all = FALSE)

# Check the Length of the sampn number of new file (it should be 269)
length(unique(sr17_place$SAMPN))
names(sr17_place)
head(sr17_place)
```


## Modal split for SR 17 corridor


Task 1: Let us generate trip frequencies by mode:

```{r sr17-5, }
## Generate frequency for trip Mode
temp1<-as.data.frame(table(sr17_place$Mode))
temp1
## Create percentage of frequency
temp1$Percent<-temp1$Freq*100/sum(temp1$Freq)
temp1
## Names 
names(temp1)
## Rename the variables
colnames(temp1)[c(1:3)] <- c("Mode","Ob_UW","Per_UW")
## Show the results
temp1
```

Now, compute weighted trip frequencies:

```{r sr17-6, }
## Create weighted frequency
temp2<-aggregate(NEXPTCFPERWGT~Mode,FUN=sum,data=sr17_place)
## Create weighted percentage
temp2$Percent<-temp2$NEXPTCFPERWGT*100/sum(temp2$NEXPTCFPERWGT)
## Names
names(temp2)
## Rename the variables
colnames(temp2)[c(1:3)] <- c("Mode","Ob_W","Per_W")
## Show
temp2
```

Finally, save these results in a `csv` file for later use:

```{r sr17-7, }
## Merge the two files and save
temp3<-merge(temp1,temp2, by ="Mode", all = TRUE,sort=FALSE)
temp3 <- temp3[,c(1,2,4,3,5)]
temp3
Mode_sr17<-temp3
temp3[,1]<-as.character(temp3[,1])
temp3<- rbind(temp3, c("Total", colSums(temp3[,2:5],na.rm = TRUE)))
temp3
## Save as csv file
write.csv(temp3,file=pto("Mode_sr17.csv"),row.names=FALSE)
mode.sr17 <- temp3
rm(temp1)
rm(temp2)
rm(temp3)
```


## Total Trip duration for all trips by mode for SR 17 corridor


Task 2: Let us apply a similar procedure to find Total Trip duration for all trips, by mode:

```{r sr17-8, }
## Generate frequency for trip Mode
temp1<-aggregate(TRIPDUR~Mode,FUN=sum,data=sr17_place)
## Create  percentage
temp1$Percent<-temp1$TRIPDUR*100/sum(temp1$TRIPDUR)
## Names
names(temp1)
## Rename the variables
colnames(temp1)[c(1:3)] <- c("Mode","TDu_UW","Per_UW")
## Show
temp1
```

Compute weighted frequencies:

```{r sr17-9, }
## Create weighted frequency
temp<-sr17_place$NEXPTCFPERWGT*sr17_place$TRIPDUR
temp2<-aggregate(temp~Mode,FUN=sum,data=sr17_place)
## Create weighted percentage
temp2$Percent<-temp2$temp*100/sum(temp2$temp)
## Names
names(temp2)
## Rename the variables
colnames(temp2)[c(1:3)] <- c("Mode","TDu_W","Per_W")
## Show
temp2
```


Task 3: Let us find average trip durations:

```{r sr17-10,}
## Merge the two files and save
temp3<-merge(temp1,temp2, by ="Mode", all = TRUE)
temp3 <- temp3[,c(1,2,4,3,5)]
temp3
# Average trip duration
temp3$avg_uw<-temp3$TDu_UW/Mode_sr17$Ob_UW
temp3
temp3$avg_w<-temp3$TDu_W/Mode_sr17$Ob_W
temp3
temp3[,1]<-as.character(temp3[,1])
temp3<- rbind(temp3, c("Total", colSums(temp3[,2:7],na.rm = TRUE)))
temp3
```

Now, save the results for later use:

```{r sr17-11,}
## Save as csv file
write.csv(temp3,file=pto("TripDUrm_sr17.csv"),row.names=FALSE)
dur.sr17 <- temp3
rm(temp1)
rm(temp2)
rm(temp3)
rm(temp)
```


### Trip distance by mode


Task 4: Total Trip distance, by mode.


```{r sr17-12,}
## Generate frequency for trip Mode
temp1<-aggregate(TripDistance~Mode,FUN=sum,data=sr17_place)
## Create  percentage
temp1$Percent<-temp1$TripDistance*100/sum(temp1$TripDistance)
## Names
names(temp1)
## Rename the variables
colnames(temp1)[c(1:3)] <- c("Mode","TDis_UW","Per_UW")
## Show
temp1
```

Now compute weighted frequencies:

```{r sr17-13,}
## Create weighted frequency
temp<-sr17_place$NEXPTCFPERWGT*sr17_place$TripDistance
temp2<-aggregate(temp~Mode,FUN=sum,data=sr17_place)
## Create weighted percentage
temp2$Percent<-temp2$temp*100/sum(temp2$temp)
## Names
names(temp2)
## Rename the variables
colnames(temp2)[c(1:3)] <- c("Mode","TDis_W","Per_W")
## Show
temp2
```

Task 5: Average trip distances:

```{r sr17-14,}
## Merge the two files and save
temp3<-merge(temp1,temp2, by ="Mode", all = TRUE)
temp3 <- temp3[,c(1,2,4,3,5)]
temp3
# Average trip distance
temp3$avg_uw<-temp3$TDis_UW/Mode_sr17$Ob_UW
temp3
temp3$avg_w<-temp3$TDis_W/Mode_sr17$Ob_W
temp3
temp3[,1]<-as.character(temp3[,1])
temp3<- rbind(temp3, c("Total", colSums(temp3[,2:7],na.rm = TRUE)))
temp3
```

Save the results for later use:

```{r sr17-15,}
## Save as csv file
write.csv(temp3,file=pto("TripDistanceM_sr17.csv"),row.names=FALSE)
dist.sr17 <- temp3
rm(temp1)
rm(temp2)
rm(temp3)
rm(temp)
```


## Trip Purpose Calculations


Task 6: Trip purpose breakdown. For this task, we need to import activity level data with appropriate weights.
Let us load and merge these two datasets:

```{r sr17-16,}
#Import the activity level variables dataframe and give a new name
library(chts2011)
act_data <- chts_act
names(act_data)
# Merge the person weight file with place level data file
act_place<-merge(act_data,nplace, by =c("SAMPN","PERNO","PLANO"), all = FALSE)
names(act_place)
```

For display purposes, we convert the purpose codes into readable strings.

```{r sr17-17,}
# Recoding activity purpose
act_place$Actpur <-
  ifelse(
    act_place$APURP==1," Personal Activities",
  ifelse(
    act_place$APURP==2," Preparing Meals/Eating ",
  ifelse(
    act_place$APURP==3," Hosting Visitors",
  ifelse(
    act_place$APURP==4," Exercise ",
  ifelse(
    act_place$APURP==5," Study/Schoolwork ",
  ifelse(
    act_place$APURP==6," Work For Pay At Home",
  ifelse(
    act_place$APURP==7,"Using Computer/Telephone/Cell",
  ifelse(
    act_place$APURP==8," All Other Activities At My Home ",
  ifelse(
    act_place$APURP==9," Work/Job Duties ",
  ifelse(
    act_place$APURP==10," Training",
  ifelse(
    act_place$APURP==11," Meals At Work ",
  ifelse(
    act_place$APURP==12," Work-Sponsored Social Activities ",
  ifelse(
    act_place$APURP==13," Non-Work Related Activities",
  ifelse(
    act_place$APURP==14," Exercise/Sports ",
  ifelse(
    act_place$APURP==15," Volunteer Work/Activities ",
  ifelse(
    act_place$APURP==16," All Other Work-Related Activities",
  ifelse(
    act_place$APURP==17," In School/Classroom/Laboratory ",
  ifelse(
    act_place$APURP==18," Meals At School/College ",
  ifelse(
    act_place$APURP==19," After School Sports/Physical Activity ",
  ifelse(
    act_place$APURP==20," All Other After SchoolActivities ",
  ifelse(
    act_place$APURP==21," Change Type Of Transportation/Transfer ", 
  ifelse(
    act_place$APURP==22," Pickup/Drop Off Passenger(S) ", 
  ifelse(
    act_place$APURP==23," Drive Through Meals", 
  ifelse(
    act_place$APURP==24," Drive Through Other (ATM, Bank) ",
  ifelse(
    act_place$APURP==25," Work-Related",
  ifelse(
    act_place$APURP==26," Service Private Vehicle",
  ifelse(
    act_place$APURP==27," Routine Shopping ",
  ifelse(
    act_place$APURP==28," Shopping For Major Purchases",
  ifelse(
    act_place$APURP==29," Household Errands", 
  ifelse(
    act_place$APURP==30," Personal Business", 
  ifelse(
    act_place$APURP==31," Eat Meal at Restaurant/Diner", 
  ifelse(
    act_place$APURP==32," Health Care", 
  ifelse(
    act_place$APURP==33," Civic/Religious Activities", 
  ifelse(
    act_place$APURP==34," Outdoor Exercise", 
  ifelse(
    act_place$APURP==35," Indoor Exercise ", 
  ifelse(
    act_place$APURP==36," Entertainment ", 
  ifelse(
    act_place$APURP==37," Social/Visit Friends/Relatives", 
  ifelse(
    act_place$APURP==38," Other ", 
  ifelse(
    act_place$APURP==39," Loop Trip", " Don't Know/Refused"
  )))))))))))))))))))))))))))))))))))))))
```

Now, merge the these data with the SR 17 data table:

```{r sr17-18,}
# Merge the sr17 sample data with activity level data
sr17_act<-merge(act_place,sr17, by ="SAMPN", all = FALSE)

# Check the Length of the sampn number of new file (it should be 269)
length(unique(sr17_act$SAMPN))
names(sr17_act)
head(sr17_act)
```

## Trip Purposes for trips along SR 17 Corridor


Let us compute weighted trip purpose results for the SR-17 corridor based on new county weights:

```{r sr17-19,}
## Generate frequency for trip purpose
temp1<-as.data.frame(table(sr17_act$Actpur))
## Create percentage of frequency
temp1$Percent<-temp1$Freq*100/sum(temp1$Freq)
## Names 
names(temp1)
## Rename the variables
colnames(temp1)[c(1:3)] <- c("Apurp","Ob_UW","Per_UW")
## Show the results
temp1
```

Next, let us find weighted trip purpose frequencies:

```{r sr17-20,}
## Create weighted frequency
temp2<-aggregate(NEXPTCFPERWGT~Actpur,FUN=sum,data=sr17_act)
## Create weighted percentage
temp2$Percent<-temp2$NEXPTCFPERWGT*100/sum(temp2$NEXPTCFPERWGT)
## Names
names(temp2)
## Rename the variables
colnames(temp2)[c(1:3)] <- c("Apurp","Ob_W","Per_W")
## Show
temp2
```

And here we save the results in a csv file:

```{r sr17-21,}
## Merge the two files and save
temp3<-merge(temp1,temp2, by ="Apurp", all = TRUE,sort=FALSE)
temp3 <- temp3[,c(1,2,4,3,5)]
temp3
temp3[,1]<-as.character(temp3[,1])
temp3<- rbind(temp3, c("Total", colSums(temp3[,2:5],na.rm = TRUE)))
temp3
## Save as csv file
write.csv(temp3,file=pto("Apurp_sr17.csv"),row.names=FALSE)
apurp.sr17 <- temp3
rm(temp1)
rm(temp2)
rm(temp3)
```


## Work Trips by Mode


Task 7. Finally, let us calculate the number of work trips by Mode for the SR-17 Corridor.

First, we reuse the merged activity table from above for trips activities (and associated trips), and subset it for work trips only:

```{r sr17-22,}
# Create a subset for work trips
actw_sr17<-subset(sr17_act,APURP==9)
head(actw_sr17$Mode)
```

Second, we find the nu,ber of work trips by mode: 

```{r sr17-23,}
## Generate frequency for trip Mode
temp1<-as.data.frame(table(actw_sr17$Mode))
## Create percentage of frequency
temp1$Percent<-temp1$Freq*100/sum(temp1$Freq)
## Names 
names(temp1)
## Rename the variables
colnames(temp1)[c(1:3)] <- c("Mode","Ob_UW","Per_UW")
## Show the results
temp1
```

We can now expand these results using the new countywide weights:

```{r sr17-24,}
## Create weighted frequency
temp2<-aggregate(NEXPTCFPERWGT~Mode,FUN=sum,data=actw_sr17)
## Create weighted percentage
temp2$Percent<-temp2$NEXPTCFPERWGT*100/sum(temp2$NEXPTCFPERWGT)
## Names
names(temp2)
## Rename the variables
colnames(temp2)[c(1:3)] <- c("Mode","Ob_W","Per_W")
## Show
temp2
```

Finally, let us save our results:

```{r sr17-25,}
## Merge the two files and save
temp3<-merge(temp1,temp2, by ="Mode", all = TRUE,sort=FALSE)
temp3 <- temp3[,c(1,2,4,3,5)]
temp3
temp3[,1]<-as.character(temp3[,1])
temp3<- rbind(temp3, c("Total", colSums(temp3[,2:5],na.rm = TRUE)))
temp3
## Save as csv file
write.csv(temp3,file=pto("Mode_wact_sr17.csv"),row.names=FALSE)
mode.wact.sr17 <- temp3
rm(temp1)
rm(temp2)
rm(temp3)
```
\fi

# Results



```{r sr17-res-mode,echo=FALSE,results='asis'}
mode.sr17 <- mode.sr17 %>%
    arrange(ifelse(Mode=="Total",1,0),desc(as.numeric(Ob_W))) %>%
    transmute(`Mode` = Mode,
              `Observed Unweighted` = obsfmt(`Ob_UW`),
              `Observed Weighted` = obsfmt(`Ob_W`),
              `Percent Unweighted` = pctfmt(`Per_UW`),
              `Percent Weighted` = pctfmt(`Per_W`)
              )
pandoc.table(
    mode.sr17,
    caption="Modal split through the SR17 corridor",
    justify=c('left','right','right','right','right'),
    split.tables=100,
    split.cells=c('40%','15%','15%','15%','15%'),
    big.mark=",",
    emphasize.strong.rows=which(mode.sr17$Mode=='Total',arr.ind=TRUE)
)

```


```{r sr17-res-dur,echo=FALSE,results='asis'}
dur.sr17 <- dur.sr17 %>%
    arrange(ifelse(Mode=="Total",1,0),desc(as.numeric(TDu_W))) %>%
    transmute(`Mode` = Mode,
              `Observed Unweighted` = obsfmt(`TDu_UW`),
              `Observed Weighted` = obsfmt(`TDu_W`),
              `Percent Unweighted` = pctfmt(`Per_UW`),
              `Percent Weighted` = pctfmt(`Per_W`)
              )
pandoc.table(
    dur.sr17,
    caption="Total trip duration of trips through the SR17 corridor",
    justify=c('left','right','right','right','right'),
    split.tables=100,
    split.cells=c('40%','15%','15%','15%','15%'),
    big.mark=",",
    emphasize.strong.rows=which(dur.sr17$Mode=='Total',arr.ind=TRUE)
)

```


```{r sr17-res-dist,echo=FALSE,results='asis'}
dist.sr17 <- dist.sr17 %>%
    arrange(ifelse(Mode=="Total",1,0),desc(as.numeric(TDis_W))) %>%
    transmute(`Mode` = Mode,
              `Observed Unweighted` = obsfmt(`TDis_UW`),
              `Observed Weighted` = obsfmt(`TDis_W`),
              `Percent Unweighted` = pctfmt(`Per_UW`),
              `Percent Weighted` = pctfmt(`Per_W`)
              )
pandoc.table(
    dist.sr17,
    caption="Total trip distance of trips through the SR17 corridor",
    justify=c('left','right','right','right','right'),
    split.tables=100,
    split.cells=c('40%','15%','15%','15%','15%'),
    big.mark=",",
    emphasize.strong.rows=which(dist.sr17$Mode=='Total',arr.ind=TRUE)
)

```



```{r sr17-res-apurp,echo=FALSE,results='asis'}
apurp.sr17 <- apurp.sr17 %>%
    arrange(ifelse(Apurp=="Total",1,0),desc(as.numeric(Ob_W))) %>%
    transmute(`Purpose` = Apurp,
              `Observed Unweighted` = obsfmt(`Ob_UW`),
              `Observed Weighted` = obsfmt(`Ob_W`),
              `Percent Unweighted` = pctfmt(`Per_UW`),
              `Percent Weighted` = pctfmt(`Per_W`)
              )
pandoc.table(
    apurp.sr17,
    caption="Total purpose splits for trips through the SR17 corridor",
    justify=c('left','right','right','right','right'),
    split.tables=100,
    split.cells=c('40%','15%','15%','15%','15%'),
    big.mark=",",
    emphasize.strong.rows=which(apurp.sr17$Purpose=='Total',arr.ind=TRUE)
)

```



```{r sr17-res-mode-wact,echo=FALSE,results='asis'}
mode.wact.sr17 <- mode.wact.sr17 %>%
    arrange(ifelse(Mode=="Total",1,0),desc(as.numeric(Ob_W))) %>%
    transmute(`Mode` = Mode,
              `Observed Unweighted` = obsfmt(`Ob_UW`),
              `Observed Weighted` = obsfmt(`Ob_W`),
              `Percent Unweighted` = pctfmt(`Per_UW`),
              `Percent Weighted` = pctfmt(`Per_W`)
              )
pandoc.table(
    mode.wact.sr17,
    caption="Travel mode for work trips through the SR17 corridor",
    justify=c('left','right','right','right','right'),
    split.tables=100,
    split.cells=c('40%','15%','15%','15%','15%'),
    big.mark=",",
    emphasize.strong.rows=which(mode.wact.sr17$Mode=='Total',arr.ind=TRUE)
)

```
