---
title: "6.8 Travel Statistics along the SR-17 Corridor"
author: "Suman Mitra"
date: "September 4, 2015"
output: html_document
---

## Summary and Methodology

We found 74 household respondents (177 person respondents) including 8 GPS respondents within 1 km buffer of each side of sr177 (see Map 1). We took the same length of SR 17 as it is in the TCR for SR 17 (provided by Caltrans). We generated tables by mode for the following: total number of trips, total trip distance, average trip distance, total trip duration, average trip duration.  We also generated tables for the total number of trips by trip purpose and for work trips by mode. Here, we actually generated tables for total number of trips by activity purpose since the trip purpose is given in the database as activity purpose and there is more than one activity in many single trips. We considered each activity as a trip. We only considered the purpose code (APURP) 9 for work trips.  We generated every table for both weighted and unweighted results. We used county wide weight to generate weighted results. 


## R Code and Implementation

The following commands are used to generate the results.

To answer the questions we need the place file with new county wide person weight that we computed in the section 4.1.

```{r sr17-1, }
rd <- "/home/crindt/chts-book"
p <- function(ps) { return(paste(rd,ps,sep="/")) }

#Import the place level with new weight csv file and give a new name
nplace<- read.csv(p("Weight/NewWeightData/nplacedata3032015.csv"), header=TRUE)
```

Next, we  convert the `MODE` codes into text descriptions.

```{r sr17-2, }
# For recoding Mode
nplace$Mode<-ifelse(nplace$MODE==1,"Walk",ifelse(nplace$MODE==2,"Bike",ifelse(nplace$MODE==3,"Wheelchair/Mobility Scooter",ifelse(nplace$MODE==4,"Other Non-Motorized",ifelse(nplace$MODE==5,"Auto/Van/Truck Driver",ifelse(nplace$MODE==6,"Auto /Van/Truck Passenger",ifelse(nplace$MODE==7,"Carpool/Vanpool",ifelse(nplace$MODE==8,"Motorcycle",ifelse(nplace$MODE==9,"Taxi/Hired Car/Limo",ifelse(nplace$MODE==10,"Rental Car/Vehicle",ifelse(nplace$MODE==11,"Private shuttle ",ifelse(nplace$MODE==12,"Greyhound Bus",ifelse(nplace$MODE==13,"Plane",ifelse(nplace$MODE==14,"Other Private Transit",ifelse(nplace$MODE==15,"Local Bus,Rapid Bus",ifelse(nplace$MODE==16,"Express Bus",ifelse(nplace$MODE==17,"Premium Bus",ifelse(nplace$MODE==18,"School Bus",ifelse(nplace$MODE==19,"Public Transit Shuttle",ifelse(nplace$MODE==20,"Air BART/LAX Fly Away",ifelse(nplace$MODE==21,"Dial-a-Ride/Paratransit", ifelse(nplace$MODE==22,"Amtrak Bus", ifelse(nplace$MODE==23,"Other Bus ", ifelse(nplace$MODE==24,"BART,Metro",ifelse(nplace$MODE==25,"ACE, Amtrak etc",ifelse(nplace$MODE==26,"Metro",ifelse(nplace$MODE==27,"Street Car",ifelse(nplace$MODE==28,"Other Rail",ifelse(nplace$MODE==29,"Ferry / Boat","Don't Know/Refused")))))))))))))))))))))))))))))
```

We  found 74 household respondents (177 person respondents) including 8 GPS respondents within 1 km on each side of the SR 17 using ArcGis and we saved it in a 'csv' file. Now we need to import that file. 

```{r sr17-3, }
#Import SR17 sample households (obtained from GIS) 
sr17 <- read.csv(p("District 4/SR 17/SR17_block1km.csv"), header=TRUE)
```

And, we merge this file with place file.

```{r sr17-4, }
# Merge the SR1 sample data with place level data
sr17_place<-merge(nplace,sr17, by ="SAMPN", all = FALSE)

# Check the Length of the sampn number of new file (it should be 269)
length(unique(sr17_place$SAMPN))
names(sr17_place)
head(sr17_place)
```


### Modal split for SR 17 corridor

First, we generate trip frequency by mode.

```{r sr17-5, }
## Generate frequency for trip Mode
temp1<-as.data.frame(table(sr17_place$Mode))
temp1
## Create percentage of frequency
temp1$Percent<-temp1$Freq*100/sum(temp1$Freq)
temp1
## Names 
names(temp1)
## Rename the variables
colnames(temp1)[c(1:3)] <- c("Mode","Ob_UW","Per_UW")
## Show the results
temp1
```

Then, we compute the weighted trip frequency.

```{r sr17-6, }
## Create weighted frequency
temp2<-aggregate(NEXPTCFPERWGT~Mode,FUN=sum,data=sr17_place)
## Create weighted percentage
temp2$Percent<-temp2$NEXPTCFPERWGT*100/sum(temp2$NEXPTCFPERWGT)
## Names
names(temp2)
## Rename the variables
colnames(temp2)[c(1:3)] <- c("Mode","Ob_W","Per_W")
## Show
temp2
```

Now we can save these results in a `csv` file for later use.


```{r sr17-7, }
## Merge the two files and save
temp3<-merge(temp1,temp2, by ="Mode", all = TRUE,sort=FALSE)
temp3 <- temp3[,c(1,2,4,3,5)]
temp3
Mode_sr17<-temp3
temp3[,1]<-as.character(temp3[,1])
temp3<- rbind(temp3, c("Total", colSums(temp3[,2:5],na.rm = TRUE)))
temp3
## Save as csv file
write.csv(temp3,file="Mode_sr17.csv",row.names=FALSE)
rm(temp1)
rm(temp2)
rm(temp3)
```


### Total Trip duration for all trips by mode for SR 17 corridor

Next, we perform a similar computation for Total Trip duration for all trips for SR 17 corridor, by mode.

```{r sr17-8, }
## Generate frequency for trip Mode
temp1<-aggregate(TRIPDUR~Mode,FUN=sum,data=sr17_place)
## Create  percentage
temp1$Percent<-temp1$TRIPDUR*100/sum(temp1$TRIPDUR)
## Names
names(temp1)
## Rename the variables
colnames(temp1)[c(1:3)] <- c("Mode","TDu_UW","Per_UW")
## Show
temp1
```

Again, we compute the weighted frequency.

```{r sr17-9, }
## Create weighted frequency
temp<-sr17_place$NEXPTCFPERWGT*sr17_place$TRIPDUR
temp2<-aggregate(temp~Mode,FUN=sum,data=sr17_place)
## Create weighted percentage
temp2$Percent<-temp2$temp*100/sum(temp2$temp)
## Names
names(temp2)
## Rename the variables
colnames(temp2)[c(1:3)] <- c("Mode","TDu_W","Per_W")
## Show
temp2
```

Now, we compute the average trip duration.

```{r sr17-10,}
## Merge the two files and save
temp3<-merge(temp1,temp2, by ="Mode", all = TRUE)
temp3 <- temp3[,c(1,2,4,3,5)]
temp3
# Average trip duration
temp3$avg_uw<-temp3$TDu_UW/Mode_sr17$Ob_UW
temp3
temp3$avg_w<-temp3$TDu_W/Mode_sr17$Ob_W
temp3
temp3[,1]<-as.character(temp3[,1])
temp3<- rbind(temp3, c("Total", colSums(temp3[,2:7],na.rm = TRUE)))
temp3
```

And as before, we save the results for later use.

```{r sr17-11,}
## Save as csv file
write.csv(temp3,file="TripDUrm_sr17.csv",row.names=FALSE)
rm(temp1)
rm(temp2)
rm(temp3)
rm(temp)
```


#### Trip distance by mode

Moving on, we determine Total Trip distance for SR 17 Corridor, by mode.


```{r sr17-12,}
## Generate frequency for trip Mode
temp1<-aggregate(TripDistance~Mode,FUN=sum,data=sr17_place)
## Create  percentage
temp1$Percent<-temp1$TripDistance*100/sum(temp1$TripDistance)
## Names
names(temp1)
## Rename the variables
colnames(temp1)[c(1:3)] <- c("Mode","TDis_UW","Per_UW")
## Show
temp1
```

Again, we compute the weighted frequency.

```{r sr17-13,}
## Create weighted frequency
temp<-sr17_place$NEXPTCFPERWGT*sr17_place$TripDistance
temp2<-aggregate(temp~Mode,FUN=sum,data=sr17_place)
## Create weighted percentage
temp2$Percent<-temp2$temp*100/sum(temp2$temp)
## Names
names(temp2)
## Rename the variables
colnames(temp2)[c(1:3)] <- c("Mode","TDis_W","Per_W")
## Show
temp2
```

Now, we compute the average trip distance.

```{r sr17-14,}
## Merge the two files and save
temp3<-merge(temp1,temp2, by ="Mode", all = TRUE)
temp3 <- temp3[,c(1,2,4,3,5)]
temp3
# Average trip distance
temp3$avg_uw<-temp3$TDis_UW/Mode_sr17$Ob_UW
temp3
temp3$avg_w<-temp3$TDis_W/Mode_sr17$Ob_W
temp3
temp3[,1]<-as.character(temp3[,1])
temp3<- rbind(temp3, c("Total", colSums(temp3[,2:7],na.rm = TRUE)))
temp3
```

And, we save the results for later use.

```{r sr17-15,}
## Save as csv file
write.csv(temp3,file="TripDistanceM_sr17.csv",row.names=FALSE)
rm(temp1)
rm(temp2)
rm(temp3)
rm(temp)
```


### Trip Purpose Calculations

Now we need trip purpose breakdown, so we need to import activity level data with the 
appropriate weights.  Here we load and merge them.


```{r sr17-16,}
#Import the activity level variables csv file and give a new name
#act_data <- read.csv("c:/Users/user/Documents/Suman_research/Caltrans Project/Data/deliv_activity.csv", header=TRUE)
library(chts2011)
act_data <- chts_act
names(act_data)
# Merge the person weight file with place level data file
act_place<-merge(act_data,nplace, by =c("SAMPN","PERNO","PLANO"), all = FALSE)
names(act_place)
```

For display purposes, we convert the purpose codes into readable strings.

```{r sr17-17,}
# Recoding activity purpose
act_place$Actpur<-ifelse(act_place$APURP==1," Personal Activities",ifelse(act_place$APURP==2," Preparing Meals/Eating ",ifelse(act_place$APURP==3," Hosting Visitors",ifelse(act_place$APURP==4," Exercise ",ifelse(act_place$APURP==5," Study/Schoolwork ",ifelse(act_place$APURP==6," Work For Pay At Home",ifelse(act_place$APURP==7,"Using Computer/Telephone/Cell",ifelse(act_place$APURP==8," All Other Activities At My Home ",ifelse(act_place$APURP==9," Work/Job Duties ",ifelse(act_place$APURP==10," Training",ifelse(act_place$APURP==11," Meals At Work ",ifelse(act_place$APURP==12," Work-Sponsored Social Activities ",ifelse(act_place$APURP==13," Non-Work Related Activities",ifelse(act_place$APURP==14," Exercise/Sports ",ifelse(act_place$APURP==15," Volunteer Work/Activities ",ifelse(act_place$APURP==16," All Other Work-Related Activities",ifelse(act_place$APURP==17," In School/Classroom/Laboratory ",ifelse(act_place$APURP==18," Meals At School/College ",ifelse(act_place$APURP==19," After School Sports/Physical Activity ",ifelse(act_place$APURP==20," All Other After SchoolActivities ",ifelse(act_place$APURP==21," Change Type Of Transportation/Transfer ", ifelse(act_place$APURP==22," Pickup/Drop Off Passenger(S) ", ifelse(act_place$APURP==23," Drive Through Meals", ifelse(act_place$APURP==24," Drive Through Other (ATM, Bank) ",ifelse(act_place$APURP==25," Work-Related",ifelse(act_place$APURP==26," Service Private Vehicle",ifelse(act_place$APURP==27," Routine Shopping ",ifelse(act_place$APURP==28," Shopping For Major Purchases",ifelse(act_place$APURP==29," Household Errands", ifelse(act_place$APURP==30," Personal Business", ifelse(act_place$APURP==31," Eat Meal at Restaurant/Diner", ifelse(act_place$APURP==32," Health Care", ifelse(act_place$APURP==33," Civic/Religious Activities", ifelse(act_place$APURP==34," Outdoor Exercise", ifelse(act_place$APURP==35," Indoor Exercise ", ifelse(act_place$APURP==36," Entertainment ", ifelse(act_place$APURP==37," Social/Visit Friends/Relatives", ifelse(act_place$APURP==38," Other ", ifelse(act_place$APURP==39," Loop Trip", " Don't Know/Refused")))))))))))))))))))))))))))))))))))))))
```

Now, merge the this data with SR 17 data table.

```{r sr17-18,}
# Merge the sr17 sample data with activity level data
sr17_act<-merge(act_place,sr17, by ="SAMPN", all = FALSE)

# Check the Length of the sampn number of new file (it should be 269)
length(unique(sr17_act$SAMPN))
names(sr17_act)
head(sr17_act)
```

### Trip Purposes for trips along SR 17 Corridor

We compute the observations for trip purpose for trips of SR 17 corridor based on new weights.

```{r sr17-19,}
## Generate frequency for trip purpose
temp1<-as.data.frame(table(sr17_act$Actpur))
## Create percentage of frequency
temp1$Percent<-temp1$Freq*100/sum(temp1$Freq)
## Names 
names(temp1)
## Rename the variables
colnames(temp1)[c(1:3)] <- c("Apurp","Ob_UW","Per_UW")
## Show the results
temp1
```

This is followed by the weighted trip purpose frequencies.

```{r sr17-20,}
## Create weighted frequency
temp2<-aggregate(NEXPTCFPERWGT~Actpur,FUN=sum,data=sr17_act)
## Create weighted percentage
temp2$Percent<-temp2$NEXPTCFPERWGT*100/sum(temp2$NEXPTCFPERWGT)
## Names
names(temp2)
## Rename the variables
colnames(temp2)[c(1:3)] <- c("Apurp","Ob_W","Per_W")
## Show
temp2
```

And here we save the results.

```{r sr17-21,}
## Merge the two files and save
temp3<-merge(temp1,temp2, by ="Apurp", all = TRUE,sort=FALSE)
temp3 <- temp3[,c(1,2,4,3,5)]
temp3
temp3[,1]<-as.character(temp3[,1])
temp3<- rbind(temp3, c("Total", colSums(temp3[,2:5],na.rm = TRUE)))
temp3
## Save as csv file
write.csv(temp3,file="Apurp_sr17.csv",row.names=FALSE)
rm(temp1)
rm(temp2)
rm(temp3)
```


### Work Trips by Mode

We want to generate work trips by Mode for  SR 17 Corridor.

First, we reuse the merged activity table from above for trips activities (and associated trips), and subset it for work trips only.

```{r sr17-22,}
# Create a subset for work trips
actw_sr17<-subset(sr17_act,APURP==9)
head(actw_sr17$Mode)
```

Now, we can determine the observations of each Mode used for work trips.

```{r sr17-23,}
## Generate frequency for trip Mode
temp1<-as.data.frame(table(actw_sr17$Mode))
## Create percentage of frequency
temp1$Percent<-temp1$Freq*100/sum(temp1$Freq)
## Names 
names(temp1)
## Rename the variables
colnames(temp1)[c(1:3)] <- c("Mode","Ob_UW","Per_UW")
## Show the results
temp1
```

Given those, we can expand them with the county wide weights.

```{r sr17-24,}
## Create weighted frequency
temp2<-aggregate(NEXPTCFPERWGT~Mode,FUN=sum,data=actw_sr17)
## Create weighted percentage
temp2$Percent<-temp2$NEXPTCFPERWGT*100/sum(temp2$NEXPTCFPERWGT)
## Names
names(temp2)
## Rename the variables
colnames(temp2)[c(1:3)] <- c("Mode","Ob_W","Per_W")
## Show
temp2
```

And again, we finish by saving the results.

```{r sr17-25,}
## Merge the two files and save
temp3<-merge(temp1,temp2, by ="Mode", all = TRUE,sort=FALSE)
temp3 <- temp3[,c(1,2,4,3,5)]
temp3
temp3[,1]<-as.character(temp3[,1])
temp3<- rbind(temp3, c("Total", colSums(temp3[,2:5],na.rm = TRUE)))
temp3
## Save as csv file
write.csv(temp3,file="Mode_wact_sr17.csv",row.names=FALSE)
rm(temp1)
rm(temp2)
rm(temp3)
```
