---
title: "2012 California Household Travel Survey Data Analysis"
author: "Craig Rindt"
date: "09/07/2015"
output:
  html_document:
    fig_caption: yes
    highlight: tango
    number_sections: yes
    toc: yes
  pdf_document:
    highlight: tango
    keep_tex: yes
    number_sections: yes
    toc: yes
    toc_depth: 5
---

```{r setup,echo=FALSE,cache=FALSE}
if ( !exists('rd') ) { rd <- '/Users/crindt/chts-book'}
source(paste(rd,'R/setup-hooks.R',sep="/"))
```

# Introduction


FIXME: TO BE ADDED

# Use of R Program for CHTS Data Analysis

FIXME: TO BE COMPLETED

Things to install (windows)

* R 
* R tools
* RStudio
* install.packages("chts2011")
* install.packages("chts2011pvt")
* install.packages("dplyr")
* install.packages("logging")
* install.packages("survey")
* install.packages("RCurl")
* install.packages("jsonlite")
* Change rd variable in chts-book.Rmd to chts-book root directory
* install.packages("chts2011wgt")

Libraries required to generate this document:

* `chts2011`
* `chts2011pvt`
* `dplyr`


# Diagnostic Checks of the CHTS

```{r child = '3.Diagnostic.Rmd'}
```

# Calculating Weights

## County Level Person Weights

```{r child = '4.1Personweights.Rmd'}
```

## Household Level Person Weights

```{r child = '4.2Householdweights.Rmd'}
```

## Weights for Other Geographies

```{r child = '4.3Other Geographies.Rmd'}
```

```{r child = '4.4Longdistanceweight.Rmd'}
```

# Linked Trip Tables

## How to Create the Linked Trip Table

```{r child = 'chts-linked-trips.Rmd'}
```

#  Creating the CHTS Summary Tables

```{r child = 'recreating-2001-tables-sltn.Rmd'}
```


# Data Analysis Requests

The following sections provide R solutions to the questions asked by various districts
during the course of the project.

## Data Request for Santa Cruz County

```{r child = '6.1SantaCruz county.Rmd'}
```

## Data Request for Del Note County

```{r child = '6.2DelNortecounty.Rmd'}
```

## Humboldt for Humboldt County

```{r child = '6.3 Humboldt county.Rmd'}
```

## Data Request for Lake County

```{r child = '6.4 Lake County.Rmd'}
```

## Data Request for Mendocino County

```{r child = '6.5 Mendocino County.Rmd'}
```

## Long Distance Trips Questions

```{r child = '6.6 Long Distance Trips questions.Rmd'}
```

## Travel Statistics along the SR-1 Corridor

```{r child = '6.7 SR1.Rmd'}
```

## Travel Statistics along the SR-17 Corridor

```{r child = '6.8 SR17.Rmd'}
```

# Conclusions

FIXME: To be added.
