---
title: "Analyzing the 2012 California Household Travel Survey using R"
author:
- name: Craig Rindt
- name: Suman Mitra
- name: Jean-Daniel Saphores
  contact: "Contact information: saphores@uci.edu; Phone: (949) 824 7334"
date: 'This version: `r (function() {f<-dir(); format(max(file.mtime(f[grepl("^.*\\.Rmd$",f)])), format="%B %d, %Y")})()`'
institution:
- Institute of Transportation Studies
- University of California, Irvine
output:
  pdf_document:
    highlight: tango
    keep_tex: yes
    number_sections: yes
    pandoc_args: --chapters
    template: custom.latex
    toc: yes
    toc_depth: 5
  html_document:
    fig_caption: yes
    highlight: tango
    number_sections: yes
    toc: yes
fontfamily: inconsolata
fontsize: 11pt
geometry: margin=1in
documentation: '`r if(exists("documentation")){documentation}else{documentation<-"true";documentation}`'
documentclass: '`r if (exists("documentation")) {"report"}else{"article"}`'
---

```{r book-setup-0,echo=FALSE,cache=FALSE}
if ( !exists("rd") ) { rd <<- gsub("(.*?chts-book[^/]*).*$","\\1",getwd()) } 
source(paste(rd,"R/setup-hooks.R",sep="/"))
write(paste("Using Root Directory",rd,"\n"),stderr())
```


# Introduction

The 2010-12 CHTS, which resulted from a statewide, collaborative
effort, enabled the collection of travel information from 42,560
Californian households. This rich dataset has helped update regional
and statewide travel and will help update environmental models.

In 2014, the Institute of Transportation Studies at Irvine (ITS) and
Caltrans initiated the "Enhancing the Value of the 2010-12 California
Household Travel Survey (CHTS)" contract.  This contract was motivated
by the idea that potential value of the CHTS is not always well
understood by Caltrans staff and that some Caltrans staff from the
Office of Travel Forecasting and Analysis may benefit from updating
their knowledge of statistical modeling to comfortably query CHTS data
and to estimate some common transportation econometrics models.

The specific elements of the contract included the following
deliverables:

1. a systematic diagnostic review of the 2010-12 CHTS database for
   unlikely observations;

2. a set of interviews with staff at Caltrans headquarters and visits to three
   selected Caltrans Districts to better understand how they could
   benefit from using 2010-12 CHTS data and to help promote its use in
   their work;

3. hands-on statistical training and consulting to selected Caltrans
   staff in the Office of Travel Forecasting and Analysis in
   Sacramento;

4. on-call statistical support to Caltrans staff from the
   Office of Travel Forecasting and Analysis; and

5. a reference book of useful statistical commands based
   on actual case studies to make it easier to put the 2010-12 CHTS to
   work for Caltrans staff.

The book you are reading now is the last of these deliverables.  It
provides numerous examples of how to perform various types of
statistical analysis on the CHTS.  In chapter 2, we discuss the `R`
language and environment for statistical analysis, which was chosen as
the primary analysis tool for this project.  The following chapters
provide specific examples of statistical analysis taken from the
contract tasks.  In all cases, the actual `R` code used to perform the
analysis is provided, along with detailed discussion of the methods
imployed.  Chapter 3 describes the Task 1 diagnostic review of the
CHTS.  In chapter 4, we demonstrate the computation of statistical
weights for various subpopulations in the CHTS---a critical component
of any analysis involving the CHTS.  In chapter 5, we cover the
creation of a "linked trip" dataset, which provides a means for
analyzing CHTS data in a manner that is compatible with conventional
4-step, trip based models.  In chapter 6, we describe the creation of
CHTS summary tables generated as part of task 3 work.  Finally,
chapter 7 describes the solution of a number of statistical queries
that were answered under task 4 statistical support tasks.



# Getting Started with R and R Studio

```{r getting-started-child,echo=FALSE,results='asis',cache=FALSE}
nestedInclude('getting-started.Rmd',1) 
```

# Diagnostic Checks of the CHTS

```{r diag-child,echo=FALSE,results='asis',cache=FALSE} 
nestedInclude('3.Diagnostic.Rmd',1)
```

# Calculating Weights

## County Level Person Weights


```{r pw-child,echo=FALSE,results='asis',cache=FALSE} 
nestedInclude('4.1Personweights.Rmd',2)
```

## County Level Household Weights


```{r hw-child,echo=FALSE,results='asis',cache=FALSE} 
nestedInclude('4.2Householdweights.Rmd',2)
```


## Weights for Other Geographies

```{r og-child,echo=FALSE,results='asis',cache=FALSE} 
nestedInclude('4.3Other Geographies.Rmd',2)
```

## Weights for Long Distance Trips

```{r ldw-child,echo=FALSE,results='asis',cache=FALSE} 
nestedInclude('4.4Longdistanceweight.Rmd',2)
```

# Linked Trip Tables

```{r lt-child,echo=FALSE,results='asis',cache=FALSE} 
nestedInclude('chts-linked-trips.Rmd',1)
```

#  Creating the CHTS Summary Tables


```{r rt-child,echo=FALSE,results='asis',cache=FALSE} 
nestedInclude('recreating-2001-tables-sltn.Rmd',1)
```

# Data Analysis Requests

The following sections provide R solutions to the questions asked by various districts
during the course of the project.

## Data Request for Santa Cruz County


```{r scc-child,echo=FALSE,results='asis',cache=FALSE} 
nestedInclude('6.1SantaCruz county.Rmd',2)
```


## Data Request for Del Norte County


```{r dnc-child,echo=FALSE,results='asis',cache=FALSE} 
nestedInclude('6.2DelNortecounty.Rmd',2)
```


## Long Distance Trips Questions


```{r ld-child,echo=FALSE,results='asis',cache=FALSE} 
nestedInclude('6.6 Long Distance Trips questions.Rmd',2)
```

## Travel Statistics along the SR-1 Corridor


```{r sr1-child,echo=FALSE,results='asis',cache=FALSE} 
nestedInclude('6.7 SR1.Rmd',2)
```

## Travel Statistics along the SR-17 Corridor


```{r sr17-cild,echo=FALSE,results='asis',cache=FALSE} 
nestedInclude('6.8 SR17.Rmd',2)
```

