---
title: "Getting Started with R and R Studio"
author:
- name: Jean-Daniel Saphores
- name: Craig Rindt
date: '`r file.mtime(knitr::current_input())`'
output:
  pdf_document:
    toc: yes
  html_document:
    highlight: tango
    number_sections: yes
    pandoc_args: --chapters
    theme: journal
    toc: yes
    toc_depth: 6
fontfamily: inconsolata
documentclass: '`r if (exists("documentation")) {"report"}else{"article"}`'
documentation: '`r if(exists("documentation")){documentation}else{documentation<-NULL;documentation}`'
---

```{r getting-started-setup,cache=TRUE,echo=FALSE}
# Set the working directory so we can access libraries
if ( !exists("rd") ) { rd <<- gsub("(.*?chts-book[^/]*).*$","\\1",getwd()) } 
source(paste(rd,"R/setup-hooks.R",sep="/"))
write(paste("Using Root Directory",rd,"\n"),stderr())
```


Following review and discussions with Caltrans staff during the
initial stages of the project a number of decisions were made
regarding how the deliverables would be satisfied.  The most
consequential of these early decisions was the recommendation by ITS
that the `R` language and environment for statistical computing is the
best choice for performing and teaching statistical analysis.  This
recommendation was based upon a number of factors.  First of all, R is
completely free, open-source, and it runs on UNIX, Windows and
Macintosh. Second, it is extremely powerful, and it has an excellent
built-in help system as well as excellent graphing capabilities.
Comparable commerical packages, such as SAS, cost hundreds of
dollars per user license.  Second, R's language has an easy to learn
syntax with many built-in statistical functions, and it is easy to
extend with user-written functions.  Third, R can be made user
friendly thanks to R studio and R commander.  Finally, R has a vibrant
user community that includes academic, governmental, and commercial
users who can and do extend R's capabilities through freely available
add-on packages and numerous on-line help forums.


# Installation of R and R Studio

To work with the examples provided in this book, you need to install
some free, open source software.  All software mentioned here is
available for computers running Microsoft Windows, Apple's MacOS, and
virtually any distribution of Linux.

The only software that is *required* to work these CHTS examples is `R`
itself along with various `R` package libraries that are downloaded
from within `R` itself (we discuss this below).  With that said, our
recommended `R` platform is `RStudio`, for its graphical user
interface and general ease-of-use.  `RStudio` is a "shell" that works
with the `R` system, providing features for editing `R` programs,
interacting with the `R` backend (e.g., running programs), and viewing
the results.  Note that you must install `R` prior to installing
`RStudio`.  You can download the most recent version of `R` and
`RStudio` from the following locations (we note the versions used in
this project in parentheses):

* R (3.2.2): https://cran.rstudio.com/
* RStudio (0.99.489): https://www.rstudio.com/products/rstudio/download/

# Other software

We used a variety of other software in working with data from the
CHTS.  None of these are *required* for any of the analysis this
documentation, but we summarize each below for completeness:

> `pandoc` (http://pandoc.org)
> : a tool for transforming documents between a variety of formats.  In
>   our case, it is used to create documents from `R` source code to
>   illustrate results.  Indeed, the book you are reading was generated
>   using `pandoc`.  Pandoc is installed with `RStudio` and is integrated
>   with `R` using the `knitr` `R` package (see below).  If you are not
>   using RStudio, you will need to install `pandoc` separately in order
>   to generate documentation from `R`.
>
> `mdbtools` (https://github.com/brianb/mdbtools)
> : provides command-line capabilities for working with Microsoft Access
>   databases.  We used mdbtools to export CHTS data from its access
>   distribution database to comma separated values (CSV) files.  This
>   allowed us to easily import data into R and other software such as
>   PostgreSQL.
> 
> `PostgreSQL` (http://www.postgresql.org/)
> : a well regarded open source, SQL database.  We used `PostgreSQL` to
>   perform a variety of data processing tasks early in the project.
>   `PostgreSQL`'s `PostGIS` extension provides powerful Geographic
>   Information System (GIS) capabilities that extend beyond what is
>   possible in `R` alone.
> 
> `git` (http://git-scm.com/)
> : the dominant free and open source distributed version control
>   system.  We used `git` to maintain and share our `R` source code and
>   documentation internally.


# R packages

In addition to the core `R` software, much of the analysis we
performed on the `CHTS` made use of freely available `R` packages
developed and vetted by members of the `R` community.  `R` packages
can be installed via the `packages` tab in `RStudio` or via the
`install.packages` command in `R`.  For instance, to install the
`survey` package used to compute statistical weights, you can simply
type the following at the `R` command prompt (or in the `R` console
window in `RStudio`):

```{r using-r-install-packages,echo=TRUE,eval=FALSE}
install.packages("survey")
```

This will download the `survey` package from the nearest Comprehensive
R Achive Network (CRAN) mirror and install it into your personal `R`
library for use in your analysis.


Other `CRAN` packages we used include:

> `knitr`
> : a package for generating final documentation from intermingled
>   source code and documentation.  This book was created using `knitr`
>   to process documentation written in a combination of `R` source code
>   and the `pandoc` variant of
>   [`markdown`](http://pandoc.org/demo/example9/pandocs-markdown.html)
> 
> `pander`
> : a package for generating `markdown` formatted tables and lists for
>   processing with `knitr`
> 
> `dplyr`
> : a package that provides a convenient filter-style idiom for
>   performing statistical computing tasks on `R` data frames.
> 
> `tidyr`
> : a package that allows for quick transformation of `R` data frames
>   between row- and column-oriented representations.
> 
> `Hmisc`
> : a package providing a wide range of common statistical functions and
>   utilities not included in the core `R` distribution
> 
> `isotone`
> : a package that, among other things, provides a function for
>   computing the weighted median of a categorical variable.


To simplify working with the CHTS in `R`, we also created a set of
custom packages involving the CHTS data set:

> `chts2011`
> : is a package containing the public tables from the CHTS 2010-2012
>   survey.  Once you have installed the `chts2011` package, These
>   tables are accessible as:
> 
>   * `chts_hh`: the household data set
>   * `chts_per`: the person data set
>   * `chts_veh`: the vehicle data set
>   * `chts_pla`: the place data set
>   * `chts_act`: the activity data set
>   * `chts_md`: a table containing CHTS "metadata" as provided by
>     Caltrans staff.  This data includes the variable names,
>     descriptions, and survey questions.
> 
> `chts2011pvt`
> : is a package containing the private tables from the CHTS 2010-2012
>   survey.  These include addresses and geo-coordinates.  This data is
>   privileged so that this package can only be obtained from Caltrans
>   directly and may not be redistributed in any form.
> 
> `chts2011wgt`
> : is a package containing various pre-calculated statistical weights,
>   including person and household weights by county, weekdays, and
>   weekends.



# Generating this document

```{r setup,cache=FALSE,echo=FALSE}
if ( !exists("rd") ) { rd <<- gsub("(.*?chts-book[^/]*).*$","\\1",getwd()) } 
source(paste(rd,'R/setup-hooks.R',sep="/"))
write(paste("Using Root Directory",rd,"\n"),stderr())
```

As noted above, this book was generated directly from the R-markdown
(Rmd) source code using the `knitr` package to create `markdown`
documentation that can then be processed using the `pandoc` program to
generate web pages, presentations, or hardcopy documentation.  The
source code for this document can be obtained from `github` by cloning
the following repository https://github.com/crindt/chts-book or
downloading the zip file
https://github.com/crindt/chts-book/archive/master.zip and unzipping
it.


```{r readme-child,echo=FALSE,results='asis',cache=FALSE}
nestedInclude('../README.Rmd',2) 
```

# Where to find help in R?

There is a wealth of tutorials, articles, and examples to help you
learn R and its extensions.

A starting point for R resources is the Comprehensive R Archive
Network (CRAN), which is located at http://cran.us.r-project.org.  It
is best queried using http://www.rseek.org/ instead of Google.

There are many free tutorials available online. Here are a few:

* Quick R: http://www.statmethods.net/;
* R-Twotorials: http://www.twotorials.com/;
* Getting started with the R Data Analysis Package by Professor Norm
  Matloff at UC Davis: http://heather.cs.ucdavis.edu/~matloff/r.html;
* UCLA Academy Technology Services: http://www.ats.ucla.edu/stat/r/;
  or simply
* http://pairach.com/2012/02/26/r-tutorials-from-universities-around-the-world/,
  which provides web addresses to R tutorials in locations around the
  world.

In addition, there is R-Bloggers (http://www.r-bloggers.com/), which
is a central hub of content collected from bloggers who write about R
(in English).

Another useful resource that relies on R Studio can be found at
https://www.rstudio.com/resources/training/online-learning/

As Professor Matloff from UC Davis said

> "There is a perception among some that R has a steep learning curve,
> but I disagree. True, R usage has its advanced aspects, but my
> recommendation is simply, just get started! Start simple, and then
> refine gradually."

Happy Learning R!

